---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 4
        toc_float: true
        mathjax: default
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  4   
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
TAGS:  ymlthis
---
~/code/MASTER_INDEX.md



### ymlthis::
NOT FINISHED.   yml object means S3(?), list with class = yml.
'yml string'  is string (character vector, length=1) that has CR formattted as "\n" or is written with CR to appear as multiline

ymlthis:: is wrapper yaml:: package, which I have not included here (yet)

Example below:  create yml with the functions ymlthis provides and THEN tack on
custom `TAGS: tag1, tag2`
```

```{r setup, include=FALSE		}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "      ##",
  error = TRUE,
  collapse = T
) # T easier to read
library(tinytest)
library(ymlthis)
library(magrittr)
library(yaml)
```

#### yml object
```{r playing}
## create .yml object
y <- yml()
is_yml(y)
# [1] TRUE
{
  str(y)
  ls.str(y)
  class(y)
  y
}

##  as_yml(.yml) is generic S3 function
methods(as_yml)

##  ad-hoc, create a yml object (here single string, yml string?)
z <- as_yml("
                TAG: tag1, tag2
                ")
is_yml(z)
# [1] TRUE


## append ad-hoc to bottom of existing .yml (y)
args(yml_toplevel)
yml_toplevel(y, z)

##  tree
draw_yml_tree()
```

#### yml_empty
```{r empty}
yml_empty()
```

```{r explain}
Z <- yml() %>%
  yml_output(pdf_document(toc = TRUE, toc_depth = 4)) %>%
  yml_latex_opts(
    fontfamily = "Fira Sans Thin",
    fontsize = "11pt",
    links_as_notes = TRUE
  ) %>%
  # adds suffix `:` to LHS
  yml_toplevel(list(TAGS = "the_tags_line")) %>%
  yml()
Z
is_yml(Z)
```
```{r distill}
distill_listing()

yml_distill_opts(y,
  categories = c("R", "envir", "DT")
)
```

```{r site}
yml_site_opts(y)
```





```{r render, eval=FALSE, include=FALSE 	} 
output_dir <- Sys.getenv("OUTPUT_DIR")
output_dir

{
  file <- ""
  file <- basename(file)
  file <- here("rmd", file)
  file
}

rmarkdown::render(file,
  output_format = "html_document",
  output_dir = "."
)


## replacement for here
if (F) {
  the_dir <- dirname(file)
  file <- file.path(the_dir, file)
}
```
