\footnotesize{footnotesize}   
Remainder of document now ....  But NOT TOC.

#--------------------------------------------------   
		PURPOSE:  data.table:  simple, reference examples 
		SEE example(data.table) (ie in code)
#--------------------------------------------------

<!--
# http://127.0.0.1:22091/library/data.table/doc/datatable-intro.html  
--> 

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = F,  
                      eval = F,
                      include = F,
  					comment="      ##",  
					error=TRUE, 
					collapse=F) 
library(data.table)
library(kableExtra)
``` 
####    References
```
example(data.table)
save(data.table::example(data.table), ascii=T, file="example.txt") Vignettes:  vignette(package="data.table") 
REF: https://stackoverflow.com/questions/32276887/use-of-lapply-sd-in-data-table-r
```

####    Definitions, Terms

```
By Reference | in-place
Deep copy
```

####    List datatable options: 
```{r dt_options}
  op = .Options[grep("datatable", names(.Options))]
  # unlist to named character vector 
  op = unlist(op)
  (dt = data.table(name = names(op), value = op))

```
#### create example dt
```{r begin1, eval=T}
options(datatable.print.class = T)
options(datatable.print.rownames = F)

create_dt  <- function() {
  data.table(x = rep(c("b", "a", "c"), each = 3),
                  y = c(1, 3, 6),
                  v = 1:9)
}
dt  <- create_dt()

dt
```
#### Notation:  columns vs variables
```
"y" or 'y' refers to column in dt
y  refers to a variable
```

#### Notation:  several ways to select columns
```{r notation}
## select cols y, v only.
dt[, c("y", "v")]
dt[, .(y,v)]  # same as dt[, list(y,v)]

w  <- c("y", "v")
dt[, ..w]

dt[, .SD, .SDcols=c("y", "v")]

## but NOT
dt[, .SD, .SDcols= ..w] # ..w not found
```

#### select: one column
```{r one_column}
    cols  <- c("y")
    dt[, .SD, .SDcols = c('y')]
    dt[, .SD, .SDcols = cols]

#options(error=recover)
## NOT (error)
    dt[, .SD, ..cols]
    dt[, .SD, .SDcols = ..cols]
```


#### Select: one col, choose:  character vector or data.table (ie list)
```{r same2}
## character vector
    dt[, y]


## list
    dt[, "y"]
    dt[, 'y']

```


#### setkey 
does NOT remove duplicate rows
keys makes merge efficient
keys(DT) returns key
```
does NOT remove duplicate rows
keys makes merge efficient
keys(DT) returns key
```

```{r key}
## note:  columns are Unquoted
    dt  <- create_dt()
    setkey(dt, y, x)
    dt
##  note:
    attributes(dt) # 'sorted'
```

#### setorder 
```{r setorder}
    dt  <- create_dt()
    setorder(dt,v) 
    dt |> head(3)

    setorder(dt, -v, y)
    dt |> head(3)
```


####  setDT create data.table from a list.
```{r list_exists}
dt1  <- setDT(x = list(1:9))
dt1

# last element
dt1[, .N] # as dt
dt1[.N] # as vector
dt1[1]
dt1[2]

dt1[-2] # omit 2nd elemnt
dt1[!2] # SAME
```

#### DT[i, j, by]
  *  R:                 i                 j        by
  *  SQL:  where | order by   select | update  group by  
  *  subset
  *  order
  *  list .()
  *  return vector vs dt
```{r meaning, include=T, eval=T, echo=T}
dt[x == "a",]
dt[dt$x == "a",]  #SAME
dt[x == "a" & y == 1,]
dt[order(y, -x),]   # y is asc, x in desc
```
### Columns
#### Columns:  Select/drop 
```{r select_col, size='small'}
dt[, list(y,v)]  

# .(y) is alias for list(y)
dt[, list(y)] # as data.table
dt[, .(y)]  # SAME, short cut   
dt[, .(y,v)]

dt[, y]  # as vector!

#  select by column range  (and including)
dt[, x:y]
dt[, .(x,y)]
dt[, .SD, .SDcols=x:y] # SAME

# by column number!
dt[, 1:2]

```
####  Select 1 column, return character vector
```{r }
dt[, x]
```

####  Select 1 column, return as column
```{r }
dt[, "x"]
dt[, c("x")]
dt[, .(x)]  
```


#### Select all, but  1st column
```{r select_a_few}

# drop first column
dt[, -(1)]
```
#### Select all , but  col x
```{r drop_a_few}
dt[, -c("x")]
```


## SAME:
    dt[, c("y", "v")]
    dt[, .(y, v)] # no quotes
    dt[, list(y,v)]

    z  <- c("y", "v")
    dt[, ..z]

    dt[, .SD, .SDcols = c("y", "v")]
    
## Returns vector, unlike data.table above
  dt[, y]   # character vector
 
## WRONG:
    #   dt[, .SD, .SDcols = .(y, v)]   ## Fails, no function for `.`


dt[!2:4]                       # all rows other than 2:4
dt[-(2:4)]                      # SAME

#### Column: calc, summarize, add new col dt[, .(x = FUN())] 
```{r calc}
dt[, sum(v)]  # summarizes, vector
dt[, sum(v), by = x]  # summarizes, as table, grouping

dt[, .(sum = sum(v)), by = x]  # summarizes, with new variable

####select col, but with new name
dt
dt[,.(z = v)]   # not permanent, only prints ONE col


## return ALL columns, rename just one
dt1 <-  dt[,.(x,y,z = v)]
dt1

## return some columns, drop 1, rename 1
dt[,.(y,z = v)][]

##  select cols based patterns=
    dt[, .SD, .SDcols=patterns('^[yv]')]

```
#### create 1 new col, drop another (as temp)
```{r new_remove}
dt  <- create_dt()
dt
dt[, .(new = y + v, x = NULL)]
```

#### create 3 new columns
```{r 3_new}
dt  <- create_dt()
dt
dt[, .(colA=x, colB=y, colC=v)]
dt[, .(x,y,v, colA=x, colB=y, colC=v)]

```


#### Add/Drop New Column (in place) dt[, .(x := FUN())]
```{r wrangle}
# ==========================
#   simple_dt
# ==========================
line  <- NULL
trace(what=quote(simple_dt), browser())
untrace(what=simple_dt)
simple_dt  <- function() {
    line  <- "TAG:    a,b,c,d,e"
    line  <- rep(line, 10)
    dt  <- data.table(line = line)
}
dt  <- simple_dt()
dt

# Add Date
dt[, date := format(Sys.Date())]
dt

# Drop Date
dt[, date := NULL]
dt

# remove TAGS:
pat  <- "^TAG:\\s*"
dt[, no_tag := sub(pat, replacement = "", line)]
dt

# Drop Column line
dt[, line := NULL]
dt
```

#### create 2 new columns, in-place
```{r multiple_new}
dt  <- create_dt()

dt[, c("newA", "newB") := .(x,y)][]
dt[, `:=` (colA =  x, colB = y)][]
dt

## but this is not in-place
dt[, .(colA =  x, colB = y, x, y)][]
dt
```


####  Add/Update/Delete Columns by reference
  *  Ref Semantics: 2A
  *  Read vignette notes!
```{r 2A}
dt = create_dt()

# Form 1, Add new col
dt[, c("new") := y+1 ]
dt

# Update col
dt[, c("new") := new -1][]

## modify only for specific values of x
dt[x=="b", new := new -1][]

# delete "new"
dt[, c("new") := NULL][]

```
####  Columns:  Select col, change col name (temporary)
```{r calc1}
dt  <- create_dt()
dt
# temp change, not in memory
dt[, .(x,y,v, new = y + v)]  # temp
dt

# no permanent change, display only new
dt[, .(new = y + v)]  # temp
dt


# group on col x, print .N, and sum each col
dt[, c(.N, lapply(.SD, sum)), by = .(x)]
#    x N  y  v
# 1: b 3 10  6
# 2: a 3 10 15
# 3: c 3 10 24

# DIFFERENT, EXPLAIN !
dt[, .(.N, lapply(.SD, sum)), by = .(x)]
#    x N V2
# 1: b 3 10
# 2: b 3  6
# 3: a 3 10
# 4: a 3 15
# 5: c 3 10
# 6: c 3 24

# 
# Return grand total of 2 columns, new (temp) name 
dt[, .(new = sum(y,v))]
dt[, .(new = sum(y,v))] #SAME
dt


# BUT: permanently create NEW col
# copy(dt):  not a second reference, actual 2nd copy 
dt1  <- copy(dt)
dt1
identical(dt, dt1)

# change in memory
dt1[, newer := y +v]
dt1
identical(dt, dt1)
rm(dt1)
```

#### Select cols to print
```{r select1}
dt[,c("x","y","v")]

# SAME, easier!
dt[, .(x,y,v)]

# shortcut?
col  <- c("x","y","v")
dt[, ..col]

# FAIL
if (F) {
  dt[, col]
  dt[, .col]
}

dt[, .(col)]
```

### Subsets

#### Ways to subset
```{r logic}
dt
dt[y <5,]

dt[, y < 5]  # vector
dt[, .(y < 5)]

dt[, .((y + v) > 10)]

dt[, .N, by=.(x >0, y> 0)]
#       x    y N
# 1: TRUE TRUE 9

## Seems to act like some key, on v, y; sort of .
dt[, .N, by=.(v >0, y> 0)]
dt[, .N, by=.(v >8, y> 0)]
#        v    y N
# 1: FALSE TRUE 8
# 2:  TRUE TRUE 1

# This:  do not understand!
dt[, c(.N, v, y), by=.(v >8, y> 0)]
```

####  subset, .SD

```{r subset}
dt[, .SD[1]]  # 1st row
dt[, .SD[1:4]]  # 1st 4 rows
```

#### head(.SD,2)
```{r head}
dt

# returns 2 rows (out of all)
dt[, head(.SD, 2), .SDcols= c("x")]

# returns 2 rows (for each group)
dt[, head(.SD, 2), by = .(x)]


# return 2 rows (for each group, but only include colv + col x)
dt[, head(.SD, 2), by = .(x), .SDcols=c("v")]

```

### Working with columns, j

#### aggregrate, .N, by
```{r aggregate}

# 
dt
# for each in Col x, total, return dt
dt[, .(.N), by = .(x)]
dt

# SAME, shortcut
dt[, .N, by = x]
dt

# group by 2 columns, return dt
dt[, .N, by = .(x,y)]
```

#### .N, count: returns integer vector
```{r N}
dt
dt[, .N]
dt[ x == "a", .N]

```

```
DT Reference Semantics
References are pointers to memory.  Pass by reference means NO copy needs to be
made.  Maniupulation being done directly on object in memory.   `:=` specifies
to create new in-place, in ram specifies to create new in-place, in ram
Vignette: https://rdatatable.gitlab.io/data.table/articles/datatable-reference-semantics.html 

Definitions:   
`LHS := RHS`, "modify-in-place",  change object in memory; all refences to
object see the change.


`LHS = RHS`,  no change to object in memory;
```


#### LHS := RHS (general form)
  *  Reference Semantics: 1B
  *  2 forms: (this is not code)
```
FORM 1
DT[, c("colA", "colB", ...) := list(valA, valB, ...)]

 
# when you have only one column to assign to you
# can drop the quotes and list(), for convenience
DT[, colA := valA]


FORM 2 
DT[, `:=`(colA = valA, # valA is assigned to colA
          colB = valB, # valB is assigned to colB
       
)]
```

####  unique, uniqueN: unique values and count 
```{r unique}
    dt  <- create_dt()

dt
##  unique values: return data.table with single col 
    dt[, .(unique(x))]

##	For each x, report number of rows and number of unique y values 
    dt[, .(nrows= .N, nY= uniqueN(y)), by="x"]

    data.table::uniqueN(dt, by = "x")
    uniqueN(dt, by = c("x", "y"))
# [1] 9

##  return dt, with unique ROWs
    unique(dt)

##  return first row b/c unique removes all duplicates
    unique(dt, by=c("x"))

# table, count for each unique rows
    dt[, .N, by = .(x,y)]
    dt[, .N, by = .(x)]
    dt[, .N, by = .(y)]

```

#### uniqueN
```{r uniqueN}
# count
    data.table::uniqueN(dt, by = "x")
    uniqueN(dt, by = c("x", "y"))

    dt  <- create_dt()
    dt[, .(nRows = .N, 
           nX = uniqueN(x),
           nY = uniqueN(y))]


    # remove duplicated rows, base on by=
    unique(dt, by = "x")
    unique(dt, by = c("x", "y"))
```



#### TODO:  all fail, select rows IF y+v > 6
```{r sum}

dt[, .(x,y,v)]
dt[(sum(y,v) > 6), .(x,y,v)]
dt[sum(y,v) > 6, ]
dt[, .SD[sum(y,v) > 6]  ]
```



#### TODO:  charcter to numeric?
FIx
```{r convert}
# x is char
dt[(sum(x,y) > 6)]
dt[with(dt, x + y) > 6, ]

```

####  Study examples
```{r begin2}
create_dt  <- function()
  DT = data.table(x=rep(c("b","a","c"),each=3), y=c(1,3,6), v=1:9)

dt  <- create_dt()
dt

example(data.table)

```

####  Order(x) (sorting )
```{r order}
dt  <- create_dt()

##  order on one column, but still displays ALL columns (optional extra comma)
dt[order(x)]
dt[order(x),]

dt[order(y),]
dt[order(y,v),]
dt[order(y,-v),]
dt[y == 6,][order(v),]

# wrong
dt[y == 6 & order(-v), ]
```

```{r challenge}

# Challenge Problem
# for each ID, count rows with non-zero values, 
  build list of types (with non-zero values)

my_tibble <- tidyr::tibble(
  ID = c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4,5,5,5,5),
  value = c(0,0,0,0,
            1,1,0,1,
            1,1,1,1,
            0,1,0,1,
            1,0,0,0),
  type = rep(c("a", "b", "c", "d"), 5)
)
my_tibble
dt  <- as.data.table(my_tibble) 

##  Each ID, show non-zero count
dt[value != 0, .N, by = ID]


##  Each ID, list the types with non-zero values
dt[value != 0, .(unique(type)), by = ID]

## TODO:
## SAME, but use list-column for the types
dt[value != 0, c(ID, types `:=`  .(type)), by = ID]
dt[]
```

#### Order, base on column
```{r order1}
create_dt()
dt[order(x)]                   # no need for order(DT$x)
dt[order(x),]   # SAME

dt[order(-x)]   # reverse
```

    ## "y" means columns
    ## y   means look for object y
    ##
    ## identical ways:  select columns 

```
#### create list-column
```{r list_column2}
dt[, .(y = .(y), v = .(v)), by="x"]
```

#### simple: reshape, list-column 
```{r simple1}
# practice dt,  use function so easy to re-create
create_dt  <- function() {
  data.table(x = rep(c("b", "a", "c"), each = 3),
                  y = c(1, 3, 6),
                  v = 1:9)
}
dt  <- create_dt()
dt
```

#### Grouping;  head(.SD, 2)
```{r}
# 3lines for each grouping
dt[,.SD, by=x]

# here limit to 2 lines, for each grouping
dt[,head(.SD, 2), by =x]

# first x=b, then x=a ...
dt[,print(.SD), by =x]

# for each grouping, show val
dt[, .(val = y), by =x]

```{r list_column}
# THINK
dt[, .(val = c(y,v)), by=x]

# THINK
dt[, .(val = list(y,v)), by=x]
dt[, .(val = list(c(y,v))), by=x]


```

#### NULL, NA, Explain !
  *  in dt, work with NULL, NA
```{r explain}
dt3  <- function() {
  data.table( ID = c(1,2,3,4,5,6),
      tags= list(
                 "a,b,c",
                 NULL,    # blank
                 NA,      # NA
                 x,       # NA
                 "x",
                 "y,z,NULL,NA")
      )
}
dt  <- dt3()
dt


# Step 1 (remove blank)

f  <- function(e) ifelse(is.null(e), return(NA), return(e))
dt[,.(ID, tags = lapply(tags, f)) ]

dt[is.na(dt$tags)]
dt[is.na(tags)]



# Explainh errors ....
if (F) {
dt[is.null(tags)]
dt[tags == NULL]
dt[tags == NA]
dt[tags == c(NA)]
}


# Explain behavior!
g  <- function(e) return(e)

dt[, .(ID, tags = lapply(tags, g))] 

# 2 ID columns
dt[, .(ID, tags = lapply(tags, g)), by=ID] 

dt[, .(ID, tags = lapply(.SD, g)), by=ID] 

if(F) dt[, tags = lapply(.SD, g), by=ID] 
```

#### remove 1 bad row
```{r lapply_trouble}
    dt  <- data.table(A = c(1,2), 
                    B = c("good", "bad")
                    )
    dt
    str(dt)

    dt[B != "bad"]

```

#### lapply - problems
```{r lapply_problems}

##  NOTICE:   dt.1, dt.2 are DIFFERENCE, why?
##  
    dt.1  <- dt[, .(A, B = lapply(B, function(e) grep(e, pattern="^good", value=TRUE)))] 
str(dt.1)

##  
    dt.2  <- dt[, .(A, 
        B = unlist(lapply(B, function(e) grep(e, pattern="^good", value=TRUE))))] 
str(dt.2)

identical(dt.1, dt.2)


# not what want:
    dt.3  <- dt[, lapply(.SD, function(e) grep(e, pattern="^good", value=TRUE)),
                .SDcols=c("B")] 
dt.3


# 
dt.4  <- dt[, B := lapply(B,  function(e) grep(e, pattern="^good", value=TRUE)),
                c("A", "B")] 
dt.4

```

### By Reference:

#### DEEP COPY (add a ref vs making a copy)  

  *  by reference | no copies make | any modification (in-place) affects both. 
  *  pass by value | make copy | leave original untouched

```{r copy}
# all dt
tables()

##  add new reference (not copy
  new  <- dt
  identical(new,dt)

##  make a copy
  old  <- data.table::copy(dt)
  old
  identical(old, dt)


tables()

## Any modification (by reference, in-place) change objects (in memory) and 
##   anything that points to it

# :=  forces by reference
new  <- dt[, new := 3*v]  

# both changed
new
dt

# no change
old

# now change back
dt  <- copy(old)
dt
```

#### By reference (no copy, in-place) AKA `Reference Semantics` (see Vign) 
  *  SYNTAX:   note!
  *  EXAMPLE:  020 
```{r more_by_reference }
dt  <- create_dt()
dt


##  Multiply columns in-place, same column names, in-place
    cols  <- c('y', 'v')
    dt[, (cols) := lapply(.SD, '*', -1), .SDcols = cols]
    dt


##  But if you do not provide := or col names, it will do it by copy,
    not in-place

    dt  <- create_dt()
    cols  <- c('y', 'v')
    dt[, lapply(.SD, '*', -1), .SDcols = cols]
    dt

##  SAME as first, but create 2 new columns (in-place)
    new_cols  <- c("one", "two")
    dt  <- create_dt()
    dt[, c(new_cols) := lapply(.SD, '*', -1), .SDcols = cols]
    dt

##  FANCY
    dt  <- create_dt()
    cols  <- c("y", "v")
    out_cols  <- paste("log", cols, sep=".")
    out_cols
    dt[, c(out_cols) := lapply(.SD, 
                               function(x) {log(x, base = exp(1))}), 
                              .SDcols = cols]
    dt


## not in-place, no change to original dt
dt  <- create_dt()
dt[, lapply(.SD, '*', -1), .SDcols = cols]
dt
```

#### lapply, Create new column in-place (changes dt in memory_) 
```{r lapply_in_place}
dt  <- create_dt()
dt
## 2nd ref to object in memory; change memory and both change. 
old  <- dt

## add suffix to col v, inplace
suffix = "bye"

dt[, c("v") := lapply(.SD, function(e) paste(e, suffix, sep="."   )), .SDcols = c("v")] 
dt
identical(old,dt)


```
### Data.table tools
```{r dt_tools}

## What's in memory?
  data.table::tables()

## For future, keep a ref, not a copy to DT
  dt.ref  <- dt         # this is another ref
  dt.copy  <- copy(dt)  # this creates 2nd copy
```

####    select rows, based on a variable (quote, eval), dt FAQ 1.6
```{r select}
##  REF  !so 15913832
    dt  <- create_dt()

    col= "b"
    col= "a"
    q = quote(col)
    eval(q)
    dt[x==eval(q)]
```


### create dt
```{r begin}
set.seed(10238)
# A and B are the "id" variables within which the
#   "data" variables C and D vary meaningfully

create_dt  <- function () {
DT = data.table(
  A = rep(1:3, each = 5L), 
  B = rep(1:5, 3L),
  C = sample(15L),
  D = sample(15L)
)
return(DT)
}

dt  <- create_dt()
dt
```

#### Count, sum columns, groupings
```{r sum_columns}
#Sum all columns
dt[ , lapply(.SD, sum)]
#     A  B   C   D
# 1: 30 45 120 120
```

#### Raw Markdown?
```{r testing, results="asis"}
###  Sum all columns
dt[ , lapply(.SD, sum)]
```

#### Count Rows
```{r  resume}
dt[ , lapply(.SD, length)]
dt[, .N, by=A]
dt[, .(.N), by=A]
dt[, .(.N), by=.(A)]
```

#### Count Rows, group by values in 1 column 
```{r count_by}
##  All same
dt[, .N, by=A]
dt[, .(.N), by=A]
dt[, .(.N), by=.(A)]
```

#### By grouping, sum all, except A
```{r except}
dt[ , lapply(.SD, sum), by = A]
#    A  B  C  D
# 1: 1 15 38 43
# 2: 2 15 30 49
# 3: 3 15 52 28
```
#### Sum all columns EXCEPT A
```{r except3}
dt[ , lapply(.SD, sum), .SDcols = !"A"]
```

#### Group on A, count
```{r count2}
dt[ , lapply(.SD, length), by = A]
```

#### Sum all columns EXCEPT A, group on B
```{r complex}
#Sum all columns EXCEPT A, grouping BY B
dt[ , lapply(.SD, sum), by = B, .SDcols = !"A"]
#    B  C  D
# 1: 1 27 29
# 2: 2 17 30
# 3: 3 33 11
# 4: 4 23 36

dt[ , lapply(.SD, length), by = B, .SDcols = !"A"]

# Count, each combination of c("A", "B")
dt[ , lapply(.SD, length), by = c("A","B")]
```

#### remove 1 col (modify in place)
```{r remove_col}
## start fressh
dt  <- create_dt()
dt
str(dt)


## in-place (in memory) remove col D
    colD  <- c("D")
    dt[, (colD) := NULL]
    dt
```
#### remove >1 cols (in-place)
```{r 2_col}
## in-place, cols B and C
    cols  <- c("B", "C")
    dt[, (cols) := NULL][]
```


### JOINS (merge)
First,  datatable::merge, based on, but not same, base::

#### Merge:
```{r inner}
dt  <- create_dt()
dt2  <- data.table(Z=c("a"))

## inner join
merge(dt,dt2, by.x = "x", by.y = "Z")

merge(dt,dt2, by.x = "x", by.y = "Z", allow.cartesian=T)


merge(dt,dt2, by.x = "x", by.y = "Z", all=TRUE)


## left join
merge(dt,dt2, by.x = "x", by.y = "Z", all.x=T)


## right join
merge(dt,dt2, by.x = "x", by.y = "Z", all.y=T)

## all
merge(dt,dt2, by.x = "x", by.y = "Z", all=T)


## cross


## natural
merge(dt, dt2)

## key
```{r less_confusion}
setkey(dt, x, verbose=T)
setkey(dt2, Z)

## check
key(dt)
key(dt2)

tables()

# FAIL	
merge(dt, dt2, by=)

```


#### More Join Detail:   $HOME/code/try_things_here/DT/4000_DT_joins.R

#### Simple Left, Inner 
```{r inner}
##  create table and its lookup
    dt  <- create_dt()
##  new dt, with limited values of x 
    dt1  <- data.table(x=c("a", "d"))
    dt1

##  (Left) All rows of dt1, match or not in dt
    dt[dt1, on="x"]

##  (Inner) Only rows of dt1  with match in dt 
    dt[dt1, on="x", nomatch=0]

```

#### Lookup Table
```{r lookup}


lookup  <- data.table(abbrev = c("ny", "or", "tx", "ca"),
                      name = c("new york", "oregon", "texas", "california")
                      )

#  --------

#  fix !
l = list(list("ny", "blue"), list("ca", "blue"))
main  <- data.table(l)

#  --------

main  <- data.table(abbrev = c("ny", "ca", "ma"),
                    party = c("blue", "blue", "blue"))

main
#    abbrev party
# 1:     ny  blue
# 2:     ca  blue

lookup
#    abbrev       name
# 1:     ny   new york
# 2:     or     oregon
# 3:     tx      texas
# 4:     ca california


#inner join, match in both
lookup[main, on="abbrev", nomatch = 0]

# all rows in main
lookup[main, on="abbrev"]



```



#### rbindlist
```{r rbindlist}

get_filenames  <- function(e) rep(names(e), 2)




# =====================
#   rbindlist()
# =====================

#  TWO dt
dt1  <- data.table(name=c("file1", "file1"),
                   tags=c("tag1", "tag2"))
dt1
dt2  <- data.table(name=c("file2", "file2", "file2"),
                   tags=c("tag3", "tag4", "tag5"))
dt2


rbindlist(list(dt1, dt2))
# =========================
# TWO lists 
l1  <-  list(name=c("file1", "file1"), tags=c("tag1", "tag2"))
l1
l2  <- list(name=c("file2", "file2", "file2"), tags=c("tag3", "tag4", "tag5"))
l2

rbindlist(list(l1,l2))
# ======================

```

#### normalize - votes
Status?   not sure.
```{r votes}
# each=2, each element repeated twice, before moving on.
votes <- function() {
  data.table(issue = rep(c("issue1", "issue2"),times=1, each=2),
             A = rep(c("Y",NA), 2),
             B = rep(c(NA,"N"),2)
             )
}
dt = votes()
dt
```


#### Use melt (wide --> long)
reprex for slack
```{r melt}
reprex(wd =  ".", venue = "slack",{


#' For those of us tidyverse-averse (or just stubborn) here is one way twith data.table :

library(data.table)
#' Create fake data  in wide format ('unnormalized')
#' Column A refers to votes of Country A, etc

votes <- function() {
  data.table(issue = rep(c("issue1", "issue2"),times=1, each=2),
             A = rep(c("Y",NA), 2),
             B = rep(c(NA,"N"),2)
             )
}
dt = votes()
dt


#' melt (wide to long, aka 'normalize' the data)  
#'  *  id.vars are the variables want to keep.  
#'  *  measure.vars are the columns to collapse.  
dt.1  <- data.table::melt(dt,
                 id.vars = c("issue"),
                 measure.vars = c("A", "B"),
                 variable.name = "country",
                 value.name = "vote",
                 na.rm = TRUE

)
dt.1
###' order by issue
dt.2  <- dt.1[order(issue)]
dt.2
             })
```


```{r}
dt  <- votes()
dt_A  <- dt[!is.na(A), .(country = "A", vote=A), by = issue]
dt_A
dt_B  <- dt[!is.na(B), .(country = "B", vote=B), by = issue]
dt_B

```
```{r render, eval=FALSE, include=F	} 
args(rmarkdown::render)
file  <- "/home/jim/code/try_things_here/DT/00291_DT_STUDY_examples.Rmd"
file  <- normalizePath(file)
file
rmarkdown::render(input= file,
                  output_format="pdf_document",
                  #output_file=here("rmd","out")
                  output_dir="~/Downloads/print_and_delete/"
                  )

#   *  if .gitignore balks,  then add !out/**   to close .github

```
