---
title: "090_data_table.R"
output:  
  pdf_document:
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
TAGS:  data.table,vignette,flights14,dplyr,magrittr,purrr
---
~/code/MASTER_INDEX.md
./README.md


### Vignette for data.table (subset ONLY)
  *  REF: [http://rdatatable.github.io](rdatatable)
  *  Best to follow vignette AND its USEFUL comments.
  *  Code here is SUBSET



```{r setup}
library(magrittr)
library(data.table)

input <- if (file.exists("./data/flights14.csv")) {
  "data/flights14.csv"
} else {
  "https://raw.githubusercontent.com/Rdatatable/data.table/master/vignettes/flights14.csv"

  # saveRDS("flights14.csv", "data/flights14.csv")
}
# 2.2 MM
flights <- fread(input)
head(flights)
DT <- flights
```

### Guess:  how many columns in result?
```{r guess}
## all columns returned; use order, not sort.
head(DT[order(origin, -dest), ])

## 2 columns, only ; use j
head(DT[, .(arr_delay, dep_delay)])

## 1 column
head(DT[, .(arr_delay)])

## atomic vector, not list
head(DT[, arr_delay])
```

### Rename, note dep_delay is not renamed; return 2 col
```{r rename}
head(DT[, .(x = arr_delay, dep_delay)])
```

### Do a calc in j (mean), return 1 col; named V1
```{r calc_j}
head(DT[, .(mean(arr_delay))])

## here we give it name.
head(DT[, .(mean = mean(arr_delay))])
```


#### group by; implict columns
```{r by}
## 2 columns
head(DT[, .N, by = .(origin)])

## 3 columns
head(DT[, .N, by = .(origin, dest)])


## 4 columns, note that j results in V1
head(DT[, .(mean(arr_delay)), by = .(origin, dest, month)])
# rename V1 to y
head(DT[, .(y = mean(arr_delay)), by = .(origin, dest, month)])
```

#### keyby
```{r keyby}
##  To fix, replace by with keyby (compare to 4 above); 4 cols
head(DT[, .(y = mean(arr_delay)), keyby = .(origin, dest, month)])


## gotcha, keyby doesn't like -dest??
## instead:  chain:   3 cols
head(DT[, .N, by = .(origin, dest)])[order(origin, -dest)]
```
####  fancier j, expressions; selecting columns
  *  DT[, .N, by=.(a,b, c> 1)]  # only c is condition, a,b, .N and c all
    selected
```{r}
## group, by 3 columns
head(DT[, .N, .(dep_delay > 0, arr_delay > 0)])
head(DT[, .N, by = .(dep_delay > 0, arr_delay > 0)])

## same, but new name
head(DT[, .N, by = .(x = dep_delay > 0, arr_delay > 0)])
```

#### .SDcols to choose output columns
```{r outpute}
## omits by col. `by` [all columns]
DT[, print(.SD), by = .(carrier)]

## no, not what I want [select only carrier columns]
DT[, print(.SD), by = .(carrier), .SDcols = c("carrier")] %>% head()

## USE .SDcols to limit output cols, NOTE: use quotes   c( "...", "...")
## gotcha!
```

#### need to take mean of 100 cols
  *  dt[, lapply(.SD, mean), by = ID] # returns all columns + ID
```{r lapply}
##  group by=
##  but only do calc for .SDcols, and output;  not all columns in .SD
DT[, lapply(.SD, mean), by = .(origin, dest, month), .SDcols = c(
  "dep_delay",
  "arr_delay"
)]

##
## START HERE
## note I added carrier!
DT[, print(.SD), by = .(carrier), .SDcols = c(
  "dep_delay", "arr_delay",
  "carrier"
)]
DT[, head(.SD), by = .(carrier), .SDcols = c(
  "dep_delay", "arr_delay",
  "carrier"
)]

# Too much output, even with head, limit to 2 rows Per Group.
DT[, head(.SD, 2), by = .(carrier), .SDcols = c("dep_delay", "arr_delay", "carrier")]
```


```{r render, eval=F, include=F}
file <- "00290_DT_vignette_intro.rmd"
file <- here("rmd", file)
file


rmarkdown::render(
  input = file,
  # output_format = "pdf_document",
  # output_format = "html_document",
  output_format = c(
    "html_document", # if want both
    "md_document"
  ),
  output_dir = "md",
  output_file = "out"
)
```
