---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.5in,top=0.25in   
TAGS: data.table,flights14
---


<!--

-----------------
SPECIAL NOTE:  
-----------------

LHS := RHS
Vignette REF Semantics is important!
But uses flights14

SEE 0291 for basic examples of LHS := RHS
(probably delete this file)
-->



```{r setup, include=FALSE		}
#knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read
library(data.table)
```



### Trival example modify-in-place, :=
```{r trival}
X  <- data.table(a = c(1,2,3))
X

X[, .(c = 3*a)]
X[, .(a, c=3*a)]

# no c here
X


# permanently alters X
X[, b := .(2 * a)]
X

```
### Get flights14.csv
```{r begin}

if (file.exists(here("data", "flights14.csv"))) {
   input  <- here("data","flights14.csv")
   flights  <- fread(input)
} else {
  input  <- "https://raw.githubusercontent.com/Rdatatable/data.table/master/vignettes/flights14.csv"
  flights  <- fread(input)
  fwrite(flights,file = here("data","flights14.csv"))

}

flights


```


## Section 1 - Reference Symantecs
(skip for now)


## Section 2 - Add/Delete/Modify - by reference
  *  DT has problem with continuation on 2 lines?
  *  `:=`  operator form solves this
```{r}
# add 2 fields, modify-in-place
flights[, `:=` (speed = distance / (air_time/60), # speed in mph (mi/h)
               delay = arr_delay + dep_delay)]   # delay in minutes
# check: 2 new fields
str(flights)
```



### Section 2b, modify-in-place
```{r section_2b}
# example hours:
# get all 'hours' in flights, note both 0 and 24
flights[, sort(unique(hour))]

# remove 24L in underlying table?
flights[ hour==24L, hour := 0]

# has underlying table changed? yes
flights[, sort(unique(hour))]

# this is not working!
flights[, .(unique(hour))][order(v1)]
flights[, .(order(unique(hour)))]
```

## Section 2d - group
  *  Note:  if tie, DT recycles to include all ties
```{r 2d}
flights[, max_speed := max(speed), by = .(origin, dest)]
# to view: 
flights[]
# seems easier, to me.
head(flights)

flights[, max_speed := max(speed), by = .(origin, dest)]
flights[max_speed > 620 & air_time < 100,] 

[, c(carrier, origin,
                                                             dest, max_speed) ]
```

### Section 2e - multiple columns, :=
  *  repeat section 2d, but also group by month
  *  Notes:  c(out_cols) or just (out_cols) tell DT that not creating a new
    column, but just updating existing multi-columns (in-place)
  *  .SDCols = tells DT which columns to work on


```{r section_2e}
in_cols  = c("dep_delay", "arr_delay")
out_cols = c("max_dep_delay", "max_arr_delay")


flights[, c(out_cols) := lapply(.SD, max), 
        by = month, .SDcols = in_cols]
flights[]
flights[, .(month,day,carrier,max_dep_delay,max_arr_delay)]
head(flights, n=10L)
```

## Section 3 := and copy

### functions 
  *  using := wil  change every reference to DT, even in different
    environment(check)
  *  using = does not
```{r}


foo <- function(DT) {
  DT[, speed := distance / (air_time/60)]   # changes all refernces to DT
  DT[, .(max_speed = max(speed)), by = month] # this env only, return copy
  only.
}

ans = foo(flights)
tables()  # now two
identical(flights, ans)

str(flights)
str(ans)

```


```{r knit_exit(), include = FALSE  , eval=FALSE } 
# knitr::knit_exit()
```


```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# in general, pdf will look nicer

file  <-  "/home/jim/code/try_things_here/rmd/0088_base_package_commands.Rmd"

rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
