	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
title: "title"
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  data.table,example,vignette,normalize
---


TODO:   Clean up; merge into 0291
PURPOSE:  simple, reference examples :
    * nest, unnest, normalize, list-column etc.

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(data.table)
```

### create dt (note:  ID)
```{r begin}
# practice dt,  use function so easy to re-create
create_dt  <- function() {
  data.table(ID = rep(c("b", "a", "c"), each = 3),
                  x = c(1, 3, 6),
                  y = 1:9)
}
dt  <- create_dt()
dt
```

#### normalize DT by concatenating TWO columns into ONE new
```{r normalize}
dt[,.SD, by=ID]

# `concatinate two chr[] for each ID` (into new variable)
dt[, .(c(x,y)), by = ID]
dt
```

#### compact:  into list-column
```{r}
dt  <- create_dt()
dt[,.(list(x,y))      , by=ID]
```

### group and head()
```{r}
# 3lines for each grouping
dt[,.SD, by=ID]


# for each ID, choose max y
dt[, .SD[which.max(y)], by = ID]

# here limit to 2 lines (not 3), for each grouping
dt[,head(.SD, 2), by =ID]

# break out by ID
dt[,print(.SD), by =ID]

# for each grouping, show val
dt[, .(val = y), by =ID]

```{r list_column}
# THINK
dt[, .(val = c(x,y)), by=ID]

# THINK
dt[, .(val = list(x,y)), by=ID]
dt[, .(val = list(c(x,y))), by=ID]
```


### Group on ID, display list-column of data.tables!
  *   make list of each SD,  then list all the results
```{r data.table}

dt  <- create_dt()
dt

# 3 the SAME
dt[, list(dt.xy=list(.SD)), by=ID]
dt[, .(dt.xy=list(.SD)), by=ID]
dt[, .(dt.xy=.(.SD)), by=ID]


# FAILS b/c do not have every subset
if (F) dt[, .(dt.xy=.SD), by=ID]
```

### undo/unwind list-column
```{r un}

dt2  <- dt[, .(dt.xy=.(.SD)), by=ID]
dt2

# each ID, pop off 1st entry (there is only 1)
dt3  <- dt2[, dt.xy[[1]], by = ID]
identical(dt,dt3)
```




```{r select}
dt[,c("ID","x","y")]

# SAME, easier!
dt[, .(ID,x,y)]

# shortcut?
col  <- c("ID","x","y")
dt[, ..col]

# FAIL
if (F) {
  dt[, col]
  dt[, .col]
}

dt[, .(col)]
```


### normalize example3
  * ONE column contain SINGLE character vector: c("a, b, c")
  * UNSPLIT and normalize so each element put on separate ROW (normaliz)
  * Study NA, NULL
  * Here we will several situations to solve
  * Unpack ONE column of muliptle strings; PUT each string inseparate ROW
    (normalize)
```{r normalize2}
#
dt3  <- function() {
  data.table( ID = c(1,2,3,4,5,6),
      tags= list(
                 "a,b,c",
                 NULL,    # blank
                 NA,      # NA
                 x,       # NA
                 "x",
                 "y,z,NULL,NA")
      )
}

dt  <- dt3()
dt


# Step 1 (cleanup:: remove blank)
f  <- function(e) ifelse(is.null(e), return(NA), return(e))
dt  <- dt[,.(ID, tags = lapply(tags, f)) ]
dt
dt[is.na(dt$tags)]
dt[is.na(tags)]


# Step 2  more cleanup: use NA_character_ version of NA
#base::strsplit() works with NA_character_, but not NA
f  <- function(e) ifelse(is.character(e),return(e), return(c(NA_character_)))
dt  <- dt[, .(ID, tags = lapply(tags, f)) ]
dt


# Step 3, split and normalized (must cleanup step 1,2 or will bauk)
split_up  <- function(e) base::strsplit(e, split="[,]")

# remove unneeded list
unlist(split_up("a,b,c"))
}

#### do it ... normalize
{
dt[, .(tags = unlist(lapply(tags, split_up) )), by=ID]

dt
    }

#### Just DETAILS below, of normalization.

{
  # test 
# list, with chr[]
split_up(c("a,b,c"))
# same
split_up("a,b,c")

split_up(c(NA, "b"))
split_up(c(NA, NA, NA, "b"))
split_up(NA_character_)
split_up(c(NA_character_))


# error
if (F){
split_up(NA)
split_up(c(NA))
split_up(c(NA, NA))
}



# Line by Line, check
# normalized! unlist
dt[ID == "1",.(ID, tags = unlist(lapply(tags, split_up) ))]

#  compare
dt[ID == "1",.(ID, tags = lapply(tags, split_up) )]




dt[ID == "2",.(ID, tags = unlist(lapply(tags, split_up) ))]
dt[ID == "3",.(ID, tags = unlist(lapply(tags, split_up) ))]
dt[ID == "4",.(ID, tags = unlist(lapply(tags, split_up) ))]
dt[ID == "5",.(ID, tags = unlist(lapply(tags, split_up) ))]
dt[ID == "6",.(ID, tags = unlist(lapply(tags, split_up) ))]

dt[, .(tags = unlist(lapply(tags, split_up) )), by=ID]

dt
```

```{r knit_exit()} 
knitr::knit_exit() 
```

\newpage

```{r render, eval=FALSE	} 
args(rmarkdown::render)
file  <- "/home/jim/code/try_things_here/rmd/00298_DT_normalize.Rmd"
file  <- basename(file)
file  <- here("rmd", file)



rmarkdown::render(input = file,
                  #output_format = "pdf_document",
                  #output_format = "html_document",
                  output_format=c("html_document",    # if want both
                                  "md_document"),
                  output_dir = "md",
                  output_file = "out")

```
