---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
        mathjax: default
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
---


###	PURPOSE:   Collect DT merge/join examples
REF: https://stackoverflow.com/questions/54864132/intuition-for-data-table-join-syntax

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = T   ) # T easier to read
library(data.table)
```

###	STOP, understand these before proceeding
```{r}
##	Which columns?
##  Then which rows?

DT  <-  data.table(r=rep(c("b","a","c"),each=3), s=c(1,3,6), t=1:9)
DT
#    r s t
# 1: b 1 1
# 2: b 3 2
# 3: b 6 3
# 4: a 1 4
# 5: a 3 5
# 6: a 6 6
# 7: c 1 7
# 8: c 3 8
# 9: c 6 9

X  <-  data.table(x=c("c","b"), y=8:7, z=c(4,2))
#    x y z
# 1: c 8 4
# 2: b 7 2


if (F) DT[X]   ## error

#	all elments of X; all columns of X except col match DT
DT[X, on=.(r=x)]
#    r s t y z
# 1: c 1 7 8 4
# 2: c 3 8 8 4
# 3: c 6 9 8 4
# 4: b 1 1 7 2
# 5: b 3 2 7 2
# 6: b 6 3 7 2


#	To prove this is NOT an inner join try
	Y  <- data.table(x=c("c","z"), y=8:7, z=c(4,2))
	DT[Y, on = .(r = x)]


# insist on column for x
	DT[X, on=.(r=x), .(r,s, t, x, y, z)]
#    r s t x y z
# 1: c 1 7 c 8 4
# 2: c 3 8 c 8 4
# 3: c 6 9 c 8 4
# 4: b 1 1 b 7 2
# 5: b 3 2 b 7 2
# 6: b 6 3 b 7 2

DT[X, on =.(r=x,  s < y)]
#    r s t z
# 1: c 8 7 4
# 2: c 8 8 4
# 3: c 8 9 4
# 4: b 7 1 2
# 5: b 7 2 2
# 6: b 7 3 2

DT[X, on =.(r=x,  s > y)]
#    r s  t z
# 1: c 8 NA 4
# 2: b 7 NA 2

DT  <- data.table(r = c(1), s=NA, t=NA)
#    r  s  t
# 1: 1 NA NA

X  <- data.table(x=c(1,2),y = c(NA, NA), z=c(NA,NA))
#    x  y  z
# 1: 1 NA NA
# 2: 1 NA NA

DT[X, on =.(r=x)]
#    r  s  t  y  z
# 1: 1 NA NA NA NA
# 2: 1 NA NA NA NA

DT[X, on =.(r==x, s > y), nomatch=NULL]
#    r  s  t  z
# 1: 1 NA NA NA
# 2: 1 NA NA NA

```	
When you have a X[Y] join it means: "For every value in Y try to join a value
from X", hence, basically this is a left join to Y and the result will be the
length of Y (I agree it's kind of counter-intuitive). – David Arenburg Feb 25 '19 at 11:17

####	Introductory Example
  *  DT[X, on= .(x, y <= foo)]
	*  `key`  means whatever we are matching on, whether or not table is actually
		keyed.
	*  Go through each row of X, find all rows matche in DT. st. c(c,b)
	*  on= means
	*  (always read left to right) DT$x == X$x and
	*  DT$y  <= X$foo
```{r simple_join1}

DT  <-  data.table(x=rep(c("b","a","c"),each=3), y=c(1,3,6), z=1:9)
#    x y z
# 1: b 1 1
# 2: b 3 2
# 3: b 6 3
# 4: a 1 4
# 5: a 3 5
# 6: a 6 6
# 7: c 1 7
# 8: c 3 8
# 9: c 6 9

X  <-  data.table(a=c("c","b"), b=8:7, c=c(4,2))
#    a b c
# 1: c 8 4
# 2: b 7 2

# Work through this:  yields understanding !
DT[X, .(x, y, z, DT.a = a, DT.b = b, DT.c = c), on= .(x=a, y <= c)]
#    x y z DT.a DT.b DT.c
# 1: c 4 7    c    8    4
# 2: c 4 8    c    8    4
# 3: b 2 1    b    7    2

DT[X]
DT[X, on=.(x=a)]

```

```
TIP:	  do not think inner, outer ... or SQL ... 

DT[X
means go through X row by row, select those only rows of DT which match  (aka WHERE)

.(x, y, z, DT.a = a, DT.b = b, DT.c = c) 
refers to columns in find result.

on= .(x=a, y <= c)   
criteria for match:  DT.x == X.a && DT.y == X.c (always left to right)



```{r simplest}
A  <- data.table(x=c(1,2,3,4))
B  <- data.table(y=c(2,3,4,5))

setkey(A,x)
setkey(B,y)
## right outer 
A[B]   
#    x
# 1: 2
# 2: 3
# 3: 4
# 4: 5

## inner
A[B, nomatch=0]
#    x
# 1: 2
# 2: 3
# 3: 4

## left
B[A]
#    y
# 1: 1
# 2: 2
# 3: 3
# 4: 4

```
#### simple DT join

  *  DT[X] is a right-join (odd, yes) on X.
  *  REF:  https://gist.github.com/nacnudus/ef3b22b79164bbf9c0ebafbf558f22a0
  *  REF:  https://stackoverflow.com/questions/67261777/data-table-join-is-hard-to-understand



 ```{r  simple_join}

DT  <-  data.table(x=rep(c("b","a","c"),each=3), y=c(1,3,6), v=1:9)
X  <-  data.table(x=c("c","b"), v=8:7, foo=c(4,2))

#    x y v
# 1: b 1 1
# 2: b 3 2
# 3: b 6 3
# 4: a 1 4
# 5: a 3 5
# 6: a 6 6
# 7: c 1 7
# 8: c 3 8
# 9: c 6 9

#    x v foo
# 1: c 8   4
# 2: b 7   2

DT[X, on= .(x, y <= foo)]
#    x y v i.v
# 1: c 4 7   8
# 2: c 4 8   8
# 3: b 2 1   7


## error, join on what??
   if (F) DT[X]

##  DEFAULT is right outer join
    DT[X, on = "x"] 
    #    x y v i.v foo
    # 1: c 1 7   8   4
    # 2: c 3 8   8   4
    # 3: c 6 9   8   4
    # 4: b 1 1   7   2
    # 5: b 3 2   7   2
    # 6: b 6 3   7   2


##
# =============================
    X  <- data.table(x = "c")

    DT[X, on=.(x)]
#    x y v
    #
vignette("datatable-secondary-indices-and-auto-indexing")
          vignette("datatable-keys-fast-subset")
vignette("Joins and rolling joins")
# 1: c 1 7
# 2: c 3 8
# 3: c 6 9

# =============================
##
    X  <- data.table(x = "c", y=c(1,3))
    DT[X, on=.(x)]
#    x y v i.y
# 1: c 1 7   1
# 2: c 3 8   1
# 3: c 6 9   1
# 4: c 1 7   3
# 5: c 3 8   3
# 6: c 6 9   3

# =============================
##
    DT[X, on=.(x,y)]
#    x y v
# 1: c 1 7
# 2: c 3 8

# =============================
##
    X  <- data.table(x = "c", y=c(1,3), foo=c(4,2)) 
#    x y foo
# 1: c 1   4
# 2: c 3   2

##
    DT[X, on=.(x)]
#    x y v i.y foo
# 1: c 1 7   1   4
# 2: c 3 8   1   4
# 3: c 6 9   1   4
# 4: c 1 7   3   2
# 5: c 3 8   3   2
# 6: c 6 9   3   2
    #
# =============================
    #
   DT[X, on=.(x,y)] #
#    x y v foo
# 1: c 1 7   4
# 2: c 3 8   2

# =============================
    

##
    DT[X, on = .(x, y <=foo)]
#    x y v i.v
# 1: c 4 7   8
# 2: c 4 8   8
# 3: b 2 1   7


```

```{r join}

a <- data.frame(
  one = 1:5,
  two_a = 10:14
)

b <- data.frame(
  one = 3:5,
  two_b = c("a", "b", NA)
)

left_join(a, b) %>%
  # but ONLY calculate this for one %in% b
  mutate(three = two_a + 2)

Desired output:

c  <- data.frame(
  one = 1:5,
  two_a = 10:14,
  two_b = c(NA, NA, "a", "b", NA),
  # this column "knows" the last three rows have data frame b info
  three = c(NA, NA, 14, 15, 16)
)
a; b; c

```

```{r so}
library(data.table)
A <- data.table(id = letters[1:10], amount = rnorm(10)^2)
B2 <- data.table(
  id = c("c", "d", "e", "e"), 
  ord = 1:4, 
  comment = c("big", "slow", "nice", "nooice")
)
A
#     id amount
#  1:  a 0.0667
#  2:  b 0.0834
#  3:  c 0.8062
#  4:  d 2.1857
#  5:  e 1.2807
#  6:  f 0.3719
#  7:  g 0.0872
#  8:  h 0.5281
#  9:  i 0.8966
# 10:  j 0.9682
#
B2
#    id ord comment
# 1:  c   1     big
# 2:  d   2    slow
# 3:  e   3    nice
# 4:  e   4  nooice

# preferred way

A[, comments := B2[A, on=.(id), x.comment]]
```

### [REF](https://stackoverflow.com/questions/21435339/data-table-vs-dplyr-can-one-do-something-well-the-other-cant-or-does-poorly/27718317#27718317)
```{r DT}

library(data.table)
example(data.table)

dt.a  <- data.table(a)
dt.b  <- data.table(b)


# best to key 1st
# setkey(DT1, x, y)

 # 1. normal join
dt.a[dt.b, ,on="one"]  ## data.table syntax
#left_join(dt.b, dt.a) ## dplyr syntax


 # 2. select columns while join    
 DT1[DT2, .(z, i.mul)]
# left_join(select(DT2, x, y, mul), select(DT1, x, y, z))

 # 3. aggregate while join
 DT1[DT2, .(sum(z) * i.mul), by = .EACHI]
 DF1 %>% group_by(x, y) %>% summarise(z = sum(z)) %>% 
     inner_join(DF2) %>% mutate(z = z*mul) %>% select(-mul)

 # 4. update while join
 DT1[DT2, z := cumsum(z) * i.mul, by = .EACHI]
 ??

 # 5. rolling join
 DT1[DT2, roll = -Inf]
 ??
  
```


## Suppose (lost)
```{r  more_tries}
dt1  <- data.table(A = c(1,2,3,2), 
                   B = c(4, 5,6, 20),
                   C = c(7, 8, 9, 21)
                   )

dt2  <- data.table(E = c(1,3,2),
                   F = c(10,12,11),
                   G = c(13, 15, 14),
                   H = c(16, 18, 17)
                   )
dt1
#    A  B  C
# 1: 1  4  7
# 2: 2  5  8
# 3: 3  6  9
# 4: 2 20 21
#

dt2
#    E  F  G  H
# 1: 1 10 13 16
# 2: 3 12 15 18
# 3: 2 11 14 17


# how to join/merge to get this?
dt3
A   B   C   H
1   4   7   16
2   5   8   17
3   6   9   18
2   20  21  17      


dt1[dt2, on= c(A = "E")]
dt1[dt2, on= c("A" = "E")]

# Why fails?
if (F)
  dt1[dt2, on= c(A = E)]

```

---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 4
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  4   
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
TAGS:  DT,joins 
---
~/code/MASTER_INDEX.md


### REF:  https://gist.github.com/nacnudus/ef3b22b79164bbf9c0ebafbf558f22a0
library(data.table)

?`[.data.table`
?data.table::":="

DT <- data.table(x=rep(c("b","a","c"),each=3), y=c(1,3,6), v=1:9)
#    x y v
# 1: b 1 1
# 2: b 3 2
# 3: b 6 3
# 4: a 1 4
# 5: a 3 5
# 6: a 6 6
# 7: c 1 7
# 8: c 3 8
# 9: c 6 9
#
X  <- data.table(x=c("c","b"), v=8:7, foo=c(4,2))
#    x v foo
# 1: c 8   4
# 2: b 7   2

colnames(DT)
# [1] "x" "y" "v"

colnames(X)
# [1] "x"   "v"   "foo"

DT[X, on="x"]            # * right join (DT is left, X is right)
                         # * SELECT DT RIGHT JOIN X ON DT$x = x$x
                         # * X$v is renamed i.v to avoid conflict with DT$v.
                         #     The i comes from the fact that X is given as the
                         #     'i' argument

X[DT, on="x"]            # left join  (DT is left, X is right)
                         # SELECT X RIGHT JOIN DT ON X$x = DT$x

DT[X, on="x", nomatch=0] # inner join
                         # SELECT DT INNER JOIN X ON DT$x = X$x

DT[!X, on="x"]           # not join   (DT is left)
                         # SELECT DT LEFT JOIN X ON DT$x != X$x

DT[X,   on=.(y<=foo)]      # NEW: non-equi join (v1.9.8+)
                         # SELECT DT RIGHT JOIN X ON DT$y <= X$foo
# * i.x and i.v are new names for the X$x and X$v, which is given as the i
#     argument, that conflict with column names in DT.  You can prove this to
#     yourself by observing that there is no row in DT where x=b and v=7.
# * X$foo isn't returned at all, which is annoying, but is consistent with the
# right-hand column being dropped in merges, where it becomes redundant in the
# output.  You can select it manually (along with any other column) like so:
DT[X, .(x, y, v, i.x, i.v, i.foo), on=.(x, y>=foo)]  # Select columns to return, including ones that would otherwise be dropped

DT[X, on="y<=foo"]                          # same as above
DT[X, on=c("y<=foo")]                       # same as above

DT[X, on=.(y>=foo)]                         # NEW non-equi join (v1.9.8+)
DT[X, on=.(y>=foo)]                         # NEW non-equi join (v1.9.8+)

DT[X, .(x,y,x.y,v,i.foo), on=.(x, y>=foo)]  # Select x's join columns as well

DT[X, on=.(x, y<=foo)]                      # NEW non-equi join (v1.9.8+)
# SELECT DT RIGHT JOIN X ON DT$x = X$x AND DT$y <= x$foo

DT[X, on=.(x, v>=v), sum(y)*foo, by=.EACHI] # NEW non-equi join with by=.EACHI (v1.9.8+)
# Is this doing GROUP BY DT$x, DT$v?
#
#
#
### Even simpler: (write SQL)
dt  <- data.table( id = c(1,2,3), a = c("a1", "a2", "a3"))
dt
#    id  a
# 1:  1 a1
# 2:  2 a2
# 3:  3 a3

X  <- data.table( id = c(1,2,4), b = c("b1", "b2", "b3"))
X
#    id  b
# 1:  1 b1
# 2:  2 b2
# 3:  4 b3



## inner join, equi-join
   dt[X, on = .(id), nomatch=0]
#    id  a  b
# 1:  1 a1 b1
# 2:  2 a2 b2

## select X right outer join DT on (all X rows)
   dt[X, on = .(id) ]
   dt[X, .(id, b), on = .(id)]


## select dt right outerjoin X (all dt rows)
   X[dt, on = .(id) ]
   X[dt, .(id, a),on = .(id) ]
