  NOTE:  Example taken from updating R packages, where .libPaths()[[2]] is read-only
-- /home/jim/code/try_things_here/DT/570_DT_update_R_TOY_EXAMPLE.qmd
<!-- TODO
- quarto, double column for simple vertical lists, tibbles, etc/

-->

##    Toy Example


###  Create toy example

<!-- {{{ -->


When I update my R packgaes, I have a problem:  my site directory is read-only.  
In a separate post, I will show how I navigate this without opening the
permissions. 

In this post, I use a toy example to illustrate the use of joins to accomplish this.

The toy example simplifies results of  `installed.packages()` and `old.packages()`

- **inst**;  currently installed packages, using installed.packages
- **cand**;  packages R thinks are old, using old.packages
- **x**;     represents pkg
- **y**;     represents directory (1 or 2)      (.libPaths() has only 2 entries)

<!-- }}} -->

<!--
###  Prepare toy example of what `installed.packages` reports:
-->
<!-- {{{ -->
```{r}
#| include: false
#| eval: true

library(data.table)
(inst = data.table(x=(LETTERS[1:4]),
       y= c(rep(1,4))
       ))
#   add 4 packages in 2
(inst = rbind(inst, data.table(x = LETTERS[5:8],y= c(rep(2,4))) ))
(inst= rbind(inst, data.table(x=c("E","H"), y=c(1,1))))
setorder(inst)
dput(inst)
#     x y
#  1: A 1
#  2: B 1
#  3: C 1
#  4: D 1
#  5: E 2
#  6: F 2
#  7: G 2
#  8: H 2
#  9: E 1
# 10: H 1
```
<!-- }}} -->


<!--
### Prepare Candidates  or old.packages for toy examples
-->
<!-- {{{ -->

```{r }
#| include: false
#| eval: true
s = "x, y
A, 1
B, 1
E, 1 
F, 2
E, 2
H, 2"
cand = data.table::fread(s)
dput(cand)
# structure(list(x = c("A", "B", "E", "F", "E", "H"), y = c(1L, 
# 1L, 1L, 2L, 2L, 2L)), row.names = c(NA, -6L), class = c("data.table", 
# "data.frame"), .internal.selfref = <pointer: 0x55578dd18740>)
#    x y
# 1: A 1
# 2: B 1
# 3: E 1
# 4: F 2
# 5: E 2
# 6: H 2
```
<!-- }}} -->

:::: {.columns}

### Create installed packages for toy example:
::: {.column width="50%"}
<!-- {{{ -->


```{r create_inst, style="background: lightgreen;" }
inst  <- structure(
      list(x = c("A", "B", "C", "D", "E", "F", "G", "H", "E", "H"), 
           y = c(1, 1, 1, 1, 2, 2, 2, 2, 1, 1)), 
      row.names = c(NA, -10L), 
      class = c("data.table", "data.frame")
      )
```
<!-- }}} -->
:::

::: {.column width="20%"}

```{r collapse = TRUE, bg-success, class.source="bg-success"}
setorder(inst)[]
```
:::

::::
### Examine install packages

<!-- {{{ -->
From inspecting inst: you can see   

- pkgs A-E  are all in directory 1;  
- pkgs E-H are in directory 2;  
- Note:  pkg **E and H are in BOTH 1 and 2**.  

This occurs in my setup when E,H in the past were upgraded.   R alerted me that
directory 2 was read only and allowed me to choose another directory.  I chose
1.
<!-- }}} -->

### Create update candidates (old.packages)
<!-- {{{ -->

```{r}
cand  <-  structure(list(x = c("A", "B", "E", "F", "E", "H"), y = c(1L, 1L, 1L,
                                                                    2L, 2L,
                                                                    2L)),
                    row.names = c(NA, -6L), class = c("data.table",
                                                      "data.frame"))
cand
```
<!-- }}} -->

### Examine candidate packages
<!-- {{{ -->


- dir 1, A,B,E are update candidates
- dir 2, E,F,H are update candidates
- Note:  E is in both 1 and 2.
- F is candidate, but not in directory 1
(- G is not candidate, so ignored.)
<!-- }}} -->

### Which packages will update?
<!-- {{{ -->

- In dir 1, all candidates will update (A,B,E,H)
- In dir 2, candidate F will update (to dir 1)

- Pkgs C,D will not update (no candidate)
- Pkgs G will not update (no candidate)
<!-- }}} -->

### Update process: cases

<!-- {{{ -->
**Case** **1** 
If the candidates are in directory 1, we can update these as is.

```{r case_1}
cand[y==1, ] 
```

However, if the candidates are in directory 2, there are a few cases.

**Case** **2A**
 If candidate was never installed in 1, upgrade to directory 1.
 Select only candidates in 2 that were NOT previously installed in 1 (anti-join).
```{r}
cand[y== 2, ][!inst[y==1, ], on=.(x)]
```

**Case** **2B**
If the candidate package is currently installed in 1, there is nothing more to check. Any update, if any, will be done in CASE I. (details below)
Case 1 will any candidate in 2 that was previously installed in 1. 
```{r}
cand[y ==2, ][inst[y == 1], on=.(x), nomatch=NULL]
```
<!-- }}} -->

***

<!--
### Actual update:

:::: {.columns}

::: {.column width="40%"}
```{r}
print("installed")
inst
```

:::

::: {.column width="60%"}
```{r}
print("cand")
cand
```
:::

::::
-->
```{r}
#| include: false
knitr::knit_exit()
```

### Details for Case 2B:
<!-- {{{ -->
  Case 2B   is really 2 cases.
  If the candidate is also a candidate in 1 then case 1 will handle and nothing more to be done.
```{r}
  cand[y==1,][cand [y == 2, ], on = .(x), nomatch=NULL]
```
And if candidate pkg is currently installed in 1, but NOT candidate in 1, then there is nothing to do.
 If the candidate is not a now candidate in 1 but was previously installed in  1 (ie past upgrade)
 false positive, ignore
```{r}
cand[y == 2] [!cand[y ==1],  on=.(x)][inst[y == 1,], on = .(x), nomatch=NULL]
```

```{r}


cand[y ==2, ][!inst[y == 1, ], on=.(x)]
```
<!-- }}} -->

```{r}
#| include: false
knitr::knit_exit()
```
```{r}

`A new "!" prefix on i signals 'not-join' (a.k.a. 'not-where'), #1384.
DT[-DT["a", which=TRUE, nomatch=0]]   # old not-join idiom, still works
DT[!"a"]                              # same result, now preferred.
DT[!J(6),...]                         # !J == not-join
DT[!2:3,...]                          # ! on all types of i
DT[colA!=6L | colB!=23L,...]          # multiple vector scanning approach
DT[!J(6L,23L)]                        # same result, faster binary search
'!' has been used rather than '-' :
* to match the 'not-join' and 'not-where' nomenclature
* with '-', DT[-0] would return DT rather than DT[0] and not be backwards
compatibile. With '!', DT[!0] returns DT both before (since !0 is TRUE in
base R) and after this new feature.
* to leave DT[+...] and DT[-...] available for future use``
```
