---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
        mathjax: default
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
    extra_dependencies: ["amssymb", "bbm", "threeparttable", "forest"]
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
TRY:   advanced-R,condition
---

###  Example of custom condition,
  *  Section    of Advanced R
  *   https://stackoverflow.com/a/15442036/5057251
  *  For my notes|tools on conditions see

---

\begin{forest}
  for tree={
    fit=band,% spaces the tree out a little to avoid collisions
  }
  [condition
    [simpleCondition, tier=condition
      [division_condition, tier=condition]
    [error, tier=abstract
      [simpleError, tier=error]
      ]
    [warning, tier=abstract
      [simpleWarning, tier=warning]
      ]
  ]
\end{forest}

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = T   ) # T easier to read

library(sloop)
```


For 1., I think of simpleCondition as illustrating how one can construct custom conditions, e.g,.
```{r from_stack }

# returns custom condition

 myCondition <-
    function(message, call=NULL, type=c("overflow", "underflow", "zero"))
{
    type <- match.arg(type)             # only allowed types past here
    class <- c(type, "my", "condition")
    structure(list(message = as.character(message), call = call), 
        class = class)
}

## is a constructor for making custom conditions

  cnd  <- myCondition("oops") 
  # <overflow: oops>
  sloop::otype(cnd) 
  class(cnd)

  myCondition("oops", type="underflow")
  # <underflow: oops>

  # bug, this will throw nasty error
  myCondition("oops", type="none")
```


These conditions can be used in tryCatch or withCallingHandlers
```{r tryCatch}

xx <- tryCatch({
    signalCondition(myCondition("oops", type="underflow"))
    }, 
    underflow=function(e) {
      message("underflow: ", conditionMessage(e))
      NA # return value, assigned to xx
    }
)
xx
```

These are S3 classes so can have a linear hierarchy -- bad and worse are both subclasses of error.
```{r linear}

myError <-
    function(message, call=NULL, type=c("bad", "worse"))
{
    type <- match.arg(type)
    class <- c(type, "error", "condition")
    structure(list(message=as.character(message), call=call),
              class=class)
}
```

One might also create an error that extends the 'simpleError' S3 class as cond <- simpleError("oops"); class(cond) = c("myerr", class(cond)

With tryCatch we just get access to a single handler, the first (in the sense described on ?tryCatch) to match the class of condition
```{r }


# class = worse 
tryCatch({
    stop(myError("oops", type="worse"))
    }, 

  bad = function(e) {
      message("bad error: ", conditionMessage(e))
  }, worse = function(e) {
      message("worse error: ", conditionMessage(e))  # here's where we end up
  }, error=function(e) {
      message("error: ", conditionMessage(e))
  }
)

# class = bad
tryCatch({
    stop(myError("oops", type="bad"))
    }, 

  bad = function(e) {
      message("bad error: ", conditionMessage(e))
  }, worse = function(e) {
      message("worse error: ", conditionMessage(e))  # here's where we end up
  }, error=function(e) {
      message("error: ", conditionMessage(e))
  }
)
# class =  "no class"
tryCatch({
    stop(myError("oops", type="no class"))
    }, 

  bad = function(e) {
      message("bad error: ", conditionMessage(e))
  }, worse = function(e) {
      message("worse error: ", conditionMessage(e))  
  }, error=function(e) {
      message("error: ", conditionMessage(e))   # <--- end up here
  }
)

```

With withCallingHandlers we have the opportunity to hit multiple handlers, provided we don't invoke a restart
```{r withCallingHandler}

withCallingHandlers({
    stop(myError("oops", type="bad"))
}, bad = function(e) {                             # here...
    message("bad error: ", conditionMessage(e))
}, worse = function(e) {
    message("worse error: ", conditionMessage(e))
}, error=function(e) {                             # ...and here...
    message("error: ", conditionMessage(e))
})                                                 # ...and top-level 'error'

withCallingHandlers({
    stop(myError("oops", type="worse"))
}, bad = function(e) {                             #  <-- not here
    message("bad error: ", conditionMessage(e))
}, worse = function(e) {
    message("worse error: ", conditionMessage(e))
}, error=function(e) {                             # <-- here and ...
    message("error: ", conditionMessage(e))
})                                                 # <-- here, top-level 'error'

```
```{r warning}

withCallingHandlers({
    x <- 1
    warning(myError("oops", type="bad"))
    "OK"
}, bad = function(e) {                     # here, but continue at the restart
    message("bad warning: ", conditionMessage(e))
    invokeRestart("muffleWarning")
}, worse = function(e) {
    message("worse warning: ", conditionMessage(e))
})

I'm not so sure about your question 2; I think this is the situation that calling handlers are designed to address -- the entire frame where the condition was invoked is poised waiting to continue, once you invoke the restart. 
```



```{r render, eval=FALSE, include=FALSE 	} 
# from ~/.Renviron
output_dir  <- Sys.getenv("OUTPUT_DIR")
output_dir
{
file  <- "/home/jim/code/book_advanced_R/R/008_custom_conditions.Rmd" 
file  <- basename(file)
file  <- here("ADVANCED_R", "R", file)
file
}


## Want github to display nicely rendered output?
##  *  include md_format
#   *  include output_dir = "out"
#   *  if .gitignore balks,  then add !out/**   to close .github

rmarkdown::render(file,
    output_dir = output_dir,
    output_format = "pdf_document")

)
```
