---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
        mathjax: default
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
    extra_dependencies: ["amssymb", "bbm", "threeparttable", "forest"]
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
TAGS:  latex, trees, forest, condition
---


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = T   ) # T easier to read
```


### Probe condition
---

\begin{forest}
  for tree={
    fit=band,% spaces the tree out a little to avoid collisions
  }
  [condition
    [simpleCondition, tier=condition
      [division_condition, tier=condition]
      ]
    [error, tier=abstract
      [simpleError, tier=error]
      ]
    [warning, tier=abstract
      [simpleWarning, tier=warning]
      ]
  ]
\begin{forest}
  for tree={
    fit=band,% spaces the tree out a little to avoid collisions
  }
  [condition
    [simpleCondition, tier=condition
      [division_condition, tier=condition]
      ]
    [error, tier=abstract
      [simpleError, tier=error]
      ]
    [warning, tier=abstract
      [simpleWarning, tier=warning]
      ]
  ]
\end{forest}

\begin{forest}
  for tree={
    fit=band,% spaces the tree out a little to avoid collisions
  }
  [things
    [cabbages, tier=vegetables
      [peaches, tier=fruits]
    ]
    [kings, tier=aristocrats]
    [sealing wax
      [queens, tier=aristocrats
        [carrots, tier=vegetables]
        [pineapple, tier=fruits]
        [aubergine, tier=vegetables]
      ]
    ]
  ]
\end{forest}


#### Probe
```{r probe}
## constructors of class with same name
  warning
  message

  simpleCondition
  simpleError
  simpleWarning
  simpleMessage

## but these have no constructor
  condition
  error

## find 4 methods associated with class = 'condition'
 sloop::s3_methods_class(x = "condition")

## generic functions + print + as.character
  sloop::ftype(conditionCall)
  sloop::ftype(conditionMessage)

## internal functions
  sloop::ftype(warning)
  sloop::ftype(signalCondition)

## methods
  methods(conditionCall)
  methods(conditionMessage)

f  <- function(...) {
  cat(nargs(), "\n")
  args  <- list(...)}

r  <- f("one")
r
## SimpleCondition
 methods(simpleCondition)
 ftype(simpleCondition)

 is.object(simpleCondition)
# [1] FALSE


## find all methods associated with a class

 args(sloop::s3_methods_class )

 sloop::s3_methods_class(x = "factor")
 sloop::s3_methods_class(x = "error")
 sloop::s3_methods_class(x = "condition")


 # why nothing?

 sloop::s3_methods_class(x = "warning")
 sloop::s3_methods_class(x = "message")
 sloop::s3_methods_class(x = "simpleCondition")
 sloop::s3_methods_class(x = "simpleError")
 sloop::s3_methods_class(x = "simpleWarning")
 sloop::s3_methods_class(x = "simpleMessage")





sloop::s3_dispatch(conditionCall( signalCondition(stop())))
sloop::s3_dispatch(conditionCall( signalCondition(stop())))
sloop::s3_dispatch(warning())


## Constructor
  simpleMessage

```

###
```{r objects}

## S3 simpleCondition
    s  <- simpleCondition("example: simpleCondition ", call = T)
    class(s)
    attributes(s)
    sloop::otype(s)
    s$message
    s$call

    x  <- sloop::s3_methods_generic("simpleCondition")
    str(x)
    sloop::s3_methods_class("simpleCondition")
    sloop::s3_methods_class("simpleError")
    sloop::s3_methods_class("factor")
    

    simpleCondition



s  <- simpleError("error")
class(s)
sloop::otype(s)



```




### UseMethod  - wrong!
```{r UseMethod}
h  <- function(generic,object ) {
  UseMethod(generic,  object) 
}

```{r skeleton, include=F, eval=F}
file="/home/jim/.config/nvim/templates/skeleton.Rmd"
```

```{r render, eval=FALSE, include=FALSE 	} 
#  [For technical notes]: (~/docs/technical_notes/001_render_output.md)


{
file <- "ADVANCED_R/R/008_condition_probe.Rmd"
file  <- basename(file)
file  <- here("ADVANCED_R","R", file)
file
}
output_dir  <- Sys.getenv("OUTPUT_DIR")
output_dir



rmarkdown::render(file,
    output_format = "pdf_document",
    #output_format = "html_document",
    #output_format=c("html_document", "md_document"),
    output_dir = output_dir)
    #output_file = "out")
```
