---
title:  ""
output: 
  pdf_document:
    toc: TRUE 
    toc_depth: 2
    latex_engine: xelatex
fontsize: 12pt
geometry: margin=0.5in,top=0.25in
#  md_document:
#  html_document:
#  github_document:
#   toc: true
#  	toc_float: true
#    css: styles.css
params:
  date: !r lubridate::today()
---
`r params$date`

(TODO:  needs cleanup)

## Chapter 13 
*  [intro, cheat sheet, see:](~/code/try_things_here/rmd/)
*  (see vctrs:: for practical S3 )
*  This file:  study hadley's tools and ideas
*  Next file:  Build a class


```{r one, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  
                      comment="      ##",  
                      error=TRUE, 
                      collapse=TRUE)
library(purrr)
library(sloop)
library(jimTools)
```

```{r summary}

####  base
is.object
methods
typeof
class


#### hadley
sloop::otype
sloop::ftype
sloop::s3_dispatch
sloop::s3_dispatch(print(f))

```


## 13.2 Basics
*  class
*  attributes

```{r 13.2}
f <- factor(c("a", "b", "c"))
f
typeof(f)   # base = int[]
class(f)	  # 'factor'
attributes(f) # includes 'levels' which hold data as presented ('a', 'b', ....)

# compare, when no longer 'factor' still has attribute "levels"
g  <- unclass(f)   # removes class, not other attributes
g
typeof(g)
class(g)
# so does 
class(1:3)

attributes(g)
```

## Find methods for a generic function
```{r methods}
methods(factor)
methods(print)   # chr[] , many!
methods(str)
methods(mean)
```

## structure
```{r structure}

# Create and assign class in one step
x <- structure(list(), class = "my_class")
class(x)

# check;
```

## S3 & generic functions(really an interface)
*  sloop::ftype (generic or not)
```{r s3}
# Given an S3 object, generic functions process  based on class.

# is a function S3 generic?  
sloop::ftype(print) # yes, if see 'generic'
ftype(str)  # yes
ftype(cat)  #no, internal
ftype(lm)   # no
ftype(list)  # no, primitve
ftype(as.character)  # primitive generic S3?? 
ftype(mean) # yes
ftype(factor) # no

```
### Don't get fooled  (don't get it) 
```{r caution}
(d  <- date())
class(d)
# tricky!  POSIXlt is S3, based on list
# BUT str() is generic and designed to fool! Hides the list!
str(d)
sloop::otype(d)
```

### s3_dispatch(f) , list methods for generic function.
  * Format for methods is <generic>.<class>
```{r dispatch}
s3_methods_generic("print")
s3_methods_generic("factor")
# <generic>.<class>
# f is factor (above); show which method selecdted
# `=>` is method UseMethod() will use
# `*`  method is not used

sloop::s3_dispatch(print(f))
sloop::s3_dispatch(print(c(1,2,3,4)))
sloop::s3_dispatch(print(1 + 1))

# knows it is a function!
sloop::s3_dispatch(print(function(x) {x}))

sloop::s3_dispatch(cat("hello"))
sloop::s3_dispatch(factor(c(1,2,3)))
```

### UseMethod  - wrong!
```{r UseMethod}
h  <- function(generic,object ) {
  UseMethod(generic,  object) 
}

f.character <- function(x) print("character")
f.numeric  <- function(x) print("numeric")
class("a")
# [1] "character"

class(2)
# [1] "numeric"

h(generic = "f",object = "a")
h("f", 2)
```

## Chapter 13 
*  (see vctrs:: for practical S3 )
*  Last file:  study hadley's tools and ideas
*  THIS file:  Build a class

## Outline:
  *   By adding class attribute to base types can create objects behave
    differently under generic functions (such as when printed).
  *  S3 is relatively simple OO, but ALSO allows easy errors; hence structure
    Hadley seeks to impose.
  * constructor (not called directly for this reason), validator (to check) and
    friendly helper to actually create object (safely).
  * need generic function (as as print) to create variants or methods.  
  * which variant is determined by dispatch | UseMethod()
  * helpers - numerous to determine dim, class, names, attributes.
  * sloop::  to help see under the hood



## Constructor, example
*  Base class (double)
*  add class
*  add structure
*  some rumentary checks
```{r constructor}

new_difftime <- function(x = double(), units = "secs") {
  stopifnot(is.double(x))
  units <- match.arg(units, c("secs", "mins", "hours", "days", "weeks"))
  structure(x,
    class = "difftime",
    units = units
  )
}

r <- new_difftime(52, "weeks")
class(r)
is.double(r)
# r or print(r)  only one method
r
print(r)   # SAME



# another
r  <- new_difftime(c(1, 10, 3600), "secs")
r
```

## More complex S3, need for validator
```{r validator}

new_factor <- function(x = integer(), levels = character()) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))

  structure(
    x,
    levels = levels,
    class = "factor"
  )
}

r  <- new_factor(c(1,10,100), levels = c("low", "medium", "high"))
class(r)
is.integer(r) #F
is.character(r) #F
is.double(r)
typeof(r)


r  <- new_factor(1:3, levels = c("low", "medium", "high"))
r
is.integer(r) #F 
is.double(r)  #F


# another, all FAIL
new_factor(1:5, c("a"))
new_factor(1:5, "a")
new_factor(0:1, "a")

```

