
--- 
title:  "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output: 
  html_document:  
    toc: true 
    toc_depth:  2 
    toc_float:  true   
    number_sections: yes
  pdf_document: 
    toc: true 
    toc_depth:  2 
    latex_engine: xelatex
    fontsize: 10pt 
    geometry: margin=0.5in,top=0.25in 
---
## Misc useful stuff
## Tags:  stop(), str_c(), function(...), row_number(),  

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read
library(forcats)
library(dplyr)
library(stringr)
library(knitr)
```

#### stop(), throws error 
```{r}  
x <- 0
if (x == 0) stop("this is msg")   
print("never gets here")
```

#### use of "..."
```{r}
stringr::str_c("a", "b")   # ab
commas <- function(...) {
        # pass ... to functions inside
        # collapse into single string
        stringr::str_c(..., collapse=", ")
}

commas(letters[1:10])           # a, b, c, d, e 

stringr::str_c("a","b", "c", sep=", ")
```
```

```{r mod}
#  using  7 %% 2 as 7 mod(2) is equal 1
#  8 %% 2 = 0

even <- 
```




```{r dplyr}
# ---- rank, row_number() -------------

t <- tibble(x = 10:1,
            letters = letters[1:10])
t %>% factor(letters)
t 

# each group, find lowest 2 scores, print original line#
df %>% as.tibble() %>%
        group_by(factor) %>%
        mutate(n = row_number()) %>%  #note:  WITHIN group
                       select(n,factor,score) %>%
       
        arrange(desc(factor), desc(score)) %>%
        top_n(-2)
       
```
```{r sQL}


# ---- show SQL -----------------
q <- df %>% as.tibble() %>%
        group_by(factor) %>%
        mutate(n = row_number()) %>%  #note:  WITHIN group
        select(n,factor,score) 
q %>% dplyr::show_query()
q %>% explain()

```
```{r render, eval=FALSE    }

file  <- here("", "")
file  <- "rmd/00010_dplyr_misc_factor_examples.Rmd"

# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
