
--- 
title:  "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output: 
  html_document:  
    toc: true 
    toc_depth:  2 
    toc_float:  true   
    number_sections: yes
  pdf_document: 
    toc: true 
    toc_depth:  2 
    latex_engine: xelatex
    fontsize: 10pt 
    geometry: margin=0.5in,top=0.25in 
---
## Misc useful stuff
## Tags:  stop(), str_c(), function(...), row_number(),  factor, forcats

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read
library(forcats)
library(dplyr)
library(stringr)
library(knitr)
```

#### stop(), throws error 
```{r}  
x <- 0
if (x == 0) stop("this is msg")   
print("never gets here")
```

#### use of "..."
```{r}
stringr::str_c("a", "b")   # ab
commas <- function(...) {
        # pass ... to functions inside
        # collapse into single string
        stringr::str_c(..., collapse=", ")
}

commas(letters[1:10])           # a, b, c, d, e 

stringr::str_c("a","b", "c", sep=", ")
```
from Grolemund - categorical variables
Ref: https://rstudio-education.github.io/hopr/r-objects.html Ch 5.5.2

>To make a factor, pass an atomic vector into the factor function. R will
>recode the data in the vector as integers and store the results in an integer
>vector. R will also add a levels attribute to the integer, which contains a
>set of labels for displaying the factor values, and a class attribute, which
>contains the class factor:
```{r factors, collapse=TRUE}
# recode chr[] as integer
gender <- factor(c("male", "female", "female", "male"))

typeof(gender)
attributes(gender)   # levels, class

# internally, stored as integer
unclass(gender)  # as  integers, with level attr


# R uses the levels attribute when it displays the factor, as you will see.

# displays as levels
gender
```


#### 0050_factors (r4ds: Chapter 15, forcats)
#### factor: Peng, think ints with each having a label
```{r factors2, collapse=TRUE}
# create factors
levels <- c("a","b")

# "c" becomes "NA"  NO ERROR
x <- c("a","b","b","b","c")

x <- c("a","b","b","b")  # chr[]


# factor: closer look
# store as integer, use levels for display
y <- factor(x, levels=levels)
y

typeof(y)
attributes(y)  # levels, class 

# also

levels(y)
class(y)
str(y)

# nice features
table(y)
names(y)
labels(y)

unclass(y)	# underneath, stored as ints
names(unclass(y))
labels(unclass(y))
```

####10_base sort
```{r base}
# create data another way
x <- rep(c("a","b"),6)  # chr[]
x

# turn into factora
levels=c("a","b")
y <- factor(x, levels=levels)
y


sort(y)  # sort the underlying ints, then display with levels
```
https://forcats.tidyverse.org/index.html
#### forcats::fct_*
```{r forcats1}
# create random data from 0:9
x <- 0:9
x

# Create sample of 100 by:
# Choose 1 value from x, then replace it.
# Repeat 100 times

set.seed(2021)
y <- sample(x, size=100, replace=TRUE)
y  # int[100]


# only change is create factor
z <- factor(y, levels=0:9)
z

# nice freq table  5 is most common (17)
table(z)  

####
# Data internally or display, does not change, added class and levels
# Only change:  order of factors
####

# order factors by freq
forcats::fct_infreq(z)
# keep largest 3, lump rest into "other"
forcats::fct_lump(z, n=3)
forcats::fct_lump(z, n=3, other_level = "Z")



# order factors by 1st appearance 6 5 ...
forcats::fct_inorder(z)

# shuffle, again no change to underlying data or display, only factors
forcats::fct_shuffle(z)

# again
a  <- forcats::fct_shuffle(z)
a
```

```{r forcats2 }
f <- factor(c("a", "b", "c"))
f

fct_rev(f) #> [1] a b c
g  <- fct_rev(f)
g
levels(g)

fct_inseq(g)
```

```{r mod}
#  using  7 %% 2 as 7 mod(2) is equal 1
#  8 %% 2 = 0

even <- 
```




```{r dplyr}
# ---- rank, row_number() -------------

t <- tibble(x = 10:1,
            letters = letters[1:10])
t %>% factor(letters)
t 

# each group, find lowest 2 scores, print original line#
df %>% as.tibble() %>%
        group_by(factor) %>%
        mutate(n = row_number()) %>%  #note:  WITHIN group
                       select(n,factor,score) %>%
       
        arrange(desc(factor), desc(score)) %>%
        top_n(-2)
       
```
```{r sQL}


# ---- show SQL -----------------
q <- df %>% as.tibble() %>%
        group_by(factor) %>%
        mutate(n = row_number()) %>%  #note:  WITHIN group
        select(n,factor,score) 
q %>% dplyr::show_query()
q %>% explain()

```
```{r render, eval=FALSE    }

file  <- here("", "")
file  <- "rmd/00010_dplyr_misc_factor_examples.Rmd"

# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
