---
title: "020_"
output: html_document
editor_options: 
  chunk_output_type: console
---

## httr:: Examples
Ref: https://httr.r-lib.org//reference/GET.html
* vignette - uses httpbin.org as practice
* Wiki for REST (an architet using, using HTTP, URIs, payloads ....)
* Study various options, httr_options()
* Specific option:   curl_docs("useragent")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(httr)
library(jsonlite)
```

## simple google search
```{r}
r<- GET("http://google.com/", 
        path = "search", 
        query = list(q = "falwell"))

# many ways to examine response

# two ways to look at header
r$headers %>% str(max.level=1 )
httr::headers(r)

# no error
httr::http_error(r)
httr::status_code(r)
httr::cookies(r)
content(r)
content(r, "text")
```


## httr::modify_url()
```{r}
the_base  <- "https://httpbin.org"
r  <- httr::GET(the_base)

# usually, we have some kind of path or endpoint
the_url  <- modify_url(the_base, path="/get")
GET(the_url)
```

variety of curl options, and how to use in R.
```{r change}
ua  <- user_agent("httr")
GET(the_url, ua )

# now without ua
GET(the_url)

#  with ua
GET(the_url, config=ua)


# without ua
GET(the_url)

# hmmm,  adds to default user_agent()
with_config(GET(the_url), config=config(user_agent("httr")))

```
temporarily change config
```{r with_config}
GET(the_url, user_agent("httr"))
with_config(the_url, user_agent("httr"))
```


Focus on httpbin.org practice
## headers
```{r}
# add header
base_url <- "http://httpbin.org"
r <- httr::GET("http://httpbin.org/get", 
               add_headers(a="ONE_header", b="TWO_header"))
r
```

```{r debug}
Verbose, useful to debug
'<-' header in
'<<' data in
'*'  msg from curl
```
## verbose, content
```{r verbose}
r  <- httr::GET(base_url, verbose())
http_type(r)
r

# vector
content(r)

# stream of text
content(r, "text")
```

Add query, parameters relatd to useful to the api
```{r query}
# http:://httbgin.org/?name=jim&occupation=student
# Note:  ? begins qery, & separates
r  <- httr::GET(base_url, query=list(name="jim", occupation="student"))
r
```

## Return json and process
```{r json}
# api says add /json
r  <- GET(modify_url(base_url, path="/json"))
http_type(r)
names(r)
content(r, as="text") %>% jsonlite::fromJSON()

```

verbose, body
```{r}
Use body, to pass parameters
r  <- with_verbose(GET(base_url, body=list(name="jim", encode="json"),
                       content_type_json()))
```

## Use S3 to examine response
```{r S3}
# make function
httpbin_api  <- function(path){
  # path = "/json"
  r  <- GET(modify_url(base_url, path=path))
  http_type(r)
  names(r)
  parsed  <- content(r, as="text") %>% jsonlite::fromJSON()

  # Make a structure
  structure (
           list(content = parsed,
                path    = "/json",
                response = r
                ),
           class = "httpbin_api"
           )
}

print.httpbin_api<- function(x, ...) {
  cat("<httpbin ", x$path, ">\n", sep = "")
  str(x$content)
  invisible(x)
} 

httpbin_api("/json")


```
```{r}

endpoint <- "https://api.noopschallenge.com/hexbot"


## get a single color, and display on screen

random.color <- function() {
    hexbot <- GET(endpoint)
    content(hexbot)$colors[[1]]$value
}

rcol <- random.color()
barplot(1, col=rcol, main=rcol, axes=FALSE)
## --------
## get n colors at once
## get n colors at once

random.colors <- function(n) {
    hexbot <- GET(endpoint, query=list(count=n))
    unlist(content(hexbot)$colors)
}

ncol <- 5
barplot(rep(1,ncol),col=random.colors(ncol), axes=FALSE)

```
