	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

 ======================================================
withr::  way to temporily change values and then reset
 ======================================================

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw

vim: to format all urls for md
s/https.*/[&]()/g
-->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
```


```{r library, include=FALSE		}
library(jimTools)
library(withr)
``` 

change state:   poor
```{r begin}
sloppy <- function(x, sig_digits) {
  options(digits = sig_digits)
  print(x)
}
# old state
old  <- getOption("digits")    # 3
old
pi

# now
sloppy(pi, 4)
# but state has changed!
pi


#  reset
options("digits"  = old)
pi
```
withr
```{r withr}
neater <- function(x, sig_digits) {
  # store old in op; 
  op <- options(digits = sig_digits)
  # put on LIFO stack, to change back when functions ends
  withr::defer(options(op))
  print(x)
}

pi
neater(pi, 7)
pi
```

One defer, put on stop of defer stack
```{r lifo}
defer_stack <- function() {
  cat("put on socks\n")
 withr::defer(cat("take off socks\n"))
}

defer_stack()
```

Two defer, use LIFO 
```{r lifo2}
defer_stack <- function() {

 cat("put on socks\n")
 withr::defer(cat("1-take off socks\n"))  # on stack

  cat("put on shoes\n")
  withr::defer(cat("2-take off shoes\n")) # on top of stack
}
defer_stack()  # use LIFO
```
Want FIFO?
```{r fifo}
# use priorty = last
defer_queue <- function() {
  cat("Adam gets in line for ice cream\n")
  withr::defer(cat("Adam gets ice cream\n"), priority = "last")

  cat("Beth gets in line for ice cream\n")
  withr::defer(cat("Beth gets ice cream\n"), priority = "last")
}
defer_queue()
```


BEGIN HERE:   stuck on environments again.
improved neater
```{r improve}
sys.nframe()  # Global
local_digits <- function(sig_digits, envir = parent.frame()) {
  op <- options(digits = sig_digits)
  print(envir)
  withr::defer(options(op), envir = envir)

}
parent.frame()  # Global

neato <- function(x, digits) {
  sys.nframe()
  local_digits(digits)
  print(x)
}
pi
local_digits(3)
neato(pi, 7)
pi


```

***

```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file <- ""
file  <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
