---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:
  html_document: 
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 2
    toc_float: true
  pdf_document:
    toc: yes
    number_sections: TRUE
geometry: margin=1in
editor_options: 
  chunk_output_type: console
---

## Collect knitr:: here

echo[TRUE] print code  
eval[TRUE] eval cod  
include[TRUE] print output  

## 0-setup, no output
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  
                      comment="      ##", 
                      z=TRUE, 
                      collapse=TRUE)
library(knitr)
library(jimTools)
```


```{r TAGS, eval=FALSE, echo=FALSE}
TAGS <- c("collapse", "chunks", "examples")
```

## 1-evaluate, but include no output
```{r test, eval=TRUE, include=FALSE}
print("hi")
x=4
show_text=FALSE
c_ECHO = FALSE
c_EVAL = FALSE
c_INCLUDE = FALSE


```

## 2-output, but do not evaluate
```{r check, eval=FALSE}
print(x)
```

## 3-nothing!
```{r check1, eval=FALSE, include=FALSE}
print(x+2)
```

3A-results='hide', code only
```{r results='hide'}
print(x+2)
print("3A-results=hide")
```

3B-NOT results='hide'
```{r, collapse, collapse=TRUE}
print(x+2)
print("3B-results=NOT hide")
```

4A-conditional
```{r conditional_block, echo=show_text, eval=show_text, include=show_text }

# see in 1-setup
print("show_text is TRUE")
```

4B-conditional
```{r conditional_blockB, echo=!show_text, eval=!show_text, include=!show_text }

# see in 1-setup
print("show_text is TRUE")
```

## 7A-collapse=TRUE
```{r collapse=TRUE }
print("one")
print("two")
```

## 7B-collapse=FALSE
```{r, label="7B", collapse=FALSE, eval=c_eval, include=c_include, echo =
  c_echo}
print("one")
print("two")
```


## 20-list extra-all chunk options
```{r chunk_options, include=show_text, collapse=TRUE}
str(knitr::opts_chunk$get())
```



## 21-knitr::read_chunk from external .R file WORKS!
[REF:] (https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html)
```{r read, cache=FALSE }

#  knitr::read_chunk('~/code/r_try_things_here/rmd/file_for_read_chunk.R')
x  <- knitr::read_chunk("file_for_read_chunk.R")


```

## 22-knit_code(), include code from label="7B"
```{r, echo = !c_ECHO, eval = !c_EVAL, include=!c_INCLUDE, code=knitr::knit_code$get("7B")}

```


(I stole this code)
figure this out!
* find all chunks
* neatly numbers them?

```{r challenge_regex, eval=FALSE , echo=FALSE  }

name_chunks_mlr3book = function() {

  here()
  # full.names = FALSE for file names only
  rmds = list.files(here("rmd"), 
    pattern = "\\.Rmd$", full.names = TRUE, recursive = TRUE)

  rmds

  pattern = "^([[:space:]]*```\\{[rR])([[:alnum:] -]*)(.*\\})[[:space:]]*$"
  
  for (rmd in rmds) {
    message(sprintf("Renaming chunks in '%s'", basename(rmd)))

    # read all the lines
    lines = readLines(rmd)

    # which lines, spaces, then ```, anything, then }
    # misses a few with ::spaces:: after final }
    ii  <- which(
                 stringr::str_detect(
                                     lines, 
                                     "^[[:space:]]*```\\{[rR].*\\}$"))


    # START HERE
    labels = sprintf("%s-%03i", stringi::stri_replace_last_fixed(basename(rmd), ".Rmd", ""), seq_along(ii))

    lines[ii] = stringi::stri_replace_first_regex(lines[ii], pattern, sprintf("$1 %s$3", labels))

    writeLines(stringi::stri_trim_right(lines), con = rmd)
  }

  invisible(TRUE)
}
```
STop
```{r}
knitr::knit_exit()
```

render from code, to html
```{r eval=FALSE}

file  <- "00064_knitr_chunk_practice.Rmd"
file  <- here("rmd", file)
file
rmarkdown::render(
	file, 
	rmarkdown::html_document(), 
#	rmarkdown::pdf_document() ,
  output_file="~/Downloads/print_and_delete/out") 
```

