---
title: "085B_R_ENV_variables.Rmd"
output:  
  html_document:
        code_folding: show
        toc: true 
        toc_depth: 3
        toc_float: true
  pdf_document:
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
TAGS:  Sys.getenv(),options() ,R environment variable, 
---

 includes:  021_efficient_coding	

### PURPOSE:  understand how R uses options() and Sys.getenv()
###  Uses of Sys.getenv() and options()
```{r setup}
knitr::opts_chunk$set(echo = TRUE, 
                      comment = "        ..##")


library(tibble)
library(tidyr)
library(dplyr)
library(magrittr)
library(xfun)
```

## R environment variables
*  See: https://stat.ethz.ch/R-manual/R-devel/library/base/html/EnvVar.html
*  get/set , listing, searching for 


### examine Sys.getenv()
```{r main, eval=TRUE, inlcude=TRUE}

l  <- c("R_HOME", "R_LIBS_USER", "R_HIST_SIZE", "R_BROWSER")
sapply(l, Sys.getenv)

.libPaths()



# examine character vector
x  <- Sys.getenv()

library(data.table)
dt  <- data.table(x)
dt[1:5,]


```

## grep on names of R environment variables. 
```{r grep}
grep(x=names(Sys.getenv()),
     pattern = "dirs",
     value=TRUE,
     ignore.case=TRUE)


# get elment #
n  <- grep(x=names(Sys.getenv()),
     pattern = "dirs",
     value=FALSE,
     ignore.case=TRUE)
# display
Sys.getenv()[n]


# for "libs"
n  <- grep(x=names(Sys.getenv()),
     pattern = "libs",
     value=TRUE,
     ignore.case=TRUE)
Sys.getenv()[n]
```




```{r .libPaths, eval=FALSE, echo=FALSE}

#### 4-.libPaths(), list all paths, purpose, name

.libPaths()
# fix:

# t_libPaths <- tibble(.libPaths()) %>% 
#   mutate(
#   purpose =  c("added packages", 
#                ".Library.site", 
#                ".Library.site", 
#                "base R pkgs"),  
# 
#   quantity  = c("many", "Empty", "Empty", "many")  ,
#   env_var = c("R_LIBS_USER","","","R_LIBS?"),
#   aka = c("",".Library.site", ".Library.site", ".Library.site, .Library"))
# 
# t_libPaths
```

## Each path in .libPaths(), list installed packages

```{r list.dirs}
.Library
if (FALSE)
  .Library
  lapply(.libPaths(), list.dirs, recursive=FALSE)
```


```{r exit, include=FALSE}
knitr::knit_exit()
```

### Session_info
```{r session_info}

# yihui's improved
xfun::session_info()

# note:  results is S3 object
(z  <- utils::sessionInfo())
sloop::otype(z)
# methods that handle this object?
sloop::s3_methods_class(z)

# stack is generic function for object z



session_info()
```





## R help
```{r}
# REF https://csgillespie.github.io/efficientR/set-up.html

?mpg   # mpg is dataset (if loaded)
?stats  # package
?reprex # function
?ggplot   # function
??reprex  # topic
??ggplot   # topic 
??pch      # topic 
??regression

help.search("pch")  # long form


## R Startup
```{r startup}

# In this order, R seeks .Rprofile $R_HOME, $HOME, project directory
?Startup
R.home() 
site_path = R.home(component = "home") # 
fname = file.path(site_path, "etc", "Rprofile.site") # /usr/lib/R/etc/Rprofile.site

file.exists(fname) # True
```


## .Rprofile
```{r Rprofile}

file.exists("~/.Rprofile") # True, also

options()
# 2 sign digits  (3.7)
options(prompt = "R> ", digits = 2, show.signif.stars = FALSE, continue = "  ")

getOption("repos")  # cran.rstudio.com

fortunes::fortune(50)  # error
.Last = function() {
        
        message("Goodbye at ", date(), "\n")
}
.Last()
```

```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

file  <- here("rmd", "00085B_R_ENV_variables.Rmd")
file 
# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# in general, pdf will look nicer

```



