---
title: "085B_R_ENV_variables.Rmd"
output:  
  html_document:
        code_folding: show
        toc: true 
        toc_depth: 3
        toc_float: true
  pdf_document:
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
---

 includes:  021_efficient_coding	

```{r setup}
knitr::opts_chunk$set(echo = TRUE, 
                      comment = "        ..##")

       #               eval = FALSE,
       #               include = FALSE)

knitr::current_input()  # NUL

library(tibble)
library(tidyr)
library(purrr)
library(dplyr)
library(magrittr)
library(xfun)
```



## ENV variables
*  See: https://stat.ethz.ch/R-manual/R-devel/library/base/html/EnvVar.html
*  get/set , listing, searching for 

```{r main, eval=TRUE, inlcude=TRUE}
# /usr/lib/R
(r_home <-R.home(component="home"))
Sys.getenv("R_HOME") 			# /usr/lib/R ; where R installed
Sys.getenv("R_LIBS_USER") # ~R/x86 .../4.0
Sys.getenv("R_HIST_SIZE")
Sys.getenv("R_BROWSER")   # croutonurlhandler

.libPaths()

# yihui's improved
xfun::session_info()

# too lengthy
if  (FALSE) Sys.getenv()		

length(Sys.getenv())  # >100

```

## grep on names of R environment variables. 
```{r grep}
grep(x=names(Sys.getenv()),
     pattern = "dirs",
     value=TRUE,
     ignore.case=TRUE)


# get elment #
n  <- grep(x=names(Sys.getenv()),
     pattern = "dirs",
     value=FALSE,
     ignore.case=TRUE)
# display
Sys.getenv()[n]


# for "libs"
n  <- grep(x=names(Sys.getenv()),
     pattern = "libs",
     value=TRUE,
     ignore.case=TRUE)
Sys.getenv()[n]
```




```{r .libPaths, eval=FALSE, echo=FALSE}

#### 4-.libPaths(), list all paths, purpose, name

.libPaths()
# fix:

# t_libPaths <- tibble(.libPaths()) %>% 
#   mutate(
#   purpose =  c("added packages", 
#                ".Library.site", 
#                ".Library.site", 
#                "base R pkgs"),  
# 
#   quantity  = c("many", "Empty", "Empty", "many")  ,
#   env_var = c("R_LIBS_USER","","","R_LIBS?"),
#   aka = c("",".Library.site", ".Library.site", ".Library.site, .Library"))
# 
# t_libPaths
```

## Each path in .libPaths(), list installed packages

```{r list.dirs}
.Library
if (FALSE)
  .Library
  lapply(.libPaths(), list.dirs, recursive=FALSE)
```


```{r exit, include=FALSE}
knitr::knit_exit()
```


TODO
##  options()
```{r options}
# options() (returns <list>)
# Tibble 87 x 2.  values is list-column.
# Goal:  longer tibble, with 1 row for each value

l  <- options()
# find TRUE


l2  <- purrr::discard(l, ~is_function(.x))
l2
l3   <- purrr::discard(l2, ~(typeof(.x) == "language"))
length(l3)

l3[31:40]
l  <- l3
t <- tibble(name = names(l), values = l)
map(t$values, ~typeof(pluck(1)))
t$typeof  <- typeof(pluck(t$values, 1)) 
head(t)

# Problem, options()[[5]] is 'closure' and not a vector, list.
t %>% tidyr::unchop(values)

options() %>% dplyr::filter(rlang::is_list(values))


typeof(options()[[5]])
t %>% unchop(values)


typeof(pluck(t,2,1))
typeof(pluck(t,2,2))
typeof(pluck(t,2,5))
pluck(t,2,5)
pluck(t,2,10, 1)
pluck(t,2,10, 1)

#### TODO
t %>% tidyr::unnest_wider(options)
tidyr::unnest_auto(t, options)

#### another way
l  <- options()
names(l) #88

#### find element of vector
l["warning.length"]
l["pdfviewer"]	  # /usr/bin/xdg-open
l["keep.source"]  # TRUE
l["browser"]	 # croutonurlhandler

```





## R help
```{r}
# REF https://csgillespie.github.io/efficientR/set-up.html

?mpg   # mpg is dataset (if loaded)
?stats  # package
?reprex # function
?ggplot   # function
??reprex  # topic
??ggplot   # topic 
??pch      # topic 
??regression

help.search("pch")  # long form


## R Startup
```{r startup}

# In this order, R seeks .Rprofile $R_HOME, $HOME, project directory
?Startup
R.home() 
site_path = R.home(component = "home") # 
fname = file.path(site_path, "etc", "Rprofile.site") # /usr/lib/R/etc/Rprofile.site

file.exists(fname) # True
```


## .Rprofile
```{r Rprofile}

file.exists("~/.Rprofile") # True, also

options()
# 2 sign digits  (3.7)
options(prompt = "R> ", digits = 2, show.signif.stars = FALSE, continue = "  ")

getOption("repos")  # cran.rstudio.com

fortunes::fortune(50)  # error
.Last = function() {
        
        message("Goodbye at ", date(), "\n")
}
.Last()
```
```{r from_tan}
library(tidyverse)
l <- options() %>% 
  list() %>% 
  tibble() %>% 
  unnest_longer(1,indices_to = "option",values_to = "value")
l
```
```{r enframe}
library(tibble)
library(purrr)
options()
# list  -> nested tibble?
t  <- tibble::enframe(options())
head(t)

```

```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

file  <- here("rmd", "00085B_R_ENV_variables.Rmd")
file 
# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# in general, pdf will look nicer

rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```



