	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

PURPOSE:  practice data.table (dt)

SEE example(data.table)
SEE vignette(package=)

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(jimTools) 
library(data.table)

# list all vignettes
if (F){
vignette(package="data.table")
}

if (F) {
vignette("datatable-intro", package = "data.table")
vignette("datatable-reference-semantics", package = "data.table")
# this opens in new Chrome Tab!
# http://127.0.0.1:22091/library/data.table/doc/datatable-intro.html  
}
``` 

```{r dt_options}
# ALL options
options()
length(options()) #93
dt  <- as.data.table(options())


getOption("datatable.print.nrows") #100
```

### create
```{r begin}
# practice dt,  use function so easy to re-create
create_dt  <- function() {
  data.table(x = rep(c("b", "a", "c"), each = 3),
                  y = c(1, 3, 6),
                  v = 1:9)
}
dt  <- create_dt()
dt

```


### setDT
```{r list_exists}
dt1  <- setDT(x = list(1:9))
dt1
```

### DT[i, j, by]

###   R:                 i                 j        by
### SQL:  where | order by   select | update  group by
  * subset
  * order
  * list .()
  * return vector vs dt
```{r meaning}
# i:  subset, or order
dt[x == "a",]

dt[dt$x == "a",]  #also works

dt[x == "a" & y == 1,]
dt[order(y, -x),]   # y is asc, x in desc


# select columns
dt[, list(y,v)]  

# .(y) is alias for list(y)
dt[, list(y)] # as data.table
dt[, .(y)]  # SAME, short cut   
dt[, .(y,v)]

dt[, y]  # as vector!

#  select by column range  (and including)
dt[, x:y]

# by column number!
dt[, 1:2]

# drop first column
dt[, -(1)]


dt[, .(x, v)]  # return BOTH as list, ie new data.table


dt[, sum(v)]  # summarizes, vector
dt[, sum(v), by = x]  # summarizes, as table

dt[, .(sum = sum(v)), by = x]  # summarizes, with new variable


## RENAME
dt
dt[,.(z = v)]   # not permanent, only prints ONE col


## return ALL columns, rename just one
dt1 <-  dt[,.(x,y,z = v)]
dt1
```
### LOGIC
```{r logic}

dt[y <5,]

dt[, y < 5]  # vector
dt[, .(y < 5)]

dt[, .((y + v) > 10)]

```

### calc
```{r calc}
dt
dt[, .(x,y,v, new = y + v)]  # temp
dt

# no permanent change, display only new
dt[, .(new = y + v)]  # temp
dt
# 
# not SAME, grand total ?
dt[, .(new = sum(y,v))]
dt

# permanently create NEW col
dt[, newer := y +v]
dt
```



### aggregrate
```{r aggregate}

# 
dt
# for each in Col x, total, return dt
dt[, .(.N), by = .(x)]
dt

# SAME, shortcut
dt[, .N, by = x]
dt

# group by 2 columns, return dt
dt[, .N, by = .(x,y)]

```


### DEEP COPY
```{r copy}
# all dt
tables()

new  <- dt
old  <- data.table::copy(dt)
old
tables()

## Because modification (by reference) changes objects and 
##   anything points to it

# :=  forces by reference
new  <- dt[, new := 3*v]  

# both changed
new
dt

# no change
old

# now change back
dt  <- copy(old)
dt

```
### unique
```{r unique}
dt

# count
data.table::uniqueN(dt, by = "x")
uniqueN(dt, by = c("x", "y"))


# table, count for each unique rows
dt[, .N, by = .(x,y)]
dt[, .N, by = .(x)]

```
### subset, .SD

```{r subset}
dt[, .SD[1]]  # 1st row
dt[, .SD[1:4]]  # 1st 4 rows

```


.N, count returns vector
```{r N}
dt
dt[, .N]
dt[ x == "a", .N]

```


TODO:  all fail, select rows IF y+v > 6
```{r sum}

dt[, .(x,y,v)]
dt[(sum(y,v) > 6), .(x,y,v)]
dt[sum(y,v) > 6, ]
dt[, .SD[sum(y,v) > 6]  ]
```



TODO:  charcter to numeric?
FIx
```{r convert}


# x is char
dt[(sum(x,y) > 6)]
dt[with(dt, x + y) > 6, ]

```



***

```{r knit_exit()} 
knitr::knit_exit() 
```

\newpage

```{r render, eval=FALSE	} 
args(rmarkdown::render)
file  <- "/home/jim/code/try_things_here/rmd/00091_basic_data_table.Rmd"
rmarkdown::render(input= file,
                  output_format="html_document")


```
