---
title: "090_data_table.R"
output:  
  pdf_document:
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
fontsize: 12pt
---

```
REF: [http://rdatatable.github.io](rdatatable)
##		Vignette:  Intro
##		example(data.table)

# =======
# SETUP
# =======
```{r setup}

library(magrittr)
library(data.table)
citation("data.table")

input <- if (file.exists("./data/flights14.csv")) {
   "flights14.csv"
} else {
  "https://raw.githubusercontent.com/Rdatatable/data.table/master/vignettes/flights14.csv"
}
# 2.2 MM
flights <- fread(input)
flights
```
```{r lazy }

# I'm lazy!  use DT[]
DT <- flights
colnames(flights)
typeof(flights) #list
class(flights) # df AND "data.table"

dplyr::glimpse(flights)

# convert df
# DT <- as.data.table(df)

# =========================================
# BASIC:    WHERE, SELECT, GROUP BY (sql)
# =========================================
# WHERE clause( subset)  -------------------## 
flights [origin == "JFK" & carrier == "AA" & year == "2014" & month == 6L ] %>% nrow()
head( flights[order(origin)])


# TODO (doesn't get to head()) -------------------## 
 flights[order(origin)] %>% nrow() %>% head()
# group, total
DT[,.N,by=orgin]

# reverse order!
DT[order(-origin), .N, by=origin]

head( flights[order(-origin)] )

# return a vector -------------------## 
head (flights[, arr_delay])

# return a list -----------------##
head (flights[, list(arr_delay)])
head (flights[, .(arr_delay)])   # .( )  alias for list

# ========
# SELECT
# ========
flights[,mean(arr_delay)]
flights[,mean(arr_delay), by = dest]
flights[origin == "JFK" && 1:5,mean(arr_delay), by = dest]

# wrong:  why giving me first & last 5?
flights[origin == "JFK" && 1:5,mean(arr_delay), by = dest]

# why is orgin EWR appear?
flights[origin == "JFK" && 1:5 ,.(origin,mean(arr_delay)), by = dest]



# SELECT 3 coluns, 1 is calc -----------------##
# CHAIN, in desc order
flights[,.(arr_delay,dep_delay,totl = arr_delay + dep_delay)] [order(-totl)]

# CHAIN
flights[,.(arr_delay,dep_delay,totl = arr_delay + dep_delay)] [totl <0,]


# SELECT when sum is less 0
flights[sum(arr_delay+dep_delay) <0,] 

# WHERE (ie subset) and SELECT length of subset -----------------##
flights[origin == "JFK" & month==6L, length(dest)]

# shortcut .N   -----------------##
flights[origin == "JFK" & month==6L, .N] 


# GROUP
flights[,.N, by=dest]  # # rows by dest

# ==========
# FOR LOOP
# ==========

df <- data.frame(x=rnorm(100))
dt  <- as.data.table(df)

# each row i, set new col v = i
for(i in 1:100)dt[i, v := i]
head(dt)

Challenge Problem
# for each ID, count rows with non-zero values, build list of types (with
non-zero values)
my_tibble <- tidyr::tibble(
  ID = c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4,5,5,5,5),
  value = c(0,0,0,0,
            1,1,0,1,
            1,1,1,1,
            0,1,0,1,
            1,0,0,0),
  type = rep(c("a", "b", "c", "d"), 5)
)
my_tibble
t  <- my_tibble
```
