	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

PURPOSE:  practice data.table (dt)

SEE example(data.table)
SEE vignette(package=)

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```


```{r library, include=FALSE		}
library(jimTools) 
library(data.table)

vignette(package="data.table")
vignette("datatable-intro", package = "data.table")

vignette("datatable-reference-semantics", package = "data.table")
# this opens in new Chrome Tab!
# http://127.0.0.1:22091/library/data.table/doc/datatable-intro.html  
``` 


```{r begin}
dt  <- data.table(x = rep(c("b", "a", "c"), each = 3),
                  y = c(1, 3, 6),
                  v = 1:9)
dt
```

```{r list_exists}
dt1  <- setDT(x = list(1:9))
dt1
```

## DT[i, j, by]

##   R:                 i                 j        by
## SQL:  where | order by   select | update  group by
```{r meaning}
# i:  subset, or order
dt[x == "a",]
dt[x == "a" & y == 1,]
dt[order(y, -x),]   # y is asc, x in desc


# select columns
dt[, list(y,v)]  

# .(y) is alias for list(y)
dt[, list(y)] # as data.table
dt[, .(y)]  # SAME, short cut   
dt[, .(y,v)]

dt[, y]  # as vector!


dt[, .(x, v)]  # return BOTH as list, ie new data.table


dt[, sum(v)]  # summarizes, vector
dt[, sum(v), by = x]  # summarizes, as table

dt[, .(sum = sum(v)), by = x]  # summarizes, with new variable
```


## DEEP COPY

```{r copy}
old  <- data.table::copy(dt)
old

## Because modification (by reference) changes objects and 
##   anything points to it

# :=  forces by reference
new  <- dt[, new := 3*v]  

# both changed
new
dt

old

# now change back
dt  <- copy(old)
dt

```
## unique
```{r unique}
dt

# count
data.table::uniqueN(dt, by = "x")
uniqueN(dt, by = c("x", "y"))


# table, count for each unique rows
dt[, .N, by = .(x,y)]
dt[, .N, by = .(x)]

```
## subset, .SD

```{r subset}
dt[, .SD[1]]  # 1st row
dt[, .SD[1:4]]  # 1st 4 rows

```


.N, count returns vector
```{r N}
dt[, .N]
dt[ x == "a", .N]

```


TODO:  all fail, select rows IF y+v > 6
```{r sum}

dt[, .(x,y,v)]
dt[(sum(y,v) > 6), .(x,y,v)]
dt[sum(y,v) > 6, ]
dt[, .SD[sum(y,v) > 6]  ]
```



TODO:  charcter to numeric?
FIx
```{r convert}


# x is char
dt[(sum(x,y) > 6)]
dt[with(dt, x + y) > 6, ]

```



***

```{r knit_exit()} 
knitr::knit_exit() 
```

\newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
