
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

## REF:  S3 vignette from vctrs:: pkg

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(jimTools) 
library(vctrs)
``` 

constructor for vctrs_percent class
```{r begin}
new_percent  <- function(x = double()){
  vec_assert(x, double())
  new_vctr(x, class="vctrs_percent")
}
new_percent()

values  <- c(seq(0,1, length.out=4), NA)
typeof(values) #double

x  <- new_percent(values) # done!
str(x)
attributes(x)


```
helper, easier for user
```{r helper}
percent  <- function(x = double()){
  x  <- vec_cast(x, double())
  new_percent(x)
}

percent()
percent(1:5)  # converts int[]

x  <- percent( c(seq(0,1,length.out=4),NA) )
x
```

is_percent, to check
```{r is_percent }
is_percent  <- function(x) {
  inherits(x, "vctrs_percent")   # based on class of x 
}
is_percent(percent(1:5))  # TRUE  
```
round to 3 signif digits
```{r signif}
signif(.0021234, 3)
signif(.021234, 3)
signif(.21234, 3)
signif(2.1234, 3)
signif(21.234, 3)
signif(212.34, 3)
signif(2123.4, 3)
signif(21234., 3)
```

char
```{r formatC}

formatC(2.1234)
formatC(2.1234, format="E")
```

paste0 to add % suffix
```{r format}

format.vctrs_percent <- function(x, ...) {
  # formatC is style, round to 3 places
  out <- formatC(signif(vec_data(x) * 100, 3))
  out[is.na(x)] <- NA
  out[!is.na(x)] <- paste0(out[!is.na(x)], "%")
  out
}
format.vctrs_percent(rnorm(10))
# x (from above)
format.vctrs_percent(x)


```

***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
