	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  data.table,example,vignette,
---

PURPOSE:  simple, reference examples 
SEE example(data.table) (ie in code)

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(jimTools) 
library(data.table)

# list all vignettes
if (F){
vignette(package="data.table")
}

if (F) {
vignette("datatable-intro", package = "data.table")
vignette("datatable-reference-semantics", package = "data.table")
# this opens in new Chrome Tab!
# http://127.0.0.1:22091/library/data.table/doc/datatable-intro.html  
}
``` 

```{r dt_options}
# ALL options
if (F) options()
length(options()) #93

# not so simple,  options is unruly list
if (F) dt  <- as.data.table(options())
getOption("datatable.print.nrows") #100
```

### create dt
```{r begin}
# practice dt,  use function so easy to re-create
create_dt  <- function() {
  data.table(x = rep(c("b", "a", "c"), each = 3),
                  y = c(1, 3, 6),
                  v = 1:9)
}
dt  <- create_dt()
dt
```


### setDT
```{r list_exists}
dt1  <- setDT(x = list(1:9))
dt1

# last element
dt1[, .N] # as dt
dt1[.N] # as vector
dt1[1]
dt1[2]

dt1[-2] # omit 2nd elemnt
dt1[!2] # SAME
```

### DT[i, j, by]

###   R:                 i                 j        by
### SQL:  where | order by   select | update  group by
  * subset
  * order
  * list .()
  * return vector vs dt
```{r meaning}
# i:  subset, or order
dt[x == "a",]

dt[dt$x == "a",]  #also works

dt[x == "a" & y == 1,]
dt[order(y, -x),]   # y is asc, x in desc


# select columns
dt[, list(y,v)]  

# .(y) is alias for list(y)
dt[, list(y)] # as data.table
dt[, .(y)]  # SAME, short cut   
dt[, .(y,v)]

dt[, y]  # as vector!

#  select by column range  (and including)
dt[, x:y]

# by column number!
dt[, 1:2]

# drop first column
dt[, -(1)]


dt[, .(x, v)]  # return BOTH as list, ie new data.table


dt[, sum(v)]  # summarizes, vector
dt[, sum(v), by = x]  # summarizes, as table

dt[, .(sum = sum(v)), by = x]  # summarizes, with new variable


### RENAME
dt
dt[,.(z = v)]   # not permanent, only prints ONE col


## return ALL columns, rename just one
dt1 <-  dt[,.(x,y,z = v)]
dt1
```
### LOGIC
```{r logic}

dt[y <5,]

dt[, y < 5]  # vector
dt[, .(y < 5)]

dt[, .((y + v) > 10)]

```

### calc
```{r calc}
dt
dt[, .(x,y,v, new = y + v)]  # temp
dt

# no permanent change, display only new
dt[, .(new = y + v)]  # temp
dt
# 
# not SAME, grand total ?
dt[, .(new = sum(y,v))]
dt

# permanently create NEW col
dt[, newer := y +v]
dt
```



### aggregrate
```{r aggregate}

# 
dt
# for each in Col x, total, return dt
dt[, .(.N), by = .(x)]
dt

# SAME, shortcut
dt[, .N, by = x]
dt

# group by 2 columns, return dt
dt[, .N, by = .(x,y)]

```


### DEEP COPY
```{r copy}
# all dt
tables()

new  <- dt
old  <- data.table::copy(dt)
old
tables()

## Because modification (by reference) changes objects and 
##   anything points to it

# :=  forces by reference
new  <- dt[, new := 3*v]  

# both changed
new
dt

# no change
old

# now change back
dt  <- copy(old)
dt

```
### unique
```{r unique}
dt

# count
data.table::uniqueN(dt, by = "x")
uniqueN(dt, by = c("x", "y"))


# table, count for each unique rows
dt[, .N, by = .(x,y)]
dt[, .N, by = .(x)]

```
### subset, .SD

```{r subset}
dt[, .SD[1]]  # 1st row
dt[, .SD[1:4]]  # 1st 4 rows

```


.N, count returns vector
```{r N}
dt
dt[, .N]
dt[ x == "a", .N]

```


TODO:  all fail, select rows IF y+v > 6
```{r sum}

dt[, .(x,y,v)]
dt[(sum(y,v) > 6), .(x,y,v)]
dt[sum(y,v) > 6, ]
dt[, .SD[sum(y,v) > 6]  ]
```



TODO:  charcter to numeric?
FIx
```{r convert}


# x is char
dt[(sum(x,y) > 6)]
dt[with(dt, x + y) > 6, ]

```

### Study examples
```{r begin2}
create_dt  <- function()
  DT = data.table(x=rep(c("b","a","c"),each=3), y=c(1,3,6), v=1:9)

dt  <- create_dt()
dt

example(data.table)

```

### order(x)
```{r order}
dt[order(x),]
dt[order(y),]
dt[order(y,v),]
dt[order(y,-v),]
dt[y == 6,][order(v),]

# wrong
dt[y == 6 & order(-v), ]
```

```{r challenge}

# Challenge Problem
# for each ID, count rows with non-zero values, 
  build list of types (with non-zero values)

my_tibble <- tidyr::tibble(
  ID = c(1,1,1,1,2,2,2,2,3,3,3,3,4,4,4,4,5,5,5,5),
  value = c(0,0,0,0,
            1,1,0,1,
            1,1,1,1,
            0,1,0,1,
            1,0,0,0),
  type = rep(c("a", "b", "c", "d"), 5)
)
my_tibble
dt  <- as.data.table(my_tibble) 

##  Each ID, show non-zero count
dt[value != 0, .N, by = ID]


##  Each ID, list the types with non-zero values
dt[value != 0, .(unique(type)), by = ID]

## TODO:
## SAME, but use list-column for the types
dt[value != 0, c(ID, types `:=`  .(type)), by = ID]
dt[]


```

### simple: reshape, list-column 

```{r simple1}
# practice dt,  use function so easy to re-create
create_dt  <- function() {
  data.table(x = rep(c("b", "a", "c"), each = 3),
                  y = c(1, 3, 6),
                  v = 1:9)
}
dt  <- create_dt()
dt
```

### group and head()
```{r}
# 3lines for each grouping
dt[,.SD, by=x]

# here limit to 2 lines, for each grouping
dt[,head(.SD, 2), by =x]

# first x=b, then x=a ...
dt[,print(.SD), by =x]

# for each grouping, show val
dt[, .(val = y), by =x]

```{r list_column}
# THINK
dt[, .(val = c(y,v)), by=x]

# THINK
dt[, .(val = list(y,v)), by=x]
dt[, .(val = list(c(y,v))), by=x]


```

```{r select}
dt[,c("x","y","v")]

# SAME, easier!
dt[, .(x,y,v)]

# shortcut?
col  <- c("x","y","v")
dt[, ..col]

# FAIL
if (F) {
  dt[, col]
  dt[, .col]
}

dt[, .(col)]
```
### Explain !
  *  in dt, work with NULL, NA
```{r explain}


dt3  <- function() {
  data.table( ID = c(1,2,3,4,5,6),
      tags= list(
                 "a,b,c",
                 NULL,    # blank
                 NA,      # NA
                 x,       # NA
                 "x",
                 "y,z,NULL,NA")
      )
}
dt  <- dt3()
dt


# Step 1 (remove blank)

f  <- function(e) ifelse(is.null(e), return(NA), return(e))
dt[,.(ID, tags = lapply(tags, f)) ]

dt[is.na(dt$tags)]
dt[is.na(tags)]



# Explainh errors ....
if (F) {
dt[is.null(tags)]
dt[tags == NULL]
dt[tags == NA]
dt[tags == c(NA)]
}


# Explain behavior!
g  <- function(e) return(e)

dt[, .(ID, tags = lapply(tags, g))] 

# 2 ID columns
dt[, .(ID, tags = lapply(tags, g)), by=ID] 

dt[, .(ID, tags = lapply(.SD, g)), by=ID] 

if(F) dt[, tags = lapply(.SD, g), by=ID] 





```
***

```{r knit_exit()} 
knitr::knit_exit() 
```

\newpage

```{r render, eval=FALSE	} 
args(rmarkdown::render)
file  <- "/home/jim/code/try_things_here/rmd/00291_DT_STUDY_examples.Rmd"
file  <- basename(file)
file
rmarkdown::render(input= file,
                  output_format="html_document",
                  output_file=here("rmd","out")
                  )




```
