	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  data.table,example,vignette,
---


TODO:   merge this into DT_reshape.Rmd
PURPOSE:  simple, reference examples :
    * nest, unnest, normalize, list-column etc.

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(jimTools) 
library(data.table)
library(knitr)

```

### create dt

```{r begin}
# practice dt,  use function so easy to re-create
create_dt  <- function() {
  data.table(ID = rep(c("b", "a", "c"), each = 3),
                  x = c(1, 3, 6),
                  y = 1:9)
}
dt  <- create_dt()
dt
```

### Normalized!
```{r normalize}
dt[,.SD, by=ID]

# `concatinate two chr[] for each ID`
dt[, .(c(x,y)), by = ID]
dt
```
```{r}
dt  <- create_dt()
dt[,.(list(x,y))      , by=ID]


# max packed
dt[,.(list(c(x,y))), by = id]

# $v1 is list of 3 num[]
str(dt[,.(list(c(x,y))), by = id]$v1)

```



### simple: reshape, list-column 

```{r simple1}
# practice dt,  use function so easy to re-create
create_dt  <- function() {
  data.table(ID = rep(c("b", "a", "c"), each = 3),
                  x = c(1, 3, 6),
                  y = 1:9)
}
dt  <- create_dt()
dt
```

### group and head()
```{r}
# 3lines for each grouping
dt[,.SD, by=ID]


# for each ID, choose max y
dt[, .SD[which.max(y)], by = ID]

# here limit to 2 lines (not 3), for each grouping
dt[,head(.SD, 2), by =ID]

# break out by ID
dt[,print(.SD), by =ID]

# for each grouping, show val
dt[, .(val = y), by =ID]

```{r list_column}
# THINK
dt[, .(val = c(x,y)), by=ID]

# THINK
dt[, .(val = list(x,y)), by=ID]
dt[, .(val = list(c(x,y))), by=ID]

```


### Group on ID, display list-column of data.tables!
  *   make list of each SD,  then list all the results
```{r data.table}

# 3 the SAME
dt[, list(dt.xy=list(.SD)), by=ID]
dt[, .(dt.xy=list(.SD)), by=ID]
dt[, .(dt.xy=.(.SD)), by=ID]


# FAILS b/c do not have every subset
if (F) dt[, .(dt.xy=.SD), by=ID]
```

### undo/unwind list-column
```{r un}

dt2  <- dt[, .(dt.xy=.(.SD)), by=ID]

# each ID, pop off 1st entry (there is only 1)
dt3  <- dt2[, dt.xy[[1]], by = ID]
identical(dt,dt3)
```




```{r select}
dt[,c("ID","x","y")]

# SAME, easier!
dt[, .(ID,x,y)]

# shortcut?
col  <- c("ID","x","y")
dt[, ..col]

# FAIL
if (F) {
  dt[, col]
  dt[, .col]
}

dt[, .(col)]
```

***

```{r knit_exit()} 
knitr::knit_exit() 
```

\newpage

```{r render, eval=FALSE	} 
args(rmarkdown::render)
file  <- "/home/jim/code/try_things_here/rmd/00291_DT_STUDY_examples.Rmd"
file  <- basename(file)
file
rmarkdown::render(input= file,
                  output_format="html_document",
                  output_file=here("rmd","out")
                  )




```
