---
title:  "0399A_yaml_utils.R"
date:   "`r format(Sys.Date(), '%d_%b_%Y)`" 
TAGS:  yaml,tags,lapply examples,utils,vignette
---

  ### Utilies for data.table
  *  which route?    dplyr, tidyr (pivots, wider, longer df) ?
  *  TODO This grab bag needs Cleanup

#'  NOTATION:   Following Hadley (purrr:::) refer to `list` as list and refer
  #'  to `atomic vectors` as character vectors, numeric vectors etc ....
  #'  FOR BREVITY:    chr[] refers to character vector,  chr[n] refers to
  #'  character vector where length of each element is n.
  #'
  #'
  #'


### Given list of this form list(file_name1 = c("a", "b", "c"))  create dt with 2 columns (file_name and tag) , normalized


### list all vignettes
```{r all_vignettes}

if (F){
vignette(package="data.table")
}
```
### Activate vignette (opens http://127 ....server)
```{r show_vignette}

if (F) {
vignette("datatable-intro", package = "data.table")
vignette("datatable-reference-semantics", package = "data.table")
# this opens in new Chrome Tab!
# http://127.0.0.1:22091/library/data.table/doc/datatable-intro.html  
}
``` 
### DT and options
```{r dt_options}
# ALL options
head(options())
length(options()) #93

getOption("datatable.print.nrows") #100
```
