
---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.5in,top=0.25in   
TAGS:  data.table,Lahman,normalize,nest,unest
---

Normalize and de-Normalize Data
REF:  https://rdatatable.gitlab.io/data.table/library/data.table/doc/datatable-reshape.html
  * Follow the tutorial
  * Conent here is LIMITED annotation, follow the tuturial

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read
library(jimTools)
library(data.table)
```


### warm-up, create list-column
REF:  
```{r create_list_col}
# l is list of 3, each also list

create_dt   <- function(){
  dt  <- data.table(
                  x = letters[1:3],
                  l = list(list(c1=6L,
                                c2=4L),
                           list(x=2L,
                                y=4L,
                                z=3L),
                           list()
                           )
                  )
}
dt  <- create_dt()
dt
```

### before unwind (unnest) - understand what we have
```{r understand}
y  <- dt$l

# y has 3 elments, unnamed, no suprise
# each elment holds another list, but these are named
y

length(y)

# examine 1st of y's 3 elments, a named list of 2 ... any suprise?
y[[1]]


## unlist(dt)
unlist(dt)


## unlist(y)
unlist(y)

# crude attempt to see .SD for each value of column x
dt[,print(.SD), by=x]
```

### unwind (unnest) the list-column
*  result:  normalized
*  effectively, keyed on x, and new column `name`
*   2nd new column will be `value`
*   no need to modify-in-place
*   magic is done by unlist()

```{r unwind}
dt

# helps to remove list[0]
dt  <- create_dt()
dt[]

# not working, change 1 elment in column
dt  <- dt[length(l) == 0, l := NA]
dt[]


# now unwind - 2 steps
# by default, unlist  keeps names use.names=T
dt[,.(name = names(unlist(l))),    by=x]
dt[,.(value = unlist(l)), by =x]

# unwind - both in 1 step
dt[,.(name = names(unlist(l)), value = unlist(l)),    by=x]

```








### Section 1
### non-normalized data
```{r}
s1 <- "family_id age_mother dob_child1 dob_child2 dob_child3
1         30 1998-11-26 2000-01-29         NA
2         27 1996-06-22         NA         NA
3         26 2002-07-11 2004-04-05 2007-09-02
4         32 2004-10-10 2009-08-27 2012-07-21
5         29 2000-12-05 2005-02-28         NA"
DT <- fread(s1)
DT
str(DT)
```


```{r knit_exit(), include = FALSE  , eval=FALSE } 
# knitr::knit_exit()
```

\newpage


```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

## In doubt?   USE  knitr and do not waste time!

## Want to embed Latex, stick to pdf output (html?   never sure!)

# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# in general, pdf will look nicer

file  <-  "/home/jim/code/try_things_here/rmd/0088_base_package_commands.Rmd"

rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "out")
```
