---
title:  "0399A_yaml_utils.R"
date:   "`r format(Sys.Date(), '%d_%b_%Y)`" 
TAGS:  yaml,tags,lapply examples,utils,vignette
---

  ### Utilies for data.table
  *  which route?    dplyr, tidyr (pivots, wider, longer df) ?
  *  TODO This grab bag needs Cleanup

#'  NOTATION:   Following Hadley (purrr:::) refer to `list` as list and refer
  #'  to `atomic vectors` as character vectors, numeric vectors etc ....
  #'  FOR BREVITY:    chr[] refers to character vector,  chr[n] refers to
  #'  character vector where length of each element is n.
  #'
  #'
  #'


### Given list of this form list(file_name1 = c("a", "b", "c"))  create dt with 2 columns (file_name and tag) , normalized

### normalize
```{r normalize}

# ==================
# clever, no cigar
# ==================

library(data.table)
l  <- list(file1 = c("a", "b", "c"))

# check that l is singleton
file_name  <- names(l)
file_name

tags = l[[1]] 
l[[file_name]]
tags
identical(tags, l[[file_name]])



normalize  <- function(file_name = character(), tags = character()) {
  n  <- length(tags)  # check > 0
  dt  <- data.table(file_name = rep(file_name,n),
                    tags = tags)
}

dt  <- normalize(file_name = file_name, tags = tags)
dt

dt1  <- normalize(file_name = names(l), 
                  tags = l[[file_name]])
dt1
identical(dt, dt1)
```


### TODO  -- this is all wrong 
### Suppose list of many
```{r normalize_many}

l  <- lines
l  <- list(file1 = c("x", "y"),
           file2 = c("A", "B"),
           file3 = c("jim", "joe", "sally"))
l



l$file3[[1]]
l[[c(3,2)]]

lapply(l, normalize, tags = function(e) e[[file_name]])
dtB  <- lapply(l, normalize, tags = function(e) unlist(e))
dtB
lapply(l, normalize, tags = function(e) e$file_name)


       file_name = function(e) names(e),
       tags = function(e) e[[file_name]]) 

lapply(l, function(e) e) 
```


<!--
# ========
#   KEEP  - useful related to vignettes
# ========
-->

### list all vignettes
```{r all_vignettes}

if (F){
vignette(package="data.table")
}
```
### Activate vignette (opens http://127 ....server)
```{r show_vignette}

if (F) {
vignette("datatable-intro", package = "data.table")
vignette("datatable-reference-semantics", package = "data.table")
# this opens in new Chrome Tab!
# http://127.0.0.1:22091/library/data.table/doc/datatable-intro.html  
}
``` 
### DT and options
```{r dt_options}
# ALL options
options()
length(options()) #93

# FAILS
dt  <- as.data.table(options())

getOption("datatable.print.nrows") #100
```
