<!--
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
   -->
--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw

vim: to format all urls for md
s/https.*/[&]()/g
-->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools)
``` 
Chapter 14  
https://bookdown.org/yihui/rmarkdown-cookbook/reuse-chunks.html
See also chunk reference labels {r name, label.ref= c('labela', 'labelb')}
in same Chapter.


Nested Chunks:  
  - top level chunk: f2c  

Calls:  
  - chunk called:  check-arg  
  - chunk called:  convert  

```{r, f2c}
F2C <- function(x) {
  <<check-arg>>
  <<convert>>
}

F2C(30)
```

First, we check if the input value is numeric:

```{r, check-arg, eval=FALSE}
  if (!is.numeric(x)) stop("The input must be numeric!")
```

Then we do the actual conversion:

```{r, convert, eval=FALSE}
  (x - 32) * 5/ 9
```

```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file <- "00301_nested_chunks_literate_programming.Rmd"
file  <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
