<!--
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
   -->
--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  create_Rmd, knit_expand 
---


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
library(jimTools)
library(knitr)
```

### PURPOSE:   From this (main) file, call a second(child) file.
  *  Goal:   modify the label of child file dyamically.


`r prefix = "JIM-"; knitr::knit(text = knitr::knit_expand(file= here("rmd/00302_child.Rmd")))`


```{r create_Rmd}
y  <- 1 
library(knitr)
writeLines(c("```{r}", "y + 1", "```"), "test.Rmd")
knit("test.Rmd", quiet = TRUE, envir = new.env())   # will find y in parent
knit("test.Rmd", quiet = TRUE, envir = baseenv())   # no y
# [1] "test.md"
cat(readLines("test.md"), sep = "\n")

# clean up
file.remove("test.Rmd")
file.remove("test.md")
```

\newpage
# 
```{r render, eval=FALSE, echo=F, incude=F	} 
file  <- "rmd/00302_knitr_modify_label_in_template.Rmd"
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "out")
```
