
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

---
othercomments
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw
-->

<!--  commments -->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools) 
``` 

match.call() 
* use inside a function.
* returns list.
* [1] is name of function.
* [2:n] are UNevaluated args.  (ie symbols are ok)
```{r match.call}
g  <- function(x, y, z){
  return(as.list(match.call()))
}

# Examples
g(1,2,3)
g(a,b,c)

# Study
l  <- g(a,b,c)
l
str(l)

# Example
l  <- g(1+3, b, c+0)
l
str(l)


match.call , in real life
```{r begin}
f <- function(x, y, z) {
  do.call("sum", as.list(match.call()[-1])) 
}

f(1, 2, 3)
```

list
```{r list}
l  <- list(a=1, b=2)
l
l[-1]   # b only
l[-2]   # a only
l[c(-1)]  # b only

# hmmm?
l[c(-1,-2)]
l[-2:-1]
l[-3:-1]
l[0]

```
random:
How to use unicode!
```{r random}
letters <- c(1, 2, 3, 4)
names(letters) <- c("a", "b", "c", "\U1F409")
letters
```
match.arg, limit possible input values
```{r match.arg}
f  <- function(type=c("a", "b"))
  type  <- match.arg(type)

(f("b"))  # chr[]   "b"
(f("c")) #Error, not "a" or "b"
```

```{r UseMethod}
f  <- function(x) {
  UseMethod(x) 
}

f.character <- function(x) print("character")
f.numeric  <- function(x) print("numeric")

f("a")
f(2)
```





***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
