---
title: "064_knitr:: chunk practice"
output:
  html_document: 
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    number_sections: TRUE
geometry: margin=1in
editor_options: 
  chunk_output_type: console
---

echo[TRUE] print code  
eval[TRUE] eval cod  
include[TRUE] print output  

# 0-setup, no output
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment="      ##", z=TRUE, collapse=TRUE)
library(knitr)
library(jimTools)
```

## 1-evaluate, but include no output
```{r test, eval=TRUE, include=FALSE}
print("hi")
x=4
show_text=TRUE
```

### 2-output, but do not evaluate
```{r check, eval=FALSE}
print(x)
```

3-nothing!
```{r check1, eval=FALSE, include=FALSE}
print(x+2)
```

3A-results='hide', code only
```{r results='hide'}
print(x+2)
print("3A-results=hide")
```

3B-NOT results='hide'
```{r collapse=TRUE}
print(x+2)
print("3B-results=NOT hide")
```

4A-conditional
```{r conditional_block, echo=show_text, eval=show_text, include=show_text }

# see in 1-setup
print("show_text is TRUE")
```

4B-conditional
```{r conditional_blockB, echo=!show_text, eval=!show_text, include=!show_text }

# see in 1-setup
print("show_text is TRUE")
```

7A-collapse=TRUE
```{r collapse=TRUE }
print("one")
print("two")
```

7B-collapse=FALSE
```{r collapse=FALSE}
print("one")
print("two")
```


20-list extra-all chunk options
```{r chunk_options, include=show_text, collapse=TRUE}
str(knitr::opts_chunk$get())
```



21-knitr::read_chunk from external .R file WORKS!
[REF:] (https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html)
```{r read, cache=FALSE }
  knitr::read_chunk('~/code/r_try_things_here/rmd/file_for_read_chunk.R')

```

22-test_a
```{r, test_a}

```


STop
```{r}
knitr::knit_exit()
```

render from code, to html
```{r eval=FALSE}
rmarkdown::render(
	"~/code/r_try_things_here/064_knitr_chunk_practice.html", 
	rmarkdown::html_document(), output_dir="./html") 
```

render from code, to pdf
```{r eval=FALSE}
rmarkdown::render(
	"~/code/r_try_things_here/064_knitr_chunk_practice.html", 
	rmarkdown::pdf_document()) 
jimTools::ren_pdf("064_knitr_chunk_practice.Rmd", "rmd")
```
