	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

## base:: files and dir commands
```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(jimTools) 
```

base:: commands (why use fs:: ?)
```{r begin}
list.files("./rmd", pattern="*.Rmd")
list.files("./rmd", full.names = TRUE ,pattern="*.Rmd")

if (FALSE)
  # file.rename()
  # file.move()
  # file.create()
  # basename()
  # dirname()
  #
  # list.dirs()
  # dir()
```

sprintf has some nice features!
```{r sprint}
sprintf("hello %s", "jim")

sprintf("hello %s", 23)

sprintf("hello %04s", 23)         # min of 4
sprintf("hello %04f", 23)         # 23.000000

sprintf("hello %04i", 23)         # int, min of 4 digits

```

regex, sub(), gsub will work
stringr is easier?
```{r}
the_dir  <- "./rmd"
the_files  <- list.files(the_dir)
the_files
library(stringr)
# regex
the_prefixes  <- stringr::str_extract(the_files, "^[0-9]*[A-Z]?_" )
the_prefixes

```


***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
