
---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.5in,top=0.25in   
---


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read
```


```{r library, include=FALSE		}
file  <- knitr::current_input() 
library(jimTools)
``` 

## Where are we now?
  * which dir are used?
  * status of pkgs in each dir
```{r now}
x  <- packageStatus()
print(x)
{
begin  <- Sys.time()

if (FALSE ) upgrade (x, ask = FALSE , checkBuild=TRUE)
upgrade (x, ask = FALSE , checkBuild=TRUE)

end  <- Sys.time()
diff  <- end - begin
diff
}
# print updated summary table
update(x, checkBuild=TRUE)

```

I installed R 4.0.4 from source on Raspberry Pi 400, which uses ARM processessors.  This seemed fine.

However, when I tried to upgrade installed packages I ran into a few problems.

My question is:  After installing R 4.0.4 but before upgrading packages, what
steps should I take to insure the package upgrade will be successful?

This is what I did do; which did not succeed:

First, I ran packageStatus to see ~29 packages needed an upgrade.
>> x  <- packageStatus()
>> print(x)
>>

Then a did the upgrade which reported ~ 31 warnings (packages that had non-zero exit status).  About 1 or 2 packages did upgrade successfully.


>> upgrade(x, ask=FALSE)
>> update(x)
>>
However, I also noted errors during the upgrade such as:
*  Wrong ELF class ELFCLASS64.
*  lazy loading failed for package:   "DT"





## Examine available.packages() (>17k)

```{r begin}
versions  <- available.packages()
t  <- as_tibble(versions)
t[,c(1,2,12:13,15:16)]
str(versions)
names(versions)
typeof(versions)
length(versions)
versions[[4]]
head(versions)
```


## Examine installed.packages
```{r installed.packages}
installed  <- as_tibble(installed.packages())
head(installed) 
str(installed)

installed[,"NeedsCompilation" == "no"]
```

## install.packages (using DT)
```{r installed_use_DT}
library(data.table)
dt  <- as.data.table(installed)
names(dt)

# remove uneeded fields
# .()  is alias for list()
dt  <- dt[, .(Package,  OS_type, NeedsCompilation, Built)]


# count, for each grouping
dt[,.(.N)    , by=.(Built)]

dt[, .N, by = .(Built)]   

dt[, .N, by = "OS_type"]  # all <NA>

dt[, .N, by = .(NeedsCompilation)]   # 141 need!

# as subset
x  <- dt[NeedsCompilation == "yes"]
x[1:100]



```


## All packages, by library location
```{r library}
library()

(.packages())
(.packages(all.available  = TRUE))
```

## Default Packages
```{r default}
getOption("defaultPackages")
```

## Basic installation
  *  sudo R CMD INSTALL *.tar.gz
```{r install}
  install.packages("igraph", verbose=TRUE, dependencies=TRUE)
```


## workhorse
```{r update}
{
begin  <- Sys.time()
.libPaths()
update.packages(ask=FALSE )
end  <- Sys.time()
diff  <- end - begin
diff
}
```

## Troubleshoot
```{r trouble}
{
begin  <- Sys.time()

# p  <- c("digest")
# p  <- c("future")
# p  <- c("httpuv")
# p  <- c("jsonlite")
# p  <- c("servr")
## p  <- c("lattice")
#p  <- c("utf8")
p  <- c("rlang")

install.packages(p, verbose=TRUE, dependencies = TRUE)
# upgrade (p, ask = FALSE , checkBuild=TRUE)

end  <- Sys.time()
diff  <- end - begin
diff
}
# print updated summary table

```


***

```{r knit_exit(), include = FALSE  , eval=FALSE } 
# knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

file  <- here("", "")
file 
# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
[``
