---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.5in,top=0.25in   
TAGS:  RPI, rpi,Rscripts
---

# =============================================
## CODE related to package mgmt SEE  jimTools
# =============================================


PURPOSE:   Tools to manage R packages.
TODO:  Study code:   
https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them

SEE ALSO:  ~/code/zsh_scripts_project/071_run_R_to_update_packages.sh


### Loads,  but does not attach; TRUE if able to load.
```{r}
check_suggests <- function(package) {
  if (!requireNamespace(package, quietly = TRUE)) {
    #stop("Suggested package '{package}' not present.")
    stop("Not found")
  }
}
check_suggests("tibble")
check_suggests("joke")
check_suggests("glue")
check_suggests("ggplot2")
```

*  May 16, 2021 - best code is in jimTools:: Rscripts
### =========================
###   UPDATE  - 2 step process?
###   jimTools - has newer version of this
### =========================

## To update `old` packages
*   Appears that need to do this first, before update others
*   To increase `old` to include those comipiled earlier minor version of R
*   Use checkBuilt= T to force

```{r update_old}
update.packages(ask = F,
                oldPkgs = old.packages(),
                checkBuilt=F)
```

## To upgrade

vs update.packages() ?
```{r upgrade}

{
begin  <- Sys.time()
    update.packages(ask = FALSE , checkBuild=TRUE)
    end  <- Sys.time()
    diff  <- end - begin
    diff
}

```



## NOTES:  re:  install on Rpi
I installed R 4.0.4 from source on Raspberry Pi 400, which uses ARM processessors.  This seemed fine.

However, when I tried to upgrade installed packages I ran into a few problems.

My question is:  After installing R 4.0.4 but before upgrading packages, what
steps should I take to insure the package upgrade will be successful?

This is what I did do; which did not succeed:

First, I ran packageStatus to see ~29 packages needed an upgrade.
>> x  <- packageStatus()
>> print(x)
>>

Then a did the upgrade which reported ~ 31 warnings (packages that had non-zero exit status).  About 1 or 2 packages did upgrade successfully.


>> upgrade(x, ask=FALSE)
>> update(x)
>>
However, I also noted errors during the upgrade such as:
*  Wrong ELF class ELFCLASS64.
*  lazy loading failed for package:   "DT"








```{r knit_exit(), include = FALSE  , eval=FALSE } 
knitr::all_labels()
# knitr::knit_exit()
```

\newpage

```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# in general, pdf will look nicer

file  <-  "/home/jim/code/try_things_here/rmd/0088_base_package_commands.Rmd"

rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "./out")
  #                output_file = "~/Downloads/print_and_delete/out")
```
