
---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.5in,top=0.25in   
TAGS:  packageDescription(),dependencies,remove.packages(),old.packages()
---




PURPOSE:   Tools to manage R packages.
TODO:  Study code:   
https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them

SEE ALSO:  ~/code/zsh_scripts_project/071_run_R_to_update_packages.sh


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read
file  <- knitr::current_input() 
library(tibble)
library(here)
```

```{r knit_exit2} 
knitr::all_labels()
knitr::knit_exit()
```

## .libPaths() to find user and `common` library locations
```{r libPaths}
# in my home directory, and
# in /usr/local/lib/R/library
.libPaths()
```

## list packages in user library
```{r list_user }
# top level dir only (so not recursive)
pkgs  <- lapply(.libPaths(), list.dirs, recursive=FALSE)

# user - 288!
length(pkgs[[1]])

# common - 30  
length(pkgs[[2]])
```

## What repos are packages from?
```{r repos}
options("repos")
```


### remove a package
```{r remove_package}
  if (F) {
    pkg = 
  remove.packages(pkg)
  }

```


## Examine installed.packages
*  See below, use DT to ask Q


```{r installed.packages}
installed  <- as_tibble(installed.packages())

# 6 x 16
head(installed) 
str(installed)

# 317
installed[,"NeedsCompilation" == "no"]

installed[,"NeedsCompilation" == "yes"]

```
## Examine available.packages() (>17k)

```{r begin}
versions  <- available.packages()

t  <- as_tibble(versions)
t[,c(1,2,12:13,15:16)]
str(versions)
names(versions)
typeof(versions)
length(versions)
versions[[4]]
head(versions)
```



## Where are we now?
  * which dir are used? (for user pkgs, for common pkgs)
  * status of pkgs in each dir
```{r now}
# packages that can be upgraded
old.packages()

# x is list
x  <- packageStatus()
typeof(x)
str(x)
print(x)

# details
summary(x)

```

### =========================
###   UPDATE  - 2 step process?
###   jimTools - has newer version of this
### =========================

## To update `old` packages
*   Appears that need to do this first, before update others
*   To increase `old` to include those comipiled earlier minor version of R
*   Use checkBuilt= T to force

```{r update_old}
update.packages(ask = F,
                oldPkgs = old.packages(),
                checkBuilt=F)
```

## To upgrade

vs update.packages() ?
```{r upgrade}

{
begin  <- Sys.time()
    update.packages(ask = FALSE , checkBuild=TRUE)
    end  <- Sys.time()
    diff  <- end - begin
    diff
}

```



## NOTES:  re:  install on Rpi
I installed R 4.0.4 from source on Raspberry Pi 400, which uses ARM processessors.  This seemed fine.

However, when I tried to upgrade installed packages I ran into a few problems.

My question is:  After installing R 4.0.4 but before upgrading packages, what
steps should I take to insure the package upgrade will be successful?

This is what I did do; which did not succeed:

First, I ran packageStatus to see ~29 packages needed an upgrade.
>> x  <- packageStatus()
>> print(x)
>>

Then a did the upgrade which reported ~ 31 warnings (packages that had non-zero exit status).  About 1 or 2 packages did upgrade successfully.


>> upgrade(x, ask=FALSE)
>> update(x)
>>
However, I also noted errors during the upgrade such as:
*  Wrong ELF class ELFCLASS64.
*  lazy loading failed for package:   "DT"






## install.packages (using DT)

```{r installed_use_DT}
library(data.table)
library(microbenchmark)

# 4.5 ms
microbenchmark( installed  <- tibble::as_tibble(installed.packages()))

dt  <- as.data.table(installed)
names(dt)

# remove uneeded fields
# .()  is alias for list()
dt  <- dt[, .(Package,  OS_type, NeedsCompilation, Built)]
dt

# count, group by version of Build
dt[,.(.N)    , by=.(Built)]

# same
dt[, .N, by = .(Built)]   

dt[, .N, by = "OS_type"]  # all <NA>

dt[, .N, by = .(NeedsCompilation)]   # 141 need!

# create subset, needs compilation
x  <- dt[NeedsCompilation == "yes"]
x[1:100]

```



## More advanced package managment:

## List all packages + description, by library location
```{r library}
library()

(.packages())
(.packages(all.available  = TRUE))
```

## List Default Packages ?? Explain
```{r default}
getOption("defaultPackages")
```

## Basic installation
  *  sudo R CMD INSTALL *.tar.gz
```{r install}
  install.packages("igraph", verbose=TRUE, dependencies=TRUE)
```


## workhorse
```{r update}
{
begin  <- Sys.time()
.libPaths()
update.packages(ask=FALSE )
end  <- Sys.time()
diff  <- end - begin
diff
}
```

## Troubleshoot:  install packages one at a time
```{r trouble}
{
begin  <- Sys.time()

# p  <- c("digest")
# p  <- c("future")
# p  <- c("httpuv")
# p  <- c("jsonlite")
# p  <- c("servr")
## p  <- c("lattice")
#p  <- c("utf8")
# p  <- c("rlang")

install.packages(p, verbose=TRUE, dependencies = TRUE)
# upgrade (p, ask = FALSE , checkBuild=TRUE)

end  <- Sys.time()
diff  <- end - begin
diff
}
# print updated summary table

```

### Pkg Dependencies
```{r dependencies}
pkg = "purrr"
packageDescription(pkg)
# subset vector
packageDescription(pkg)[c("Imports", "Suggests")]
```


## sessioninfo::package_info()
```{r package_info}
# For Installed Packages, yields dependencies
sessioninfo::package_info(pkg="tibble")
```

***

```{r knit_exit(), include = FALSE  , eval=FALSE } 
knitr::all_labels()
# knitr::knit_exit()
```

\newpage

```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# in general, pdf will look nicer

file  <-  "/home/jim/code/try_things_here/rmd/0088_base_package_commands.Rmd"

rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "./out")
  #                output_file = "~/Downloads/print_and_delete/out")
```
