---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
    fontsize: 11pt   
    geometry: margin=0.5in,top=0.25in   
TAGS: lapply,sapply,named,unamed,simplification
---

## PURPOSE:   Examples of lapply, sapply, vapply
  *  Must understand lists vs atomic, named


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = T   ) # easier to read
```

### lapply: function works on list, returns a list.
```{r simple_list}

# named list
l  <- list(a = rnorm(5), b = rnorm(5))
l

# named list  
ans  <- lapply(l, mean)
is.list(ans)
ans


args(lapply)
```

### simplification
  *  several functions try to simplify structure
  * c()
  * unlist()

```{r simplify}

# named
x  <- c(a="ONE", b="TWO")
x

# c() returns simplified
x  <- c(1,2, c(3,4))
x

# c() simplies tries to keep names
x  <- c(1, 2, d = c(3,4))
x

x  <- c(1, 2, c(3,4))
is.list(x)   # F
is.integer(x) # F
is.double(x)   # TRUE
```


### unlist as simplifier

```{r unlist}

# unlist() is another simplifier , here returns atomic vector rather than list
l  <- list(a=c(1,2), b=c(3,4))
l

# pattern ?!
ans  <- unlist(l)
ans

is.double(ans)  # T
names(ans)
length(ans)  # 4
```


### sapply:  
  *  Rather than list, try to return vector, matrix, or array (if simplify=TRUE)
```{r sapply}
args(sapply)

# sapply is 'smarter', if see scalers and returns vector[] instead.
sapply(v, mean) # return vector!

# named list 
l <- list(a=1:5, b=6:10)

sapply(l, mean) # named atomic


# first, do not simplify
ans  <- sapply(l, `[`, simplify=F)
class(ans)
attributes(ans)
ans


# second, do simplify (default) Surprise !
ans  <- sapply(l, `[`)
class(ans)
attributes(ans)
ans
```


```{r runif}


#-----------------
# runif | runif(n) returns double[] n values from uniform distr
# ---------

# to check
args(runif)

runif(3)

lapply(1:4,runif) # treats 1:4 as list, retuns list
```

Replicate():
  *  replicate is version of sapply
  *  simply=FALSE ; returns list of vectors
  *  DEFAULT: would return array or matrix

```{r }
n  <- 3

# returns matrix
l  <- replicate(3, rnorm(5))
is.matrix(l)

# returns list
l  <- replicate(3, rnorm(5), simplify = FALSE )
is.list(l)
l

lapply(l, mean)
```


### vapply(x, fun(x), typeof(x))
```{r vapply}

#  integer(1) creates 0
#  integer(2)         0 0
#
n  <- 10^3

# create int[] of length n, filled with 1:n
if (FALSE )
  seq_len(n)

# fills x 1:n
x  <- vapply(seq_len(n), function(i) {i}, integer(1))
head(x)
x  <- NULL
```
### vapply examples
```{r vapply_examples}
n  <- 10^3
# creates all zeroes
x  <- vapply(seq_len(n), function(i) {integer(1)}, integer(1))
head(x)


# remove entries 
v  <- double()
v  <- c(1,2,3, 2019, 2020, 2021, 2022)
typeof(v)

vapply(v, function(x) {if (x > 2020) x  <- NA; x}, double(1))
```


```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

file  <-"rmd/0089_lapply_sapply.Rmd" 

# in general, pdf will look nicer
rmarkdown::render(basename(file),
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
