
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

---
othercomments
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw
-->

<!--  commments -->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools) 
``` 


```{r begin}
```

Structure
non-binary

dput
```{r}

y  <- list(a=1, b=2)
dput(y)

# keeps meta data
dput(mtcars[1,])
structure(y)

## for binary
## saveRDS

***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
