---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.5in,top=0.25in   
TAGS:  curve(),latex,math
---
* using curve() to explore districtions.

* MOVE 028_  here
## Roughly follows:   OpenInfo
###  Begin, simple Z, T
###  Also, FAT tail discussion

Need X11

### curve() - very easy to generate distribution plots:  Examples:
```{r curve, eval=FALSE }
curve(sin, -1, 1)

curve(dnorm, -3, +3)


# CDF - beautiful!
curve(pnorm, -3, +3)
curve(pnorm)  # clips the x axis


curve(qnorm)  # indep variable is now alpha, (area) (0,1)
```

$$e^2$$


$$\frac{a}{b}$$

$$\frac{xe^-x}{x+1}$$
$$\frac{xe^x}{x+1}$$


\begin{equation}
y=x
\end{equation}

left align:
$y=x$

\(y=x\)


center?

\[y=x\]

### Need latex for this!
```{r use_latex}

g  <- function(x) {
  num  <- (x)/(x+1) *exp(-x)
  den  <- 1 
  return(num/den)
}

curve(g, .01,10)
```

### T Distribution
```{r t_distribution}
x  <- function(x) dt(x, df=25)
curve(x, -3, 3,)
knitr::knit_exit()

```

### Chi-squared
```{r chi-squared}
curve(
			dchisq(x, df=3),
			xlim=c(0,10),
			ylim=c(0,1),
			col="blue"
)

curve(
			dchisq(x, df=6),
			xlim=c(0,10),
			ylim=c(0,1),
      add=TRUE,
			col="green"
)

# use ADD to include 2nd graph
curve(
			pchisq(x, df=3),
			xlim=c(0,10),
			add=TRUE,
			col="red"
)
```



## Practice with normal
### 00-Given Area(1 tail), find z
```{r}
qnorm(p=.5, mean=0, sd=1)	# z=0
qnorm(p=.05)	# z=-1.64
qnorm(p=0.025)	# z=-1.96
```

### 01-Given T (or quantile) value, find Area to left
```{r}
# Ex 7.2
# area=?
(area <- pt(q=-2.10, df=18 ))	# 0.025
```

### 02-Given T (or quantile) value, find Area to left (Ex7.3)

```{r}
# Ex 7.2
# area=?
(area <- pt(q=-1.65, df=20 ))	# 0.057

# if were normal
(area <- pnorm(q=-1.65 ))	# 0.0495
```
### 02A - speaking of normal:
```{r}
# for 2 tail normal, find area within distance 1 sd of 0
1-(area <- 2*pnorm(q=-1 ))	# 0.68 (1- area in both tails)
1-(area <- 2*pnorm(q=-2 ))	# 0.954	 (1- area in both tails)
1-(area <- 2*pnorm(q=-3 ))	# 0.997 (1- area in both tails)
```

##### 02B - TODO plot - 2 tails!
```{r}

```

### 03-(Ex 7.4) Given T (or quantile) value, find Area in tails
```{r}
# so this is area above or below 3 sd fromd from 0 
# area=?
(area <- 2*pt(q=-3, df=2 ))	# 0.0955

(area <- 2*pnorm(q=-3 ))	# 0.027 (so T has more in far tails)
```

### Ex 7.5
```{r}
# area above -1.79?
1-(area <- pt(q=-1.79, df=19)) # 0.955
```


### 7.14 Constrct: 1-sample confidence
```{r}
# CI = pt est + t*StandardError
# known: x_mean = 4.4, s = 2.3, n=19
(SE = 2.3/sqrt(19))  # 0.528

# assume CI should be 95%,  ie 2.5% each tail
T <- qt(0.025,df=18,lower.tail=TRUE) # -2.1    and (symmetry) +2.1

#
CI = 4.4 + c(1,-1)*2.1*.528  # c(3.291, 5.509)
```


### 7.16 1-sample T
```{r}
# compute T value, difference/SE
SE=16.98/sqrt(100 )   # 1.698

t <- (97.32-93.29)/SE    # 2.373

# this corresp to area (above 2.373), assume 2-tail
(area <- 2*pt(-2.373,df=99))   # 0.0196, which is p-value, smaller than conf level 0.05 , reject H0
```

````

MOTIVATION:  You are finding sample mean.
DATA shows a large deviation. Is a RARE events?   For example, a 4-sigma event
in your sample?  (ie very long tail contribution)

Is it 2 2-sigma events  OR  1
4-sigma events.  

REF: https://david-salazar.github.io/2020/04/17/fat-vs-thin-does-lln-work/

Compare:  For each x, probability of 2 meaurements > 1x stand dev  vs
probability of 1 measurement > 2x stand dev

NOT Surprise:   For very rare events, ratio is large!  So the large deviation
probably caused by 2 2-sigma events, not 1 4-sigma events.  ie  Safe to ignore
tails - in normal distribution.
````

```{r fat_tails}
library(tidyverse)
library(hrbrthemes)

y  <- seq(0,4, 1)
(1-pnorm(y))


x <- seq(0, 3.45, 0.05)


NUM  <- (1 - pnorm(x))^2
DEN   <- (1-pnorm(2*x))

ratio  <- NUM/DEN
#ratio <-((1 -pnorm(x))^2)/(1-pnorm(two_x))
ratio

g  <- {
tibble(ratio) %>% 
  cbind(x) %>% 
  rename('k' = x) %>% 
  mutate(k = as.double(k)) %>% 
  arrange(k) %>% 
  ggplot(aes(x = k, y = ratio)) +
    geom_point(color = 'dodgerblue4') +
  hrbrthemes::theme_ipsum_rc(grid = "Y") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "k in terms of sigma",
       title = "Mediocristan", 
       subtitle = "Ratio of S(X)^2/S(2X) from a Standard Normal")
}
g	
```
```{r print, eval=FALSE}

{
file  <- "rmd/01010_basic_distributions.Rmd" 
file  <- basename(file)
file  <- here("rmd", file)
file
}

# in general, pdf will look nicer
rmarkdown::render(file,
                  output_format = "pdf_document")
                  #output_format = "html_document")
```
