---
title:  "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: hide
        toc: true 
        toc_depth: 2
        toc_float: true
        highlight: pygments
        theme: cerulean
editor_options: 
  chunk_output_type: console
---


<!--
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
-->
## PURPOSE: Basic sampling, 
* Verbs:   sample, rep, repetition
* Collect all examples HERE
* [source Rafal Chapter 14:](https://rafalab.github.io/dsbook/probability.html#discrete-probability)

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```


## replicate vector
```{r rep}
# return a vector, containing x  3 times (3 elments)
rep(x=5,times = 3)  # 5 5 5	

# return vector of "j" "k"  6 times (12 elements)
v <- rep(x=letters[10:11], times=6)
```

## generate seq
```{r seq}
seq(from = 1, to = 10, by=0.1)

x <- seq(1:4)
x
rep(x, 3) # 1...10 1 ...10  1...10
```

## take a sample, returns vector
*  size = repeat this many times (length of return vector)
```{r sample}
sample(x=c(0,1), size= 7, replace=TRUE)

sample(x=c("blue","red"), size=20, replace=TRUE, prob=c(0.8,0.2))

sample(x=1:5, size=50, replace=TRUE)

# note:
m <- matrix(data=1:100, nrow=25) # fills 1st column, then second ...
head(m)

```


## replicate, apply verb repeatedly
* n = repeat number of times
* expr = ususally a call, such as mean(rnorm(10)) to repeat
```{r replicate}

replicate(n =3 , mean(1:100))
replicate(n =3,  mean(rnorm(10)))

```

Construct list of vectors for lapply
  *  replicate is version of sapply
  *  simply=FALSE ; returns list of vectors
  *  DEFAULT would return array or matrix

```{r }
n  <- 3
l  <- replicate(3, rnorm(5), simplify = FALSE )
l

lapply(l, mean)
``
game
```{r game}
# game:  roll 2 dice, 7 or 11?
# construct matrix of 2 cols, so simulate throwing 2 die, 50x.

B  <- 50  # experiments (rows)
n  <- 2   # each experment, 2 dice

s <- sample(1:6,B*n, replace=TRUE) # returns vector
m <- matrix(s, nrow= B, ncol=n)
head(m)
```

replicate is wrapper to repeatedly apply sapply
```{r replicate}
# replicate
```

```{r norm}

#### 0050_from_norm
df <- data.frame(matrix(rnorm(1000), ncol =10))
head(df)	

```
```{r coin flip1}
N  <- 10000
roll <- sample(x=c(-1,1),size=N,replace=TRUE, prob=c(0.5,0.5))
roll  <- as.integer(roll)
is_integer(roll)
is_vector(roll)# T
is_atomic(roll)
str(roll)


t  <- tibble(raw =roll,
	     cum_total = cumsum(raw))
print(t, n=20)
tail(t)
plot( x=1:N, y=t$cum_total)

```
```{r coin_flip}


#### 0060_coin_flip_running_avg
# 1/n converges slowly!
size <- 100000
t <- tibble(x=sample(x=c(1,0),size=size, replace = TRUE),
			run_avg = dplyr::cummean(x)
			) %>% 
	mutate(line = 1:size)

t %>% head()

t$run_avg %>% tail()
t %>% ggplot(aes(x=line, y=run_avg)) +
	geom_point()
	

# create 'urn' with 5 balls
beads <- rep(c("red", "blue"), times = c(2,3))
beads

# pick sample (of 1)
event <- sample(beads,size=1)
event

# many events
set.seed(2012)
B <- 10000
events <- replicate(B,sample(beads,size=1))
table(events)  #  5988 blue, 4012 blue
t <- tibble(events)
	
# shortcut...
events1 <- sample(beads, size = B, replace = TRUE)
table(events1)  # 5988, 4012

x <- rnorm(10000,0,1)
mean(x <= 0)  # cute,  returns area, percentile
```




```{r knit_exit(), eval = FALSE   } 
knitr::knit_exit() 
```

```{r render, eval=FALSE, include=FALSE 	} 
file <- "rmd/01050_basic_sampling.Rmd" 
file  <- basename(file) 
dir <-"rmd"

# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

file  <- here("", "")
file

# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
