---
title:  "`r knitr::current_input()`"
date: "`r lubridate::today()`"
output:
    html_document:
        toc: true 
        toc_depth: 3
        toc_float: true
font_size: 10
---

## Binomial Sampling
Purpose:   Examples of sampling from binomial distribution


### Notation:
  -  p prob success, just 1 event
  -  n obs in 1 sample (size)
  -  k success in 1 sample
  -  B number of repeat samples (n)

### rbinom(n = B, size = n, p = p) returns SUCCESSes

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=FALSE  ) 

library(tibble)
```


##  Toss coin 100 times, # heads?
```{r rbinom}
# 1 sample(set 100 obs)

n  <- 100 # in 1 sample
p  <- 0.5
B  <- 1

set.seed(2020) # 52 successes

x <- rbinom(n = B , size=n, p=p)
x #52
```


## Repeat 40 times, each one is 100 flips, # heads? 
```{r repeat_experiment}
B  <- 40

x <- rbinom(n=B, size=n, p=0.5)
x # int[B], vector of size B, number of SUCCESSes in each  

hist(x)
```

## Want to see each H,T for 100 throws?
```{r sample}

n  <- 100
B  <- 1

x  <- rbinom(n = n, size = B, p =p)
head(x)
```
### Same, using sample() for nicer way to see H,T for 100 throws.
```{r using_sample}
# syntax a little differnt (no B, repeated samples)
x  <- sample(x =  c("H","T"),
		size = n,
		replace = TRUE,
		prob = c(0.5,0.5)
		)
head(x)

# creates 1 long string
# y  <- stringr::str_c(x, collapse="")		# collapse
# head(y)

knitr::knit_exit()
```

```{r counting}

# of the 40 experiments, how many returned > 55 heads?
knitr::knit_exit()
table(x) # of 40 samples, distribution of success totals 

table(x  >  55)   # of 40 samples(each 100), how many more than 55 successes

table(x  >  57) # 

table(x > 60) # 0
```


## Repeating Sampling
```{r }

# ----Series of Bernoulli Trials			----

# int[10]
n  <- 100 # (sample size)
p  <- 0.5
B  <- 10  # take B samples (each n)

# throw dice 100 times, count heads (repeat 10)
rbinom(n = B,size=n, prob=0.5)
```



```{r knit_exit(), eval=FALSE } 
knitr::knit_exit() 
```

\newpage
BEGIN HERE, CI
```{r BEGIN}

# FOR REAL
set.seed(1492)
nExp = 100
nThrows = 1000 # IN EACH TRIAL
X  <- sample(x = c(1,0),
		size= nThrows* nExp,# NUMBER THROWS
		replace= TRUE,
		prob= c(0.5,0.5)
		)
#X # VECTOR_DBL, ATOMIC

#M <- AS.MATRIX(X, NROW = NtHROWs, ncol=nExp, byrow=TRUE)
## matrix fill first by column, then next row
m <- matrix(rbinom(n=nExp*nThrows, 
		   size=1,
		   0.5), 
	    ncol=nExp) # n total number of random
m[1,1]
## test
m1 <- matrix (rbinom(n=nExp*nThrows,size=1,
		     0.5),
	      nrow=nThrows)

mean(colSums(m))
sd(colSums(m))

sqrt(250)  # 15.8

plot(colSums(m))
hist(colSums(m))

head(m)
dim(m)

## What does CI (Confidence Interval)  mean?
#  ============================================
## Suppose want to find percentage of students (many) with tatoos
##
## How ?   SAMPLE
p = 0.4  # (usually do not know)
n = 75   # students, sample size
## FACT
## se_bar = sqrt( p_bar * (1 - p_bar)/n)
##

## Take a Sample  (ie string of yes/no, 1,0)
r  <- sample(x = 0:1,
             size = 75,
             replace = TRUE, 
             prob = c(0.6, 0.4))
r
sum(r) #31
p_bar = sum(r)/n
p_bar   # 0.413

## And the 95% CI?
se_bar = sqrt( p_bar * (1 - p_bar)/n)
se_bar   # 0.0569
L  <- p_bar - 1.96*se_bar
U  <- p_bar + 1.96*se_bar
cat("p_bar = ", p_bar, "\n", "CI = ", "(", L, ", ", U, ")",  "\n")
contains_TRUE =  p_bar >= L   && p_bar <= U
contains_TRUE

#####   OR 
# ===========================================
## What does 95% confident actually mean??
## rbinom (n=1, size = 75, prob= 0.4)retrun string of yes/no, 1,0  (size # trials)
## To me:  n=1 means 1 SAMPLE, 
## size=throws in 1 SAMPLE, 
## return is # successes in 1 SAMPLE
##
## n =2 means, return will be 2 values, one for each sample
##
###  FUNCTION
###
 N=75
t  <- tibble::tibble()

one_sample  <- function(DUMMY = null) {  # 1 sample

    prob = 0.4
    p_bar  <- rbinom(n=1, size=75, prob = prob)/N
    ## And the range?
    se_bar = sqrt( p_bar * (1 - p_bar)/N)
    L  <- p_bar - 1.96*se_bar
    U  <- p_bar + 1.96*se_bar
    # cat("p_bar = ", p_bar, "\n", "CI = ", "(", L, ", ", U, ")",  "\n")
    contains_TRUE =  (L < prob)   && ( prob < U )
    new  <- tibble::tibble(p_bar = p_bar, L=L, U=U, contains_TRUE = contains_TRUE)
    t <<- rbind(t, new)
    invisible()
}


ans  <- lapply(0:99, one_sample)

# view
t
print(t, n=100)

# to see only FALSE   
t[!t$contains_TRUE,]
t[!t$contains_TRUE,]

# order
print(t[order(-t$L),], n=100)

###
f  <- function(x) invisible(print("hi"))
f()
r  <- lapply(0:1, f)


#### ARGH!  need dummy variable
f  <- function(dummy) {print("hi")}

l  <-   lapply (1:2, f)
```
## plots
```{r plot}

file  <- "out"
png(filename = paste0("~/Downloads/print_and_delete/" , file,".png"),
         width = 480, height = 480, units = "px", pointsize = 12,
          bg = "white",  res = NA,
         type = c("cairo", "cairo-png", "Xlib", "quartz")
)

 hist(x)
# boxplot(x, col="green", names=c("UrOx"))

stripchart(x = x, 
           ylim=c(0, 225),
           pch = 1, 
           cex = 0.5,
           frame = TRUE, 
           vertical = TRUE,
           main = "24-hour Ur Ox")

abline(h = 35, col=2)

# plot(x, fake, ylim = c(-5, 5), sub = "t-value = 21.5", type = "p")
# abline(h = fake_mean, col=2)
dev.off()
```

```{r render, eval=FALSE  }
file  <- "rmd/01060_sampling_BINOM.Rmd"
rmarkdown::render(file,  
                  output_format = "html_document",
                  output_file="~/Downloads/print_and_delete/out") 
```

