
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools) 
``` 
### Basics & trivial

logical(0) means?
* vector, logical, but length=0
* not TRUE or false
```{r logical}
x <- logical(length = 0)
is.logical(x)  # TRUE
is.object(x)   # FALSE   !
is.null(x)     # FALSE  
```

""
```{r empty}
"" == TRUE
"" == FALSE
is.na("")
is.null("")
```

NULL
* special type
```{r null}
length(NULL)  #0
is.list(NULL) # FALSE

# If list has no names,  then names returns NULL.
# To catch, this fact is important
is.null(NULL)   # TRUE
```

NA
```{r na}
na<- x
is.na(x)
```
Even though cheating, purrr is best way to see how various pieces work
```{r purrr }
L  <- list(NULL, "", NA, logical(0))
purrr::map(L, is.na)

purrr::map(L,is.na)
g  <- function(x) x == ""

purrr::map(L, g)

```

STUDY: https://github.com/r-lib/gh/blob/master/R/utils.R
ORIGINAL version
```{r has_name }
## original version
## as seen in purrr, with the name `has_names()`
has_name_orig <- function(x) {
 
  nms <- names(x)
  if (is.null(nms)) {
    rep_len(FALSE, length(x))
  } else {
    !(is.na(nms) | nms == "")
  }
}


# IMPROVED version
has_name <- function(x) {
  if (!is.list(x)) stop("Please enter a list")
  nms <- names(x)
  if (is.null(nms)) {
    rep_len(FALSE, length(x))
  } else {
    !(is.na(nms) | nms == "")
  }
}

```

All possible cases to test?
has_name


COLLECT all the test cases for has_name()
## MUST know what each of these return to names()
## Then must know how names() responses to each piece
* is.null(names())
* is.na(names())
* names() == ""
```{r}

x  <- NULL
x   <- ""
x   <- NA
x  <- "hello"
x   <- list()
x    <- list(1,2)
x  <- list(a=1, 2)
x  <- list(a=1, b=2)

L1  <- list( "", NA, "hello", list(), list(1,2), list(a=1,2),
           list(a=1, b=2))
L  <- list(NULL, "", NA, "hello", list(), list(1,2), list(a=1,2),
           list(a=1, b=2))
names("")
names(NA)
```

Purrr is best way to see how everything behaves
```{r purrr_is_great}

purrr::map(L, names)
purrr::map(L, length)
purrr::map(L, is.list)
purrr::map(L, is.na)
```

Must modify get_name()
```{r begin_tests}
# logical(0) has lenth(0), but neither TRUE or FALSE   
# NULL is not a list
# so this a problem.


has_name_orig(NULL)

```

Use improved has_name()
* test collection with vapply
* purrr
```{r improved}

has_name(NULL)
has_name("")
has_name()
has_name(list())
has_name(list(a=1, b=2))
has_name(list(a=1, 2))
has_name(list(1, 2))



vapply(L1, has_name, logical(1))
vapply(list(1,2), has_name, logical(1))

```


remove?
```{r begin}
f  <- function(x = NULL) {
  paste0("x has length= ", length(x), " and is.list? ", is.list(x))
  cat(is.list(x), " ", "length= ", length(x), "\n")
  cat(is.na(x), "\n")
}
# x undefined
x
f(x)

x  <- NULL
f(x)

x  <- ""
f(x)

# ERROR
is.list()
f()

x  <- list()
f(x)

na<-x
f(x)
is.na(x)
```

```{r vapply}
x  <- list(NULL, "", NA, list())
vapply(x, length, logical(c(0,1)))
vapply(x, f, logical(1))
```


Once we have names, use purrr to run against each step in code.
* use purrr to run names() against test collection
* use purrr to run is.list, has_names() against test collection

```{r nms}
x  <- list()
x  <- list(1,2)
x  <- list(a=1, b=2)
x   <- list(1, b=2)

L  <- list(list(1,2), list(a=1, b=2), list(1, b=2))

x  <- NULL
x  <- ""
x  <- NA

names(list(1,2)) == ""
nms  <- names(list(1,2))
nms
is.na(nms)
nms==""

vapply(L, names, character)
purrr::map_chr(L, names)
nms  <- names(x) 
nms
length(nms)
is.na(nms)
nms == ""


}
```


CLEAN UP ABOVE
Then work here.

unlist
```{r unlist}
x  <- list(1, 2, 3)
x
y  <- unlist(x)  # returns dbl atomic vector
typeof(y)

x  <- list(c(1, 2), c(3, 4), c(5, 6))
y  <- unlist(x)  # flattens to single dbl vector 1 2 3 4 5 6
y
typeof(y)


x  <- list(list("a","b"), c(3,4), c(5,6))
x
y  <- unlist(x)				# flatten to character vector
typeof(y)
y

LEGACY, from advancedR  006_functions.R


# 1st unlist
l  <- list(a=c(10,20,30), b=c(TRUE,FALSE), c=c("hello", "bye"))
# return named chr[] vector
y  <- unlist(l)
names(y)
y

# 2nd code:
base::Filter
lapply


# return list
lapply(l, is.numeric)

# extracts only numeric, returns as atomic 
Filter(is.numeric, l)

# to see why, run Filter's code, step-by-step

    f  <-  function(x) is.numeric(x)
    unlist(lapply(l,f))# atomic, logical, plus names
    as.logical(unlist(lapply(l,f))) # atomic, logical vector, no names

ind  <- as.logical(unlist(lapply(l,f))) 
which(ind) # return index which are TRUE
# same as Filter:
l[which(ind)] # subset original list, see all atomic elemtns
```
```{r list_of_lists}

l1  <- 0:9
l2   <-  letters[10:19]

L  <- list(l1, l2)
L
str(L)
is.null(l1)
names(l1)
is.na(names(l1))
```

has_name
```{r has_name}
has_name  <- function(x) { 
if (!is.list(x)) stop("Please provide list")
nms  <- names(x)
if ( is.null(nms) ) {
  base::rep_len(FALSE  , length(x))
}
else !(is.na(x) | nms=="")
}

has_name(NULL)
has_name("")
has_name(list())
has_name(logical(0))
has_name(NA)

# problems
has_name(list(a=1, 2))

v  <- ""
is.null(v)
is.na(v)

v  <- NULL
is.null(v)
is.na(v)


```
***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
