--- 
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  html_document:
    math_jax:  default
  pdf_document: 
    latex_engine: xelatex
    toc: TRUE 
    toc_depth:  4 
fontsize: 10pt 
geometry: margin=0.5in,top=0.25in 
TAGS:   LaTeX,math_jax,pdf,html
---

## Simple latex
  *  PDF, will work
  *  HTML, no guarantees and still confused.
  *  rmarkdown knows to use math_jax for html !

REF: https://gist.github.com/derekmcloughlin/896da22518ef2f3d81b0

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
library(jimTools)
```

## Best, so far:   https://www.overleaf.com/learn

### Symbols
$\emptyset$  
$\forall$  
$\infty$  
$\exists$  
$\nexists$
$\Box$  
$\alpha  $\beta  $\gamma  $\epsilon  
$\sigma$   $\mu$ 

$\bar{x_{i}}$  
$\mu_X$  
$\mu_x$  
$\hat{\theta}$  





### Simple Functions

$\sin{x}$  

$\sum x_{i}$

### Independent Samples

$$\mu_{\bar{x_{1}} - \bar{x_{2}}} = \mu_{1} - \mu_{2}$$


### Standard Errors

$$s_{e} = \sqrt \frac {SSResid}{n - 2}$$


### Expressions  
\[ \sum_{i=1}^{\infty} \frac{1}{n^i} \]   

\[ \frac{\sigma}{N^\frac{1}{2}} \]  

\[ \frac{\sigma}{\sqrt N} \]  

### Integrals
\[ \int\limits_0^1 x^2 + y^2 \ dx \]

***
$$e^2$$


$$\frac{a}{b}$$

$$\frac{xe^-x}{x+1}$$
$$\frac{xe^x}{x+1}$$


\begin{equation}
y=x
\end{equation}

***

left align:
$y=x$

\(y=x\)

***

center?

\[y=x\]
```{r knit_exit(), eval=F, include=F} 
knitr::knit_exit()
```


```{r render, eval=FALSE, include=F	} 
args(rmarkdown::render)
{
file <- "0200_math_latex.Rmd"
file  <- basename(file)
file  <- here("rmd", file)
file
}

rmarkdown::render(input = file) 
    #output_format = c("pdf_document", "md_document"))
    #output_format = "pdf_document",
    #output_dir="pdf")
```
