
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw

vim: to format all urls for md
s/https.*/[&]()/g
-->


## PURPOSE:  Put TinyTex, Blogdown helpers HERE

READ:  TinyTex and why Latex can be a pain
https://yihui.name/tinytex/r/

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
library(tinytex)
```

<!--  RENDER:  see bottom -->

install tinytex (R pkg)
```{r install, eval = FALSE }
remotes::install_github('yihui/tinytex')
install.packages('tinytex')
    ## update, SILENT, and ~ 3 min  - PATIENCE!
    ## had to sudo chmod -R a+w in directories

tlmgr_update()
```

## Update tinytex::, per Yuhi (if already installed)
```{r update}
tinytex::tlmgr_update()

# if this fails, follow this:
tinytex::reinstall_tinytex()
```

## To remove tinytex::
```{r remove}
uninstall_tinytex(force = FALSE )
```




Misc tinytex 
```{r}
library(tinytex)
?tinytex::tlmgr

# read log
tinytex::parse_packages()
```

missing ulem.sty?  (easy way)
```{r add_package}
tlmgr_install( pkgs = c("ulem"))
```


Missing .sty ? (more complete way)

```{r missing, include=FALSE, eval=FALSE		}
library(tinytex)
 

# Missing .sty?
# 1. find package where .sty located
tlmgr_search("/lmodern.sty")

# 2. install that package
tlmgr_install("lm")

# ... 3.  update
tlmgr_update()
``` 

latexmk ... only for .tex files
bauks if sees any .md file 
```{r compile}
# this is wrapper
?latexmk

latexmk(file=file)

# works 
tinytex::pdflatex("~/code/tex_project//0001_hello.tex")

```



***

```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file  <- ""0220_tiny_tex.Rmd
rmarkdown::render(file,
                  output_format = "pdf_document",
#                 output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
