
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw

vim: to format all urls for md
s/https.*/[&]()/g
-->

READ:  TinyTex and why Latex can be a pain
https://yihui.name/tinytex/r/

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools)
``` 

install
```{r install, eval = FALSE }
remotes::install_github('yihui/tinytex')

# update, SILENT, and ~ 1 min
tlmgr_update()
```
help
```{r}
library(tinytex)
?tinytex::tlmgr

# read log
tinytex::parse_packages()
```

missing ulem.sty?
```{r add_package}
tlmgr_install( pkgs = c("ulem"))
```


latexmk ... only for .tex files
bauks if sees any .md file 
```{r compile}
?latexmk

latexmk(file=file)

```



***

```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file <- ""
file  <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
