---
title: "032_regex_stringr_examles.Rmd"
output:  
  html_document:
    toc: yes
    toc_depth:  1
    number_sections: no 
  pdf_document:
    toc: yes
geometry: margin=0.5in, bottom=.75in
editor_options:
  chunk_output_type: console
---

### CONSOLIDATE here:   regex,stringr (+ related testthat)
### some REQUIREs  X11
### REF:   
###		stringr & regex cheatsheet (rstudio)
###   stat545.com/block028_character-data.html
###		JENNYBC
### =====================================

#### 000_setup
```{r setup, include=FALSE}
	# collapse=TRUE, combines code and output one block
	knitr::opts_chunk$set(echo = TRUE, comment="        ...##",
							collapse=TRUE)
	library(tidyverse)
	library(testthat)
	library(here)
```

##### base::trimws()
```{r}
	x <- c(" hello", " hello ", "hello ")
	base::trimws(x)
```

#### 002_basics_using HTML (see stringr cheat)
```{r basics, eval=FALSE, echo=FALSE, include=FALSE}
# from cheatsheet, render as html (cute!)

see <- function(rx) stringr::str_view_all("abc ABC 123\t.!?\\(){}\n", pattern = rx)
see1 <- function(rx) stringr::str_view_all("abc ABC 123 ^ \t.!?\\(){}\n", pattern = rx)

see("a") # a
see("\\.")
see("[[:blank:]]")
see("[[:space:]]")    # note!   \n  is a 'space' charcter, 
```

#### 003_string_is_a_string  !# {{{
```{r}
info  <- function(x) {

	# bug .. what is 'TRUENULL'
	print(
				cat(is.character(x), 
						"|", 
						is.vector(x)	)
	)
	length(x) }

# all TRUE
info("a")
info("aa")
info(c("a","aa"))
```# }}}


#### 003_how_R_sees_string_AFTER_parsing# {{{
```{r}
# writeLines(<type this>)   #     R sees this is regex	

# writeLines("<regex>")    # 
writeLines("\\")   # \
writeLines(".")    # .
writeLines("[[:alpha:]]") # [[:alpha:]]
```# }}}



#### 002A-stringr, detect META char
```{r}
s <- "abc^def^^" 
stringr::str_detect(s, "\\^+") # TRUE
stringr::str_detect(s, "\\*") # FALSE

# WORKS
stringr::str_replace_all(s, "\\^+", "_")
```


#### 0004_find_replace_spaces
```{r}
s <- "one      two three "
stringr::str_view_all(s, "[[:blank:]]+")

# works!
stringr::str_replace_all(s, "[[:blank:]]+", "_")
```

#### 0005_find_replace_lead_digits
```{r}
s <- "012345_junk_99"
stringr::str_replace_all(s, "^[[:digit:]]{4,6}", "")
```

#### 0006_test_that
```{r}
s <- "012345_junk_99"
stringr::str_replace_all(s, "^[[:digit:]]{4,6}", "")


f <- function(string="abc",
			  pattern="[[:alpha:]]", 
			  with="!"){
					stringr::str_replace_all(string, pattern, with)
}

f()
testthat::test_that("remove leading digits", {
	expect_equal(f(string="012345_junk_99",
				   pattern = "^[[:digit:]]{4,6}",
				   with = ""),
				 
				   "_junk_99")
	
	expect_equal(f(string="012345_junk_012345",
				   pattern = "^[[:digit:]]{4,6}",
				   with = ""),
				   "_junk_012345")
}
)
```

####010_regex_in_R_help
```{r}
help.search("^geom_[abcd]",
			package="ggplot2")


```


####007_replace_multiple_dashes
```{r}
s <- "a____b_c__d_"
stringr::str_replace_all(s, "_+","_")  # a_b_c_d_   (_ after d, yes)

s <- "a'b'c"
stringr::str_replace_all(s, "'","_")  # a_b_c   

s <- 'a"b"c'
stringr::str_replace_all(s, '"',"_")  # a_b_c   

s <- "a'b'c"
writeLines("'")
writeLines('"')

# not quite
stringr::str_replace_all(s, "'",'"')  # "a\"b\"c"   is what R displays

# remove NA, unless followed by alphanumeric
s <- "one_NA_two_NAME"
stringr::str_replace_all(s, "NA[^[[:alnum:]]]","")  # "one_two_NAME"
```
#### 0010.tidyr::extract.with regex
```{r tidry::extract}
t <- tribble(
			 ~code, 
			 "S01E01",
			 "S02E01",
			 "S03E02"
			 )
t %>% tidyr::extract(code, c("season","number"), 
					 regex="S(.*)E(.*)",convert=TRUE,
			  			remove=FALSE)  
```
####008_regex_tibble_column
```{r}
t <- tibble(original = c("one_two____three",
						 "NA_john_denver"))
t<- t %>% mutate(col2 = stringr::str_replace_all(original, "_+", "_") ) 
t
```
```{r render,include=FALSE, eval=FALSE}
#rmarkdown::render("032_regex_stringr_examples.Rmd", output_dir="html")
#rmarkdown::render("032_regex_stringr_examples.Rmd", output_format="pdf_document", output_dir="pdf")
```

#### misc stringr:: practice ####
```{r eval=FALSE}
commas <- function(...) stringr::str_c(..., collapse = ", ")
commas1 <- function(...) stringr::str_c(...)
what_is_it(letters)
print (letters)  # vector of 26 letters
commas1(letters)
commas1(letters, collapse = ", ")
what_is_it(commas(letters))	# char vector of length 1, ie string

commas1(letters, collapse = "-")	# works

# char vector becomes 1 string
letters # vector, 26
paste(letters, collapse = "+")

# remove 1 layer, list of lists, into simple vector
?purrr::flatten_dbl

# z is list containing 2 lists, each of these is single list
z = list(a = list(as.double(c(1,2))), 
	 b = list(as.double(c(10,11)))
	 )
	 
z[["a"]][[1]][2]	# 2
what_is_it(z[[1]])	# list of 1 elemnt, a list
what_is_it(z["a"])
z["a"]
what_is_it(z[a]) # fails
what_is_it(z)
y <- flatten(z["a"])
what_is_it(y)
y
str(y)

#################
# ---- fruit ----
#################
fruit    # character vector
what_is_it(fruit)  # returns char vector, size 80
stringr::str_detect(string=fruit, pattern="fruit") # returns logical 
what_is_it ( stringr::str_detect(string=fruit, pattern="fruit") )# logical

# returns character vector
(my_fruit<-stringr::str_subset(string=fruit, pattern="fruit"))
what_is_it (my_fruit)  # character vector, length 8

(z<-str_split(my_fruit, pattern=" "))
what_is_it (z) # a list of char vectors,  fruits containing space, get split

(z1 <-str_split_fixed(my_fruit, pattern=" ", n=2))
what_is_it(z1)  # vector of typke character, length = 16

# collapse
str_c(my_fruit,collapse= " ")
str_c(my_fruit,collapse= ".")
str_c(my_fruit,collapse= "|")
```
#### 0009_EXPERIMENTS,  NASTY string replacements
```{r}

{
	remove_weird  <- function(v) { 
	step2 <- stringr::str_replace_all(v,"[:punct:]+","_") 
	return (step2)
	}
}
# Works
	expr = "!@#%*-?,'"
	remove_weird(expr)

# Fails
	# ^$.()
	remove_weird("^")  # FAILS

# TODO must remove ^
	{
	test_that("remove_weird removes '", {
							expect_equal(remove_weird("abc'def''ghi"),"abc_def_ghi")
							expect_equal(remove_weird("abc'''def''ghi'"), "abc_def_ghi_")
							expect_equal(remove_weird("%%"), "_")
							expect_equal(remove_weird(expr), "_")
							})
	}

# FAILS
	stringr::str_replace_all("!?",fixed("!\\?"),"_")
stop()
```

#### stringr, grepl (need X11, html)
```{r}
val <- function(r) stringr::str_view_all("1.11.111.011", r)
val("[1]{2}")
val("^1{2}")
val("^[1]{2}")
val("[1]{2}")
# render as html
val <- function(r) stringr::str_view_all("1.11.111.011", r)
val("[1]{2}")
val("^1{2}")
val("^[1]{2}")
# render as html
val <- function(r) stringr::str_view_all("1.11.111.011", r)
val("^[1]{2}")
val("\^[1]{2}")
val("[1]{2}")
val("^1{2}")
val("^[1]{2}")
val("a")
# render as html
val <- function(r) stringr::str_view_all("1.11.111.011.a1", r)
val("a")
# render as html
val <- function(r) stringr::str_view_all(c("1.11.111.011.a1"), r)
val("[1]{2}")
# render as html
val <- function(r) { stringr::str_view_all(
c("1.11.111.011", "a1"), r)
}
val("[1]{2}")
val("^1{2}")
val("a")
val("^a")
grepl(pattern="[01]+", c(1,"2","a1","1",00,01,11,000,010,011))
grepl(pattern="^[0,1]{2,3}", c(1,"2","a1","1",00,01,11,000,010,011))
grepl(pattern="^[01]{2,3}", c(1,"2","a1","1","00",01,11,000,010,011))
grepl(pattern="^[0,1]{2,3}", as_character(c(1,"2","a1","1",00,01,11,000,010,011)))
grepl(pattern="^[0,1]{2,3}", c(1,"2","a1","1",00,01,11,000,010,011))
grepl(pattern="^[0,1]{2,3}", as.character(c(1,"2","a1","1",00,01,11,000,010,011)))
grepl(pattern="^[01]{2,3}", c(1,"2","a1","1","00",01,11,000,010,011))
as.character(c("1",1,9))
grepl(pattern="^[01]{2,3}", as.character(c(1,"2","a1","1",00,01,11,000,010,011)))
grepl(pattern="^[01]{2,3}", c(1,"2","a1","1","00",01,11,000,010,011))
as.character(c(1,"2","a1","1",00,01,11,000,010,011))

```

#### STRINGR
# ============

#### 001-stringr, regex:  find patterns
```{r}

# view matches (needs browser, html)
s <- "a.aa.aaa.aaa"
stringr::str_view_all(s, "a{3}")

# or TRUE or FALSE
stringr::str_detect(s,"a{3}")
```

#### 002-stringr, regex practice
```{r}

```

#### 002A-stringr, detect META char
```{r}
s <- "abc^def^^" 
stringr::str_detect(s, "\\^+") # TRUE
stringr::str_detect(s, "\\*") # FALSE

# WORKS
stringr::str_replace_all(s, "\\^+", "_")


```
#### 003 stringr, remove number
```{r}
# remove prefix
s <- "00001_file"
s <- stringr::str_remove(s,pattern = "^[:digit:]{4,6}")
s

# prepend prefix
s <- stringr::str_c("00001",s)
s
```

####004_stringr practice with tribble - practice
```{r}
toy <- tribble(
  ~x,
  "apple",
  "0001cherry",
  "000002orange"
)

# new cleaner col
toy %>% mutate(new = stringr::str_remove(x,  "^[:digit:]{4,6}"))
```

###  BELOW:   Development work for pkg_mp3_files
# =================================================
```{r}


#### 0000_PATH (TO DO:  should not need)
setwd("~/code/r_try_things_here")
getwd()
here::here()

#### 0001_function
# What does terminal think?
get_terminal_dir <- function(){ list(proj_path(), proj_get() )}
get_terminal_dir()

##### 0010_simple cleaning example ----
# =============================
t <- tibble::tibble(original = c("one_two____three", "NA_john_denver"))
t
t<- t %>% mutate(col2 = stringr::str_replace_all(original, "_+", "_") ) 
t

# BEGIN
# =======

#	remove_comma
# ==============
remove_comma  <- function(v){  stringr::str_replace_all(v,",+","_") }
remove_comma(c("abc,def,,ghi", ",,,",",jim,"))

remove_underscore <- function(v){  stringr::str_replace_all(v,"_+","_") }

# remove_appostrope
# ===================
remove_appostrope  <- function(v){stringr::str_replace_all(v,"'+","_")}
remove_appostrope(c("abc'def'ghi", "'''","'jim'"))

# remove_caret (special META)
# ==============
remove_caret  <- function(v){stringr::str_replace_all(v,"\\^+", "_")}
s <- "abc^def^^" 
remove_caret(s)

remove_dash  <- function(v){stringr::str_replace_all(v,"\\-+", "_")}


# example: str_detect
# =====================
s <- "abc^def^^" 
stringr::str_detect(s, "\\^+") # TRUE
stringr::str_detect(s, "\\*") # FALSE


# PUT it together?   (try glue?   fancier way to concatinate multiple regex?)
v  <- "-abc__d-ef^^_gh'_^,'"
remove_underscore(remove_dash(remove_caret (remove_appostrope(remove_comma(v)))))


# remove_multiple
# =================
# any 1 or more of _ ' , ^ - <space>     Last 3 are META, so esccape
remove_multiple  <- function(v){stringr::str_replace_all(v,"[_',\\^\\-]+","_")}
remove_multiple(v)

# Add error checking
# ====================

# even handles a char[]
remove_multiple1  <- function(v){
	if(is.null(v)) stop("please supply a string")
	if(!is.character(v)) stop("please supply a character string")
	stringr::str_replace_all(v,"[_',\\^\\-\\s]+","_")}
}
v  <- c("-abc__d-ef^^_gh'_^,'","'''"," 3    ")
remove_multiple1 (v)

# Testing (move to test folder)
# =========
test_that("remove_comma removes ','", {
						expect_equal(remove_comma("abc,def,,ghi"),"abc_def_ghi") })

test_that("remove_appostrope removes '", {
						expect_equal(remove_appostrope("abc'def''ghi"),"abc_def_ghi")
						expect_equal(remove_appostrope("abc'''def''ghi'"), "abc_def_ghi_")
						})



test_that("str_length is number of characters", {
  expect_equal(str_length("a"), 1)
  expect_equal(str_length("ab"), 2)
  expect_equal(str_length("abc"), 3)
	expect_equal(str_length("abcd"),4)
})
