
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
---
title: "050A_basic_samping"
output: 
  html_document: 
    highlight: pygments
    theme: cerulean
    toc: yes
editor_options: 
  chunk_output_type: console
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw
-->

<!--  commments -->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools) 
``` 


```{r begin}
```

***


/newpage
```{r setup}
### Basic sampling, percentile, quartile
### source: Rafal Chapter 14: https://rafalab.github.io/dsbook/probability.html#discrete-probability
			
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```
```{r rep}

#### 0010_rep()
# return vector of 5s,   3 times
rep(x=5,times = 3)  # 5 5 5	
```
```{r seq}

#### 0020_seq()
seq(from = 1, to = 10, by=0.1)

x <- seq(1:10)
x
rep(x, 3) # 1...10 1 ...10  1...10
```
```{r sample}

#### 0030_sample()
sample(x=c(0,1), size= 7, replace=TRUE)

sample(x=c("blue","red"), size=20, replace=TRUE, prob=c(0.8,0.2))

#### 0040_matrix()
m <- matrix(data=1:100, nrow=25) # fills 1st column, then second ...
m

# 7 or 11
sample(1:6,2)

# game:  roll 2 dice, 7 or 11?
# construct matrix of 2 cols, so simulate throwing 2 die, 50x.
B  <- 50  # experiments (rows)
n  <- 2   # per die experiment

s <- sample(1:6,B*n, replace=TRUE) # returns vector
m <- matrix(s, nrow= B, ncol=n)
m
```
```{r norm}

#### 0050_from_norm
df <- data.frame(matrix(rnorm(1000), ncol =10))
head(df)	

```
```{r coin_flip}


#### 0060_coin_flip_running_avg
# 1/n converges slowly!
size <- 100000
t <- tibble(x=sample(x=c(1,0),size=size, replace = TRUE),
			run_avg = dplyr::cummean(x)
			) %>% 
	mutate(line = 1:size)

t %>% head()

t$run_avg %>% tail()
t %>% ggplot(aes(x=line, y=run_avg)) +
	geom_point()
	

# create 'urn' with 5 balls
beads <- rep(c("red", "blue"), times = c(2,3))
beads

# pick sample (of 1)
event <- sample(beads,size=1)
event

# many events
set.seed(2012)
B <- 10000
events <- replicate(B,sample(beads,size=1))
table(events)  #  5988 blue, 4012 blue
t <- tibble(events)
	
# shortcut...
events1 <- sample(beads, size = B, replace = TRUE)
table(events1)  # 5988, 4012

x <- rnorm(10000,0,1)
mean(x <= 0)  # cute,  returns area, percentile
```

Create SAMPLE
```{r assess_normality}
# REF:  hannay https://faculty.nps.edu/rbassett/_book/continuous-random-variables.html
# Chapter 8.5 Assessing Normality
set.seed(2020)
B  <- 1000
fake.data  <- rnorm(B) # 100 points

# B bins
# exclude 0, 1 (z not finite)
seq(1/(B+1), B/(B+1), 1/(B+1))


```
```{r histogram}

# histogram

```
Some ratios
```{r ratios}

# Actuals
sd(fake.data) # 1.12 (vs 1.0 theory)

quantile(fake.data, 1/4)
quantile(fake.data, 3/4)

quantile(fake.data, 3/4) - quantile(fake.data, 1/4)
IQR(fake.data) # 1.3 (vs 1.3 theory

IQR(fake.data)/sd(fake.data)   # 1.16 (vs 1.3 theory)
# 
```{r plot}
t  <- tibble(fake.data = fake.data)

g  <- ggplot(t, aes(sample=fake.data) ) +
	stat_qq() +
	stat_qq_line()
# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")
```

TODO:  
Follow Hannay (above)
But replace with UNFIMR distr
```{r quantile}

# 100 points from UNIFORM DISTR, 0-1
set.seed(1492)
n  <-20 
t  <- 30*tibble(sample = sort(runif(n))  )

# Find sample quantiles corresponding to 1/20, 2/20 % ... 20/20 %  etc
probs  <-  seq(0 +1/n, 1, 1/n)

# corresponding pts, % points to left
x  <- quantile(t$sample, probs=probs)
x
t  <- t %>% mutate(quantiles = x, percent=probs)
t

# PLOT just sample points
g  <- ggplot(t, aes(x=sample,y=1) ) +
geom_point()
# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")

# PLOT quantiles corres to specific percentages
s  <- t %>% filter(percent == c(0.25, .50, .75, 1.00))


## NEXT Plot both sample data AND indicate quantiles.
g  <- ggplot(s, aes(x=quantiles,y=0, color="blue") ) +
geom_point() +
geom_point(aes(x=t$sample, y=1)) + 
# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")

```

```{r qq_plot}
g  <- ggplot(mtcars, aes(sample=mpg) ) +
	stat_qq() +
	stat_qq_line()
# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")


# 32 x 1
t  <- tibble(mpg = mtcars$mpg)
q_sample  <- quantile(mtcars$mpg, probs = seq(0, 1, 0.05), na.rm = FALSE,
              names = TRUE, type = 7 )

p_normal  <- pnorm(seq(-2, 2., 0.05))
p_normal
q_normal  <- qnorm(seq(0,1,0.05))

q_sample  <- quantile(mtcars$mpg, probs = seq(0, 1, 0.05), na.rm = FALSE,
              names = TRUE, type = 7 )
q_normal

```


```{r knit_exit()} 
knitr::knit_exit() 
```

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
