-https://mblz.io/x6TUAA--
title: "053_md_examples.Rmd"
output: 
  html_document: 
    highlight: espresso
    theme: readable
editor_options: 
  chunk_output_type: console  
params:  
  d:  !r Sys.Date()
---
\underline{\textsc{\today}}
```{r}
params$d
```

**To Render**
```{r ren_pdf, eval=FALSE	 }
library(here)
ren  <- function() {
	rmarkdown::render(
									here::here("r_try_things_here","053_md_examples.Rmd"), 
									output_format="pdf_document", 
									output_dir="~/Downloads/print_and_delete/"
									)
}
ren()
```

<!-- This is a comment -->

---

Another comment.  Be sure leave blank lines

---

Practice:  .md, knitr

See:  
064_  for Knitr chunks, YAML

***  
###	Header 3  
####	Header 4  
#####	Header 5  

***

This is **ordered list**

1.	Item One
2.	Item TWo
3.	Item Three
	i) Sub-item one
	ii) Sub-item two
4.	Item Four

***

This is `verbatim code`

>This is block quote
>

>This is longer block quote. This is longer block quote.  This is longer block
quote.  This is longer block quote. This is longer block quote. This is longer
block quote. This is longer block quote. This is longer block quote. 
>


```
Leave blank lines so knitr isn't confused.
This is longer block quote. This is longer block quote.  This is longer block
quote.  This is longer block quote. 

```


***

`***` adds a horizontal line

***

~~~
fenced by ~
~~~

***

This is UNordered list

*	Item One
	+ Sub-Item One
	- Sub-Item Two
	* Sub-Item Three
*	Item Two

****

Definition 1

:  Definition Unlawful is ....

Definition 2
:  Definition 2 means ...

****

And now an R chunk?

``` {r eval=FALSE}
rmarkdown::render("053_md_examples.Rmd", output_dir="html")
```
```{r render,include=FALSE, eval=FALSE}
#rmarkdown::render("032_regex_stringr_examples.Rmd", output_dir="html")
#rmarkdown::render("032_regex_stringr_examples.Rmd", output_format="pdf_document", output_dir="pdf")
```


\underline{underline}
\textsc{\today}
\pagebreak
