---
title: "066_fs_examples.Rmd"
output:
  pdf_document:
    extra_dependencies: ["booktabs"]
    toc: yes
  html_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

 =====================
Rename:  knitr chunks
=====================
*  Other regex, see 0032
Using regex, stringr to gather knitr chuncks
Neatly rename the chunks!

```{r include=FALSE }
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(here)
library(fs)
library(lubridate)
library(magrittr)
library(tibble)
library(dplyr)
```

print current dir, clean up time
```{r}
knitr::opts_chunk$set(echo= TRUE, include= TRUE)

# dir_info, capture as_tibble
# convert Positct date-time to simple date

t <- as_tibble(dir_info()) %>% 
  mutate(modification_time = lubridate::as_date(modification_time)) %>% 
  select(path, size, modification_time )
print(t, n=500)
`'`

get names of files in rmd
```{r get_files}
# Remove path info with path_file() 
t <- as_tibble( fs::dir_info(here("rmd"))) %>% 
  mutate(path = fs::path_file(path),
    modification_time = lubridate::as_date(modification_time)) %>% 
	select(path, size, modification_time )

t
```

(I stole this code)
figure this out!
* find all chunks
* neatly numbers them?

```{r challenge_regex}

name_chunks_mlr3book = function() {

  here()
  # full.names = FALSE for file names only
  rmds = list.files(here("rmd"), 
    pattern = "\\.Rmd$", full.names = TRUE, recursive = TRUE)

  rmds

  pattern = "^([[:space:]]*```\\{[rR])([[:alnum:] -]*)(.*\\})[[:space:]]*$"
  
  for (rmd in rmds) {
    message(sprintf("Renaming chunks in '%s'", basename(rmd)))

    # read all the lines
    lines = readLines(rmd)

    # which lines, spaces, then ```, anything, then }
    # misses a few with ::spaces:: after final }
    ii  <- which(
                 stringr::str_detect(
                                     lines, 
                                     "^[[:space:]]*```\\{[rR].*\\}$"))


    # START HERE
    labels = sprintf("%s-%03i", stringi::stri_replace_last_fixed(basename(rmd), ".Rmd", ""), seq_along(ii))

    lines[ii] = stringi::stri_replace_first_regex(lines[ii], pattern, sprintf("$1 %s$3", labels))

    writeLines(stringi::stri_trim_right(lines), con = rmd)
  }

  invisible(TRUE)
}
```

```{r}
knitr::knit_exit()

```

