---
title: '063_grolemund_map.Rmd'
output: html_document    
date:  '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://resources.rstudio.com/webinars/how-to-work-with-list-columns-garrett-grolemund"
```{r message=FALSE}
library(tidyverse)
library(babynames)
Sys.Date()
```

create tibble
```{r}
y <- tibble(
		a = c(1.0, 2.0, 3.14),
		b = c( "a","b","c"),
		c = c(TRUE, FALSE, FALSE),
		d = list(as.integer(c(1,2,3)), TRUE, 2L)
	)
y
```

dplyr::mutate: tibble->tibble;   
sqrt: vector -> dbl vector
```{r}
y %>% mutate(asc_a =  sqrt(a)) %>% print()
```
*
Error:  sqrt rejects list
```{r error=TRUE}
y %>% mutate(asc_d = sqrt(d) %>% print())
```

- **instead**, dplyr::map applies list to sqrt, element-by-element and repackages as list.
```{r}
y %>% mutate(asc_d = map(d, sqrt)) %>% print()
```

## Babynames, filter names present every year
```{r}
everpresent <- babynames %>%
	group_by(name, sex) %>%
	summarize(years = n()) %>%
	ungroup %>%
	filter(years == max(years))
```

## Keep all rows (and fields) in x which match group_by in y
```{r}
babynames <- babynames %>%
	semi_join(everpresent)
```

## most popular, each year
```{r}
babynames %>%
	group_by(year,sex) %>% 
	filter(prop == max(prop))
```
## (test)
```{r}
mtcars %>%
	group_by(cyl)  %>%
	top_n(1,hp)
```
```{r}
babynames %>%
	group_by(year,sex) %>% 
	top_n(1,prop) %>%
	arrange(desc(year))
```

## select joe, all years
```{r}
joe <- babynames %>%
	filter(name == "Joe")
joe %>%
	ggplot(aes(x=year,y=prop)) +
	geom_point() +
	geom_line() +
	geom_smooth(method=lm,se=FALSE)

```

## is linear a good fit for joe?
```{r}
fit <-  lm (prop ~ year, data=joe)
library(broom)
pluck(coef(fit),"year")
pluck(glance(fit), "r.squared")
```

@32:00
model for every name in babynames
```{r}
babynames %>%
	group_by(name,sex) %>%
	nest()
```

retrieve Mary
```{r}
babynames %>%
	group_by(name,sex) %>%
	nest() %>%
	pluck("data") %>%
	pluck(1)
```

use map to run lm interatively over list "data"
```{r}
d<-babynames %>%
	group_by(name,sex) %>%
	nest() %>%
	mutate(model =
		   	map(data,
				 ~lm(prop ~ year, data=.x)),
		   slope =
		   	map_dbl(model,
		   			~pluck(coef(.x), "year")),
			r2 =
				map_dbl(model, 
					~pluck(glance(.x), "r.squared"))
	)
save(d,file="baby_model")

```

verify "Mary"

```{r}
d %>% pluck("name") %>% pluck(1)
d %>% pluck("model") %>% pluck(1)
```