---
title: '063_grolemund_map_list_column.Rmd'
output: html_document    
date:  '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: 
  chunk_output_type: console
---

* typeof() - built-in R structures
* attributes - can add META data.
* class - is particular attribute, and makes object S3.

REF:  video-
https://resources.rstudio.com/webinars/how-to-work-with-list-columns-garrett-grolemund"
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
#library(babynames)
```

```{r basics}
x  <- c(1L, 2L, 3L)
typeof(x) # integer
```
Create new data types by adding class or other attribute
Date
```{r date}
# make it Date (actually S3) 
class(x)   <- "Date"

# but still integer[]
typeof(x) # STILL an integer
attributes(x) # only class
```
Array
```{r array_matrix}
# make an array
  a  <- array(1:8)
  a

# What is array?
  typeof(a)   # integer[]
  attributes(a) # dim: 8 
  class(a)    # array
```
Matix
```{r matrix}
# array again
  a  <- array(1:8)

# make a matrix
  attr(a, "dim")  <- c(2,4)

# What is matrix?  
  typeof(a)  # integer[]
  class(a)   # matrix, array
  attributes(a) # 2, 4
```

matrix - slight difference
```{r matrix}
# Begin with integer[]
x  <- c(1L, 2L, 3L)


# Add attributes
  dim(x)  <- c(3,1)
  # same as
  # attr(x, "dim")  <- c(3,1)

# What is matrix? 
  typeof(x)   # STILL an integer
  class(x)    # matrix, array
  attributes(x)  #  now 3,1

```

factor or categorical variable
```{r categorical}
# Begin with integer[]
x  <- c(1L, 2L, 3L)

# Add attributes & class
  levels(x)  <- c("BLUE", "BROWN", "BLACK")
  class(x)  <- "factor"

# What is factor?
  typeof(x)   # STILL an integer
  attributes(x) # now 2!, levels and class=factor
```


HOW TO MIX data types in 1 container?
use list
```{r list}
# Begin list of named atomic vectors.
l  <- list(a=c(1,2,3), b=c(TRUE, TRUE, FALSE	))

# Check
typeof(l)  # list
attributes(l) #names
```

data tables, is  a named LIST of vectors.
```{r table}
# Begin list of named atomic vectors.
l  <- list(a=c(1,2,3), b=c(TRUE, TRUE, FALSE	))

# Add class, attributes
class(l)  <- "data.frame"
rownames(l)  <- c("1", "2", "3") 

# What is data.frame?
typeof(l) # STILL a list
attributes(l)   #names, row and class
```


data.frame with list as column works, but problem
```{r}
# lists ARE vectors and therefore can be  column  table. 
l  <- list(a=c(1,2,3), b=c(TRUE, TRUE, FALSE	))
l$d  <- list(p=1:3, q=4:5, r=c(letters[10:12]))

# Add class, attributes, as above
class(l)  <- "data.frame"
rownames(l)  <- c("1", "2", "3") 


# What is data.frame?
typeof(l) # STILL a list
attributes(l)   #names, row and class

# Here's the problem:
# Did what we said, made column d list, but not so easy to manipulate.
l
```
try again, but as tibble  - not exactly
```{r tibble}
l  <- list(a=c(1,2,3), b=c(TRUE, TRUE, FALSE	))
l$d  <- list(p=1:3, q=4:5, r=c(letters[10:12]))

m  <- l

# Add class, attributes, as above
class(l)  <- c( "tbl_df", "tbl", "data.frame")

# next line works
attr(l, "row.names")  <- c("1", "2", "3") 
# not this one.
# rownames(l)  <- c("1", "2", "3")   Depreciated
l
```

tibble: much better, why?
now comes in as <named list>
```{r tibble_better}
# Better way
m  <- as_tibble(m)

# What is tibble?
typeof(m)
attributes(m)

m
```

BEGIN HERBEGIN HEREE.
```{r}

# Tibble does it better than df.
y <- tibble(
		a = c(1.0, 2.0, 3.14),
		b = c( "a","b","c"),
		c = c(TRUE, FALSE, FALSE),
		d = list(as.integer(c(1,2,3)), TRUE, 2L)
	)
y   # note d  is a list-column
```

R has tools for atomic vectors and for data tables.
Less so for lists.

dplyr::mutate: tibble->tibble;   
```{r}
# challenge:   convert to 10 x 3 tibble, all int[]
test <- tibble(a = 1:10, b = tibble(x = 11:20, y = 21:30))
typeof(test)
str(test)
```

sqrt: vector -> dbl vector
```{r}
y %>% mutate(asc_a =  sqrt(a)) %>% print()
```
*
Error:  sqrt rejects list
```{r error=TRUE}
y %>% mutate(asc_d = sqrt(d) %>% print())
```

- **instead**, dplyr::map applies list to sqrt, element-by-element and repackages as list.
```{r}
y %>% mutate(asc_d = map(d, sqrt)) %>% print()
```

## Babynames, filter names present every year
```{r}
everpresent <- babynames %>%
	group_by(name, sex) %>%
	summarize(years = n()) %>%
	ungroup %>%
	filter(years == max(years))
```

## Keep all rows (and fields) in x which match group_by in y
```{r}
babynames <- babynames %>%
	semi_join(everpresent)
```

## most popular, each year
```{r}
babynames %>%
	group_by(year,sex) %>% 
	filter(prop == max(prop))
```
## (test)
```{r}
mtcars %>%
	group_by(cyl)  %>%
	top_n(1,hp)
```
```{r}
babynames %>%
	group_by(year,sex) %>% 
	top_n(1,prop) %>%
	arrange(desc(year))
```

## select joe, all years
```{r}
joe <- babynames %>%
	filter(name == "Joe")
joe %>%
	ggplot(aes(x=year,y=prop)) +
	geom_point() +
	geom_line() +
	geom_smooth(method=lm,se=FALSE)

```

## is linear a good fit for joe?
```{r}
fit <-  lm (prop ~ year, data=joe)
library(broom)
pluck(coef(fit),"year")
pluck(glance(fit), "r.squared")
```

@32:00
model for every name in babynames
```{r}
babynames %>%
	group_by(name,sex) %>%
	nest()
```

retrieve Mary
```{r}
babynames %>%
	group_by(name,sex) %>%
	nest() %>%
	pluck("data") %>%
	pluck(1)
```

use map to run lm interatively over list "data"
```{r}
d<-babynames %>%
	group_by(name,sex) %>%
	nest() %>%
	mutate(model =
		   	map(data,
				 ~lm(prop ~ year, data=.x)),
		   slope =
		   	map_dbl(model,
		   			~pluck(coef(.x), "year")),
			r2 =
				map_dbl(model, 
					~pluck(glance(.x), "r.squared"))
	)
save(d,file="baby_model")

```

verify "Mary"

```{r}
d %>% pluck("name") %>% pluck(1)
d %>% pluck("model") %>% pluck(1)
```
