
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```


```{r library, include=FALSE		}
library(jimTools) 
``` 

```{r begin}
#' @title Name all chunks
#'
#' @description
#' Names all chunks of all Rmd files using the pattern `[file-name]-[number]`.
#'
#' @export
name_chunks_mlr3book = function() {
  root <- "~/code/r_try_things_here"
  the_dir  <- "rmd/"
  path <-  file.path(root, the_dir)
  rmds = list.files(
    path, pattern = "\\.Rmd$", 
    full.names = TRUE, 
    recursive = TRUE)
  rmds
  pattern = "^([[:space:]]*```\\{[rR])([[:alnum:] -]*)(.*\\})[[:space:]]*$"

  for (rmd in rmds) {
    rmd  <- rmds[[1]]
    message(sprintf("Renaming chunks in '%s'", basename(rmd)))

    lines = readLines(rmd)
    # ii = line # with begin with pattern
    ii = which(
      stringi::stri_detect_regex(
        lines, 
        "^[[:space:]]*```\\{[rR].*\\}$")
    )
    # removes .Rmd, appends index, saves in labels
    labels = sprintf(
      "%s-%03i", 
      # replace str, pat, with, 
      stringi::stri_replace_last_fixed(
        basename(rmd), 
        ".Rmd", 
        ""), 
    seq_along(ii))

    # prep new lines
    lines[ii] = stringi::stri_replace_first_regex(
      lines[ii], 
      pattern, 
      sprintf("$1 %s$3", labels))

    #writeLines(stringi::stri_trim_right(lines), con = rmd)
  }

  invisible(TRUE)
}

```

***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
