---
title:  "011B_misc_fun_operators.Rmd"		
output: 
  pdf_document:
    toc: TRUE 
    toc_depth: 1
    latex_engine: xelatex
fontsize: 12pt
geometry: margin=0.5in,top=0.25in
#  md_document:
#  html_document:
#  github_document:
#   toc: true
#  	toc_float: true
#    css: styles.css
---
#### basic traceback, debug



> Tools:
> base::browser()
> base::debug(FUN)
-	base::undebug(FUN)
> base::trace(FUN, exit=<action>)
rlang::trace_back()
options(error= browser, trace_back etc)

#### From Gillespie, Appendix E# {{{
```{r setup}
```

```{r library}
here()
```

```{r render, eval=FALSE,	 include=FALSE, echo=FALSE	}
file="010_oo.Rmd"
dir="book_advanced_R/R"
ren_pdf(file,dir)
ren_github(file, dir)

#knitr::knit_exit()
```
```





```{r 1 }
## IMPORTANT TOOLS
# there is no bug()
{
	print("jim")
first <- function() second()
second <- function() third()
third <- function() fourth()
fourth <- function() fifth()
fifth <- function() {
	bug()}

}
#

tracingState() # this a Global value, either on or off

# browser
# Must insert browser() into code.

# on error, use traceback()
first()
traceback()
lobstr::cst()

# use debug
debug(fifth)
first()
undebug(fifth)

#use trace
trace(fourth, browser)
first()
untrace(fourth)

#trace, another way
trace("third", print("inside 3th"))
first()
untrace(c("third", "fourth"))
first() # runs to bug()


#trace, using recover()
trace(third, recover())


rlang::last_error()
rlang::last_trace()
```# }}}

#### E2, browser(), pauses R, view in function's env


#### E4, debug(FUN), acts as browser() when entering FUN# {{{
```
debug(FUN)  - debug a FUN
undebug(FUN) - turn off
debugonce(FUN) - debug a FUN and turn if off
```

```{r E4 }
debug(second)
isdebugged(second)
isdebugged(third)

first()
undebug(second)
```# }}}

#### E5, trace(FUN,browse, at=)# {{{
```
trace inserts a command in FUN, at line #
untrace(FUN) to end
```

```{r E5}
trace(print)
print("hello")

trace(print, browser)
print("hello now")

untrace(print)
print("hi")

trace(three, print("entering 3"))
```# }}}



#### E6 recover()
```
```

#### examine class=rlang_error; code to generate error()# {{{
```{r eval=FALSE	 }
generate an error;
STUDY!
```

```{r}
## dummy function
fun <- function(x, error = FALSE) {
  x <- x + 1
  if(error)
    abort("Error!", x = x)
  return(x)
}

## no error
fun(1, error = FALSE)

## error 
fun(2, error = TRUE)

## display last error (list)
last_error()

## Examine this error
l  <- last_error()
str(l)

# message, trace, parent, x, rlang & inheritance
str(l, max.level=1)

attributes(l)
l$trace

class(l)

```# }}}


