---
title:  basic t-test
---



1030_basic_t_test.R

Base:: plots
- set the params in pdf(...)
- plot
- dev.off() to complete
```{r plots, eval=FALSE }

pdf("~/Downloads/print_and_delete/04_lm_residuals.pdf")
plot(x, r, col = 2, ylim = c(-5, 5))
abline(h = 0)
dev.off()
```


Create fake random normal data about 2.0
- if "TRUE mean" is 0, why is the t-value?
```{r t_test}

N  <-  100
fake_mean = 2.0
fake  <- rnorm(n = N, mean = fake_mean, sd= 1 )
fake

x  <- rep(0, times=N)
x

file  <- "1030_basic_t_test"
pdf(paste0("~/Downloads/print_and_delete/" , file, ".pdf"))
plot(x, fake, ylim = c(-5, 5), sub = "t-value = 21.5", type = "p")
abline(h = 0)
abline(h = fake_mean, col=2)
dev.off()

# ratio of "error" (difference of expected) to "width" (standard error)
t_value  <- (mean(fake) - 0)/ (sd(fake)/sqrt(N))
t_value # 21.5  large!

```


Recent NY Times article claimed the number of non-US born "Genius" awards (as
defined by ...... Foundation) greatly exceeds the expected number, based on
proportion of US population that is non-US born.   (The award is only open to
US residents or citizens.)

 Using the information in the article, I want to get check this (avoiding
 introducing any political arguments).


 The proportion of non-US born people in the US is 13 %, from 2010 US census.  
 This is population proportion or p0 = 0.13.


 Out of N=965 "genius" awards given since inception of the foundation,  209 or
 21.7% were born outside the US.   So p_hat = 0.217


 Is this significant?


 Hypothesis is H0 that p_hat = p0   vs   H1: p_hat != p0

 For a proportion,  Standard Error is given by 
 SE = sqrt (p0(1-p0)/n)

 and the Z-test is   (p_hat -  p0) / SE
```{r}

N <- 965
p0 <- 0.13
p_hat <- 209/N  # 0.217

SE <- sqrt(p0*(1-p0)/N)    # 0.011

#
t <- p_hat/SE  # 20.006

# So   reject H0,   so clearly can be RANDOM group.

pdf("~/Downloads/print_and_delete/1030_basic_t.pdf")
plot(x = c(0,0), y=c(p0, p_hat), xlim = c (-2,2), ylim= c(0, .5))
abline(h=p0)
abline(h=p_hat, col=2)
dev.off()

# Method 2
N <- 965
# but 
US <- 350 * 10^6
p0 <- N/US  # 2.77 * 10^-6

p_hat <- 209/(.13 * US) # 4.59 * 10^-6

d <- p_hat - p0  # 1.84 * 10^-6

SE <- sqrt(p0*(1-p0)/N)  # 5.35 * 10 ^-5

t <- d/SE  # 0.034
#####  Binomial #####

# pmf - binomial, x=success, size=throws

dbinom(x=0,size=2, prob=0.5) # n=2, x=0  prob = 0.25
dbinom(x=1, size =2 , prob=0.5) # n=2, x=1 # 0.5
sum(dbinom(x=0:2,size=2, prob=0.5)) # SUM = 1

```
