---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
    fontsize: 11pt   
    geometry: margin=0.5in,top=0.25in   
---


1030_basic_t_test.R

Base:: plots
- set the params in pdf(...)
- plot
- dev.off() to complete


## Use curve to plot t distribution
     curve(expr, from = NULL, to = NULL, n = 101, add = FALSE,
           type = "l", xname = "x", xlab = xname, ylab = NULL,
           log = NULL, xlim = NULL, ...)
```{r plot_t_dist}
x  <- function(x) dt(x, df=25)
y  <- function(x) dt(x + 0.2, df=25) # shift "mean"

curve(x, -3, 3, ylab="dt(df=25)", col="blue", )
curve(y, -3, 3,  col="red", add = TRUE)
# knitr::knit_exit()
```

```{r plots, eval=FALSE }
pdf("~/Downloads/print_and_delete/04_lm_residuals.pdf")
plot(x, r, col = 2, ylim = c(-5, 5))
abline(h = 0)
dev.off()
```


Create fake random normal data 
*  Ask:
*  H0   mean =0
*  H1   mean != 0

```{r t_test}
set.seed(4643)
N  <- 25 
fake_mean = 0.0
sd  <- 1
fake1  <- rnorm(n = N, mean = fake_mean, sd= 1 )
fake2  <- rnorm(n = N, mean = fake_mean, sd= 1 )

m  <- c(mean(fake1),mean(fake2))
m
s  <- c(sd(fake1), sd(fake2))

se  <- s/sqrt(N)
t  <- (m-c(fake_mean, fake_mean))/se
t
```

```{r plot}
plot(x = jitter(rep(c(0,1), each = N)), c(fake1, fake2), 
     xlim = c(-1,2), xlab="", ylab="", axes=FALSE )


box()
axis(side=1, at=0:1, labels=c("fake1", "fake2"))
axis(side=2, at=seq(-3,3, 1))
points(0, m, pch=15, col="red")
# arrows(x0=c(0,1), y0=m-se, y1=m+se, code=3, angle=90, length=.15)kj
# box()

knitr::knit_exit()
```
```{r next}

if (FALSE ){
axis(side=1, at=0:1, labels=c("nonsmokers", "smokers"))
axis(side=2, at=seq(14,22,2))
points(c(0,1), m, pch=15, col="red")
arrows(x0=c(0,1), y0=m-se, y1=m+se, code=3, angle=90, length=.15)
}
if (FALSE ){
pdf(paste0("~/downloads/print_and_delete/" , file, ".pdf"))
plot(x, fake, ylim = c(-5, 5), sub = "t-value = 21.5", type = "p")
abline(h = 0)
abline(h = fake_mean, col=2)
dev.off()
}

# ratio of "error" (difference of expected) to "width" (standard error)
t_value  <- (mean(fake) - 0)/ (sd(fake)/sqrt(n))
t_value # 21.5  large!

```


Recent NY Times article claimed the number of non-US born "Genius" awards (as
defined by ...... Foundation) greatly exceeds the expected number, based on
proportion of US population that is non-US born.   (The award is only open to
US residents or citizens.)

 Using the information in the article, I want to get check this (avoiding
 introducing any political arguments).


 The proportion of non-US born people in the US is 13 %, from 2010 US census.  
 This is population proportion or p0 = 0.13.


 Out of N=965 "genius" awards given since inception of the foundation,  209 or
 21.7% were born outside the US.   So p_hat = 0.217


 Is this significant?


 Hypothesis is H0 that p_hat = p0   vs   H1: p_hat != p0

 For a proportion,  Standard Error is given by 
 SE = sqrt (p0(1-p0)/n)

 and the Z-test is   (p_hat -  p0) / SE
```{r}

N <- 965
p0 <- 0.13
p_hat <- 209/N  # 0.217

SE <- sqrt(p0*(1-p0)/N)    # 0.011

#
t <- p_hat/SE  # 20.006

# So   reject H0,   so clearly can be RANDOM group.

pdf("~/Downloads/print_and_delete/1030_basic_t.pdf")
plot(x = c(0,0), y=c(p0, p_hat), xlim = c (-2,2), ylim= c(0, .5))
abline(h=p0)
abline(h=p_hat, col=2)
dev.off()

# Method 2
N <- 965
# but 
US <- 350 * 10^6
p0 <- N/US  # 2.77 * 10^-6

p_hat <- 209/(.13 * US) # 4.59 * 10^-6

d <- p_hat - p0  # 1.84 * 10^-6

SE <- sqrt(p0*(1-p0)/N)  # 5.35 * 10 ^-5

t <- d/SE  # 0.034
#####  Binomial #####

# pmf - binomial, x=success, size=throws

dbinom(x=0,size=2, prob=0.5) # n=2, x=0  prob = 0.25
dbinom(x=1, size =2 , prob=0.5) # n=2, x=1 # 0.5
sum(dbinom(x=0:2,size=2, prob=0.5)) # SUM = 1

```

```{r render, eval=FALSE    }

file  <- here("rmd", "1030_basic_t_test.Rmd")

# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
