---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
    fontsize: 11pt   
    geometry: margin=0.5in,top=0.25in   
---



### Note:  
Base:: plots  
- With knitr:  nothing special needed  
- For standalone example (see bottom)  


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE   ) # FALSE  easier to read
```
### Use curve() to plot t distribution
*  curve(expr, from = NULL, to = NULL, n = 101, add = FALSE,
           type = "l", xname = "x", xlab = xname, ylab = NULL,
           log = NULL, xlim = NULL, ...)
*  N=25 for most of this file, df = 24 (N-1)
```{r plot_t_dist}
N  <- 25
df  <- N-1
x  <- function(x) dt(x, df=24)
y  <- function(x) dt(x + 0.2, df=24) # shift "mean"

curve(x, -3, 3, ylab="dt(df=24)", col="blue", )
curve(y, -3, 3,  col="red", add = TRUE)
```

### Use  base::plot to :
*   plot a number of t values  
*   indicate (in red) a few quantiles for t
```{r annotate_t_plot}
x  <- seq(-3, +3, 0.1)
y  <- dt(x, df=df)

base::plot(x, y,
     xlim = c(-3,3), xlab="t", ylab="density", axes=FALSE )

box()
axis(side=1, at=seq(-3,3, 1))
#  need to ADD quantiles:   95%,  90%,  32%, 50%
t_values  <- qt(p=c(0.025, 0.05, 0.16, .50, 0.84, 0.95, 0.975 ), df=df,
                lower.tail=TRUE)
t_values
y  <- rep(c(0), length(t_values))
y
points(x =t_values, y=y, pch=15, col="red")

```

### Create sample(s) random normal data  
$$\mu=0$$
$$\sigma=1$$

*  Then use t-test to evaluate:  
*  H0   mean =0   
*  H1   mean != 0   



```{r samples}
set.seed(4643)
N  <- 25 
pop_mean = 0.0
pop_sd  <- 1
sample1  <- rnorm(n = N, mean = pop_mean, sd= pop_sd )
sample2  <- rnorm(n = N, mean = pop_mean, sd= pop_sd )

# knitr::knit_exit()
```
### Plot samples
```{r plot_sample}

m  <- c(mean(sample1),mean(sample2))
s  <- c(sd(sample1), sd(sample2))
se  <- s/sqrt(N)
x  <- rep(c(0,1), each = N)
plot(x = jitter(x), y = c(sample1, sample2), 
     xlim = c(-1,2), xlab="sample", ylab="value", axes=FALSE )

box()
axis(side=1, at=0:1, labels=c("sample1", "sample2"))
axis(side=2, at=seq(-3,3, 1))

# mean, 1 sd
points(x =c(0,1), m, pch=15, col="red")
arrows(x0=c(0,1), y0=m-se, y1=m+se, code=3, angle=90, length=.15)
```


### Calc se, t-value, 95% CI 
```{r t-test}

# first repeat above graph
plot(x = jitter(x), y = c(sample1, sample2), 
     xlim = c(-1,2), xlab="sample", ylab="value", axes=FALSE )

box()
axis(side=1, at=0:1, labels=c("sample1", "sample2"))
axis(side=2, at=seq(-3,3, 1))

# mean, 1 sd
points(x =c(0,1), m, pch=15, col="red")
arrows(x0=c(0,1), y0=m-se, y1=m+se, code=3, angle=90, length=.15)
# end of repeat graph


m
s
se  <- s/sqrt(N)
se

# t is actually the 'random variable', and should fit T distribution.
t  <- (m - 0)/se
t
t_95  <- qt(p=c(0.025, 0.975), df=df)
t_95

upper  <- m+ t_95*se
lower  <- m-t_95*se
lower[1]
upper[1]


# 95% CI (in green)
points(x = c(0,1), upper, pch=15, col="green")
points(x = c(0,1), lower, pch=15, col="green")
```

### Put mean and CI on density graph
```{r compare}

N  <- 25
df  <- N-1
x  <- function(x) dt(x, df=24)

curve(x, -3, 3, ylab="dt(df=24)", col="blue", )
points(x = c(lower[1], upper[1]),y=c(0,0),  pch=15, col="green")
points(x = c(t_95[1], t_95[2]),y=c(0,0),  pch=15, col="red")

knitr::knit_exit()
```

## t.test()
Compare to my calc (above)
     t.test(x, y = NULL,
            alternative = c("two.sided", "less", "greater"),
            mu = 0, paired = FALSE, var.equal = FALSE,
            conf.level = 0.95, ...)
```{r  t_test2}
t.test(x = sample1)
```


```{r }

knitr::knit_exit()
```
  
```{r next}

if (FALSE ){
axis(side=1, at=0:1, labels=c("nonsmokers", "smokers"))
axis(side=2, at=seq(14,22,2))
points(c(0,1), m, pch=15, col="red")
arrows(x0=c(0,1), y0=m-se, y1=m+se, code=3, angle=90, length=.15)
}
if (FALSE ){
pdf(paste0("~/downloads/print_and_delete/" , file, ".pdf"))
plot(x, fake, ylim = c(-5, 5), sub = "t-value = 21.5", type = "p")
abline(h = 0)
abline(h = fake_mean, col=2)
dev.off()
}

# ratio of "error" (difference of expected) to "width" (standard error)
t_value  <- (mean(fake) - 0)/ (sd(fake)/sqrt(n))
t_value # 21.5  large!

```


Recent NY Times article claimed the number of non-US born "Genius" awards (as
defined by ...... Foundation) greatly exceeds the expected number, based on
proportion of US population that is non-US born.   (The award is only open to
US residents or citizens.)

 Using the information in the article, I want to get check this (avoiding
 introducing any political arguments).


 The proportion of non-US born people in the US is 13 %, from 2010 US census.  
 This is population proportion or p0 = 0.13.


 Out of N=965 "genius" awards given since inception of the foundation,  209 or
 21.7% were born outside the US.   So p_hat = 0.217


 Is this significant?


 Hypothesis is H0 that p_hat = p0   vs   H1: p_hat != p0

 For a proportion,  Standard Error is given by 
 SE = sqrt (p0(1-p0)/n)

 and the Z-test is   (p_hat -  p0) / SE
```{r}

N <- 965
p0 <- 0.13
p_hat <- 209/N  # 0.217

SE <- sqrt(p0*(1-p0)/N)    # 0.011

#
t <- p_hat/SE  # 20.006

# So   reject H0,   so clearly can be RANDOM group.

pdf("~/Downloads/print_and_delete/1030_basic_t.pdf")
plot(x = c(0,0), y=c(p0, p_hat), xlim = c (-2,2), ylim= c(0, .5))
abline(h=p0)
abline(h=p_hat, col=2)
dev.off()

# Method 2
N <- 965
# but 
US <- 350 * 10^6
p0 <- N/US  # 2.77 * 10^-6

p_hat <- 209/(.13 * US) # 4.59 * 10^-6

d <- p_hat - p0  # 1.84 * 10^-6

SE <- sqrt(p0*(1-p0)/N)  # 5.35 * 10 ^-5

t <- d/SE  # 0.034
#####  Binomial #####

# pmf - binomial, x=success, size=throws

dbinom(x=0,size=2, prob=0.5) # n=2, x=0  prob = 0.25
dbinom(x=1, size =2 , prob=0.5) # n=2, x=1 # 0.5
sum(dbinom(x=0:2,size=2, prob=0.5)) # SUM = 1

```
## Examples  standalone plots
```{r plots, eval=FALSE, include=FALSE   }
pdf("~/Downloads/print_and_delete/04_lm_residuals.pdf")
plot(x, r, col = 2, ylim = c(-5, 5))
abline(h = 0)
dev.off()
```

```{r render, eval=FALSE    }
file  <- "1030_basic_t_test.Rmd" 
# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
