
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

REF:   See Rlang Cheatsheet
REF:   Advr v2  Chaptr 17+


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools) 
``` 

expr()
```{r begin}
# a, b are undefined
# expr() simply stores expression for future evaluation
e   <- expr(a + b)
```


evaluate e
```{r evaluat}
a  <- 1 
b  <- 2
eval(e)

a  <- 10
b   <- 11
eval(e)
```

rlang::enexpr
```{r enexpr}
# PROBLEM:  inside a function, expr
# returns x, regardless of what user sent
f  <- function(x) { expr(x)}

f(3)  # returns x


# fix, use enexpr
g  <- function(x) {enexpr(x)}

g(3)  # returns 3
g(a+b) # a+b, even though not defined.
a  <- 4
g(a)   # still get a
```
a few properties of expr()
```{r expr_are_lists}
f  <- expr(f(x=1, y=2))
f
typeof(f) # language

# add element
f$z  <- 3
names(f)

f[[1]]   # name of the call
f[[2]]

```

love the ast tree
```{r tree}
lobstr::ast(f("a","b"))
lobstr::ast(a*2+b)
lobstr::ast(b+ a*2)
```
17.4 create a funcction call
```{r call2}

rlang::call2("f",1,2,3)
rlang::call2(f,1,2,3)

```
shortcut, !!a (use expression a)
```{r bang-bang}
a  <- expr(x+y)
b   <- expr(x-y)

# Build a new expression from 2 simpler expressions.
expr(sin(!!a))
call2("sin", a)

expr(sin(!!a/!!b))
# getting longer ....
call2("sin",call2("/", a, b))

expr(!!a/!!b)
expr(!!a + !!b)
```

example, build new expression, from simpler ones.
```{r exampl}
# calc sd()/mean()

cv  <- function(x) {
  var <-  enexpr(x) # capture user's input
  expr(sd(!!var)/mean(!!var))
}
cv(x + y)
```



expressions
```{r b}
expr(log(e))    # log(e)
expr(log(!!e))  # log(a + b)


expr(log(!!a + b))   # log(10 + b)
expr(log(!!(a+b)))   # log(21)

# use with care...
log(!!(a+b))   #   0
eval(log(!!(a+b)))  #0
eval(3)							# 3
```

use eval to get number, result
```{r get-result}

log(eval(e))   # 1.1


# error:
log(eval(!!e))
eval(log(e))
eval(log(!!e))
```


k



***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
