
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  match.call, do.call, useMethod, match.args
---

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(jimTools) 
```


### match.call() 
* use inside a function.
* returns list.
* [1] is name of function.
* [2:n] are UNevaluated args.  (ie symbols are ok)
```{r match.call}
g  <- function(x, y, z){
  return(as.list(match.call()))
}

# Examples
g(1,2,3)
g(a,b,c)

# Study
l  <- g(a,b,c)
l
str(l)

# Example
l  <- g(1+3, b, c+0)
l
str(l)


### match.call , in real life
  *  use case:
  *  do.call(what, args), where `what` is function, args is list
  *  here:  match.call()[-1] puts args in a list

```{r begin}
f <- function(x, y, z) {
  do.call("sum", as.list(match.call()[-1])) 
}

f(1, 2, 3)
```

### match.arg, 
  *   limit possible input values

```{r match.arg}

f  <- function(type=c("a", "b")){
  type  <- match.arg(type)}

(f("b"))  # chr[]   "b"
(f("c")) #Error, not "a" or "b"
```








### Unicode:
  *   How to use unicode!
```{r random}
letters <- c(1, 2, 3, 4)
names(letters) <- c("a", "b", "c", "\U1F409")
letters
```


\newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
