
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  match.call, do.call, useMethod, match.args
---

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
library(jimTools) 
```


#### substitute, inside function retrieve value user sent
  *  deparse returns char[] 
```{r substitute}

# =========================
#deparse, substitute KEEP
# =========================
f  <- function(x)	substitute(x)
g   <-function(x) deparse(substitute(x)) 

f(unknown) # returns unquoted
# unknown
g(unknown) # returns in quotes, as char[1]
# [1] "unknown"


                      
f(a + b + c)
g(a + b + c)

# quite differnt
typeof(f(a + b + c))
typeof(g(a + b + c))
```

```{r quote_eval}
## setup
  x <-6
  y  <- quote(x)


## examine
  y
  eval(y)
```



```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
