  vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  advanced, functionals
---

<!-- 
vim: to format all urls for md
s/https.*/[&]()/g
-->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
```


```{r library, include=FALSE		}
library(jimTools)
  

#### Didn't know FUN(...) can be interpreted as ..1 ..2 etc

```{r dots}
f  <- function( ...) {..1 + ..2 + ..3 + ..4}
f(1,2,3,4)
# [1] 10
```

```{r }
##  IDEA is to pick correct fix_function based on param's name.
## Then apply the function to param's value.
## Repeat for all the parameters.

# fake data
raw_params  <- list(pm1 = "A", pm2 ="B" , pm3 ="C")

#   GOAL:     input_params:   list (pm1 = "1A", pm2 = "2B", pm3 = "3C")

# helps to separate into two vectors, param names & values
names(raw_params )  # char
values  <- unlist(raw_params)  
values


# match each param to fix function
fix_funs  <-  list( pm1 = function(x) paste0("1",x),
                    pm2 = function(x) paste0("2",x),
                    pm3 = function(x) paste0("3",x)
                    )
fix_funs


# given param, return its fix function
pick_function  <- function(name = x) {
  fix_funs[[name]]
}


# TWO-steps, get right fix function (based on param name), then apply to value 
input_params  <- map2(values, names(raw_params),
                     ~ pick_function(.y)(.x)) 
# matches GOAL
input_params

```


```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file <- ""
file  <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
