
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 

--- 
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
    fontsize: 11pt   
    geometry: margin=0.5in,top=0.25in   
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw

vim: to format all urls for md
s/https.*/[&]()/g
-->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "      ##",
  error = TRUE,
  collapse = FALSE
) # easier to read
```


```{r library, include=FALSE		}
file <- knitr::current_input()

library(jimTools)
library(forcats)
``` 


```{r begin}
print(file)
```

from Grolemund - categorical variables
Ref: https://rstudio-education.github.io/hopr/r-objects.html Ch 5.5.2

>To make a factor, pass an atomic vector into the factor function. R will
>recode the data in the vector as integers and store the results in an integer
>vector. R will also add a levels attribute to the integer, which contains a
>set of labels for displaying the factor values, and a class attribute, which
>contains the class factor:
```{r factors, collapse=TRUE}
# recode chr[] as integer
gender <- factor(c("male", "female", "female", "male"))

typeof(gender)
attributes(gender) # levels, class

# internally, stored as integer
unclass(gender) # as  integers, with level attr


# R uses the levels attribute when it displays the factor, as you will see.

# displays as levels
gender
```


#### 0050_factors (r4ds: Chapter 15, forcats)
#### factor: Peng, think ints with each having a label
```{r factors2, collapse=TRUE}
# create factors
levels <- c("a", "b")

# "c" becomes "NA"  NO ERROR
x <- c("a", "b", "b", "b", "c")

x <- c("a", "b", "b", "b") # chr[]


# factor: closer look
# store as integer, use levels for display
y <- factor(x, levels = levels)
y

typeof(y)
attributes(y) # levels, class

# also

levels(y)
class(y)
str(y)

# nice features
table(y)
names(y)
labels(y)

unclass(y) # underneath, stored as ints
names(unclass(y))
labels(unclass(y))
```

####10_base sort
```{r base}
# create data another way
x <- rep(c("a", "b"), 6) # chr[]
x

# turn into factora
levels <- c("a", "b")
y <- factor(x, levels = levels)
y


sort(y) # sort the underlying ints, then display with levels
```
https://forcats.tidyverse.org/index.html
#### forcats::fct_*
```{r forcats1}
# create random data from 0:9
x <- 0:9
x

# Create sample of 100 by:
# Choose 1 value from x, then replace it.
# Repeat 100 times

set.seed(2021)
y <- sample(x, size = 100, replace = TRUE)
y # int[100]


# only change is create factor
z <- factor(y, levels = 0:9)
z

# nice freq table  5 is most common (17)
table(z)

####
# Data internally or display, does not change, added class and levels
# Only change:  order of factors
####

# order factors by freq
forcats::fct_infreq(z)
# keep largest 3, lump rest into "other"
forcats::fct_lump(z, n = 3)
forcats::fct_lump(z, n = 3, other_level = "Z")



# order factors by 1st appearance 6 5 ...
forcats::fct_inorder(z)

# shuffle, again no change to underlying data or display, only factors
forcats::fct_shuffle(z)

# again
a <- forcats::fct_shuffle(z)
a
```

```{r forcats2 }
f <- factor(c("a", "b", "c"))
f

fct_rev(f) #> [1] a b c
g <- fct_rev(f)
g
levels(g)

fct_inseq(g)
```
## Begin examples from FORCATS:: cheat sheet
```{r begin_cheat}
library(forcats)
f <- base::factor(c("a", "c", "b", "a"),
  levels = c("a", "b", "c")
)
f

levels(f)

unclass(f) # structure
```

```{r fct_count}
fct_count(f)
fct_count(f, sort = FALSE)
```

```{r fct_unique}
fct_unique(f)
```
Concatenate data
```{r combine}
f1 <- factor(c("a", "c"))
f1
f2 <- factor(c("b", "a"))
f2

fct_c(f1, f2)
```

manually, reorder factors
```{r fct_relevel}
fct_relevel(f, c("b", "c", "a"))
```

highest is first
```{r fct_infreq}
f
fct_infreq(f)

f3 <- factor(c("c", "c", "a"))
f3
fct_infreq(f3)
```

set factors in order of appearance in data
```{r fct_inorder}
f2
fct_inorder(f2)
```

fct_reorder
```{r fct_reorder}
str(iris) # Species is categorical (factor)

# initial ordering
levels(iris$Species)
head(iris$Species, n = 20L)

pdf("~/Downloads/print_and_delete/forcats.pdf")
# plot(x, r, col = 2, ylim = c(-5, 5))
# abline(h = 0)
boxplot(data = iris, Sepal.Width ~ Species)
# boxplot(data = iris, Sepal.Width ~ fct_reorder(Species, Sepal.Width))
dev.off()
```
## Redo boxplot but  in asc or desc order
```{r reorder}
# .f = factor
# for each .f, group .x (a variable) and apply .fun
# sort based on .fun(.x) in asc order
# fct_reorder(.f, .x, .fun=median, .desc = FALSE)

f <- iris$Species # a factor
x <- iris$Sepal.Width
library(dplyr)

x <- fct_reorder(iris$Species, iris$Sepal.Width)
levels(x)
pdf("~/Downloads/print_and_delete/forcats.pdf")
boxplot(data = iris, Sepal.Width ~ fct_reorder(Species, Sepal.Width))
dev.off()
```

  


***

```{r knit_exit(), include = FALSE  , eval=FALSE } 
# knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

file <- here("", "")
file

# in general, pdf will look nicer
rmarkdown::render(file,
  # output_format = "pdf_document",
  output_format = "html_document",
  output_file = "~/Downloads/print_and_delete/out"
)
```
