---
title:  "00097_basic_trace_debug.Rmd"
output: 
  pdf_document:
    toc: TRUE 
    toc_depth: 1
    latex_engine: xelatex
fontsize: 10pt
geometry: margin=0.5in,top=0.25in
TAGS:  traceback, debug
---
#### REF:   Gillespie, Appendix E, PENG R programming, Ch18 
#### basic traceback, debug
```
 Tools:
	base::browser()
	base::debug(FUN)
	base::undebug(FUN)
	base::trace(FUN, exit=<action>)
	rlang::trace_back()
	options(error= browser, trace_back etc)
```


####	traceback()  WHERE is error
```{r traceback}
#	example:  x,y are undefined || error and warnings:
lm(y ~ x)

#	run immediately, error occured 7 levels deep
traceback()


##	another example, no function bug()
```{r 1 }
# there is no bug()
{
	first <- function() second()
		second <- function() third()
			third <- function() fourth()
				fourth <- function() fifth()
					fifth <- function() {
						bug()
					}

}

#	error
	first()

#	at level 5, function fifth()
	traceback()

```

####	interactive debugger aka browser in R 
```{r debug }
##	function to trigger debugger (ie browser opens) 
debug(lm)

lm(y ~ x)
undebug(lm)
```

####	debugger, 2nd example
```{r debug2}

f  <- function(){
	g  <- function() y  <- 10
	x  <- 5
	g()
}
debug(f)
f()
undebug(f)

}
debug(f)
first()
```

####	USE variable .DEBUG to indicate debug mode is on/off

			REF:  !so 6286092
```{r}

##	NEED: get/set function, plus function to check state
#   
# A slightly fancier version of Dirk's answer:


is_debug_mode <- function()
{
  exists(".DEBUG", envir = globalenv()) && 
    get(".DEBUG", envir = globalenv())
}

set_debug_mode <- function(on = FALSE)
{
  old_value <- is_debug_mode()
  .DEBUG <<- on
  invisible(old_value)
}

# Usage is, e.g.,

if(is_debug_mode())
{
  #do some logging or whatever
}

# and

set_debug_mode(TRUE)   #turn debug mode on
set_debug_mode(FALSE)  #turn debug mode off

is_debug_mode()

```
#### no bug
```{r 1 }
# there is no bug()
{
	print("jim")
	first <- function() second()
		second <- function() third()
			third <- function() fourth()
				fourth <- function() fifth()
					fifth <- function() {
						bug()
					}

}
#

tracingState() # this a Global value, either on or off

# browser
# Must insert browser() into code.

# on error, use traceback()
first()
traceback()
lobstr::cst()

# use debug
debug(fifth)
first()
undebug(fifth)

#use trace
trace(fourth, browser)
first()
untrace(fourth)

#trace, another way
trace("third", print("inside 3th"))
first()
untrace(c("third", "fourth"))
first() # runs to bug()


#trace, using recover()
trace(third, recover())


rlang::last_error()
rlang::last_trace()
```

#### E2, browser(), pauses R, view in function's env


#### E4, debug(FUN), acts as browser() when entering FUN# 
```
debug(FUN)  - debug a FUN
undebug(FUN) - turn off
debugonce(FUN) - debug a FUN and turn if off
```

```{r E4 }
debug(second)
isdebugged(second)
isdebugged(third)

first()
undebug(second)
```# 

#### E5, trace(FUN,browse, at=)# 
```
trace inserts a command in FUN, at line #
untrace(FUN) to end
```

```{r E5}
trace(print)
print("hello")

trace(print, browser)
print("hello now")

untrace(print)
print("hi")

trace(three, print("entering 3"))
```# 



#### E6 recover()
```
```

#### examine class=rlang_error; code to generate error()# 
```{r eval=FALSE	 }
generate an error;
STUDY!
```

```{r}
## dummy function
fun <- function(x, error = FALSE) {
  x <- x + 1
  if(error)
    abort("Error!", x = x)
  return(x)
}

## no error
fun(1, error = FALSE)

## error 
fun(2, error = TRUE)

## display last error (list)
last_error()

## Examine this error
l  <- last_error()
str(l)

# message, trace, parent, x, rlang & inheritance
str(l, max.level=1)

attributes(l)
l$trace

class(l)

```# 


