<!--
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
-->	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

```{r setup, include=FALSE		}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "      ##",
  error = TRUE,
  collapse = TRUE
)
```

```{r library, include=FALSE		}
library(jimTools)
``` 

### When a function returns ....
  *  can return an object.
  *  may *also* have side effect, such as print to console.
  *  many base R functions do both.
  *  invisible() supresses print to console.
     This function can be useful when it is desired to have functions
     return values which **can be assigned,** but which **do not print** when they
     are not assigned.
  *  invisible() has another differnce from return().  return() halts
     execution, returns *control*.   invisible() does not return *control*.
  *  Good practice when using return is return(invisible(x)) so only the
     object is returned.
```{r invisible}
print(x = 3)

{
  # print returns 3, also prints 3 ...this can lead to duplication
  y <- print(x = 3)
  y
}

f <- function(x) {
  return(x)
}

# not assigned, but prints
f(2)

# assigned normally, does not print
y <- f(2)
y


# SAME, but do NOT send to console
g <- function(x) {
  invisible(x)
}

# not assigned, no print
g(3)

y <- g(3)
y
```
