
---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
TAGS:  typeof,collection, symbol, base types,
---


## COLLECION of R base \code{types}

### Try to collect info re:  base R types,  that is all.
### R basic types
  *  REF:  [hadley Chapter 12]( https://adv-r.hadley.nz/base-types.html )

#### (from help) unlikely to be seen at user level 

  *    symbol, 
  *    pairlist, 
  *    promise,
  *    language, 
  *    char, 
  *    ..., 
  *    any, 
  *    expression, 
  *    externalptr,
  *    bytecode 
  *    weakref  

(Stole from maya?)
::: WIP, construct an assorted list of `typeof` objects :::

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read

```


```{r typeof} 

rm(x)
z  <- 1
ex  <- list( `[`, `[[`, `c`, `+`, sum, c(1,2), c(TRUE,FALSE), 
            as.name('x'), quote(x), quote(x+1), quote(z) ,
            function(x) {x+1},  "abcd", environment())

g  <- function(x) { 
  # If expression is unevaluated, Deparse returns char[1] (string)
  paste(deparse(x), "=", typeof(x))
}

purrr::map(ex, g)

lapply(ex, g)
```

### language
```{r language}
w  <- quote(x+y)

typeof(w)
# [1] "language"


typeof(unlist(w))
# [1] "language"
```



### symbol    
REF:  https://stackoverflow.com/a/20764194/5057251
```{r symbol}
    x=as.name('y')
# y

x
    # y

if (F) y   # error

mode(x)
  # [1] "name"

  class(x)
# [1] "name"

  typeof(x)
# [1] "symbol"

  storage.mode(x)
# [1] "symbol"

```

### Pairlist  
  *  Ref Adv-R Chapter 18.6  
  *  Lists know length;  pairlist only knows link to next (tranverse to get
    length)  
  *  However, pairlists are easy to append/remove without reallocating space.
    (Heavily used internallly)  
```{r setup, include=FALSE		}
f  <- function(x) x^2

body(f)
# x^2

typeof(body(f))
# [1] "language"

formals(f)
# $x
# 
typeof(formals(f))
# [1] "pairlist"

```


```{r library 		}
# print all labels
file  <- knitr::current_input() 
print(knitr::all_labels())
library(jimTools)
``` 


```{r skeleton}
file="/home/jim/.config/nvim/templates/skeleton.Rmd"
```

```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

## In doubt?   USE  knitr and do not waste time!

## Want to embed Latex, stick to pdf output (html?   never sure!)

# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# NOTE:    md_document is considered HTML, and latex commands may fail.
# in general, pdf will look nicer


## GOOD PRACTICE:
#  Refer to files relative to project root, which should remain as working dir.
#  So why using `here` ?

{
file  <- "" 
file  <- basename(file)
file  <- here("rmd", file)
file
}


## Want github to display nicely rendered output?
##  *  include md_format
#   *  include output_dir = "out"
#   *  if .gitignore balks,  then add !out/**   to close .github

rmarkdown::render(file,
    #output_format = "pdf_document",
    #output_format = "html_document",
    output_format=c("html_document", "md_document"),
    #output_dir = "out",
    output_file = "out")

# ==========================================
#  Example 1:  pdf| same name as source file | in wd |
# ==========================================

rmarkdown::render(file,
    output_format = "pdf_document")

# ============================================================
#  Example 2:  pdf | same name as source file | in pdf/|
# ============================================================
# pdf/  will be created

if (knitr::is_latex_output())  (
rmarkdown::render(file
    output_format = c("pdf_document") ,
    output_dir = "pdf")
)

if (knitr::is_html_output()) (

rmarkdown::render(file,
    output_format = c("html_document", "md_document"), 
    output_dir = "html")

)
```
