	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

SCRATCH

If it is good code,  move it!
OTHERWISE ... it is junk

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)

library(jimTools)

```

Do NOT UNDERSTAND this
```{r}
x  <- 1
e  <- new.env()
e$x  <- 2
e$g  <- function() x + 1

# interesting!
e$g()   # 2, not 3

# however, these make sense
e$x     #2
ls(e)
get("x", envir=e)  # 2
```


makes sense
```{r}
x  <- 1
g  <- function() {
  x  <- 2
  x + 1
}
g()   # 3

```


```{r enclosing}
e  <- new.env()
e$g  <- function(x) x + 1
e$x  <- 2

environment(g)   # error, no g
environment(e$g)  # global 
environment(e)   # NULL

# all errors
if (FALSE ){
  g(x)
  g(1)
  g()
  e$g()
  e$g(x)
}

e$g(5)   # 6
e$g(e$x) #3


environment(e$g)  <- e
environment(e$g)   # 0x59dd....  (not global)
e$g(e$x)  # 3
e$g(5)   # 6

```



get(x, envir=environment(pos), pos= -1 ...)
```{r}
x  <- 2
get("x") # req'd to be string #2

# in calling env of  get
get("x", envir=as.environment(-1)) #2


f  <- function() {
  x  <- 4
  get("x", envir=as.environment(-1)) 
}
f()  # 4

f  <- function() {
  x  <- 4
  get("x", envir=globalenv())
}
f()   # 2

f2  <- function() {
  get("x", envir=as.environment(-1)) 
}
f2()  # 2

```


Change f's environment
```{r}
x  <- "hello"
f  <- function() x
f()  # "hello"


# create practice env, e
e  <- new.env()
e$x  <- "bye"

environment(f)  <- e
f()  # "bye"
```

change a function env
```





Does what?
```{r begin}

get_environment  <- function() {

# each row is char[1]	, 82 x 1
# 
t <- tibble::tibble(environment = Sys.getenv())

# z  <- tibble::tibble(env=names(t$environment),
#                   values = t[[1]])

}
```

***

```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file <- ""
file  <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
