	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: `knitr::content_input()` 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)


```

Environment consists of 
  - frame:  set of named objects
  - pointer:  to enclosing frame.


Visually, environemnts appear tree-like
TODO:  insert latex


<e> notation for environment

Scope: Rules R uses to find objects by name
  - R first looks in runtime <e>
  - Then to parent.env()

```{r env}
e  <- environment()
parent.env(e)
```

Two environment equal if:
```{r identical}
e  <- environment()
f   <- environment()
identical(e, f)
```

Special Environments
```{r special}
emptyenv()
baseenv()
globalenv()

search()
```

See objects in Environment
```{r ls}
ls(envir = environment())
ls(envir = emptyenv())
head(ls(envir = baseenv()), n = 10L)
ls(envir = parent.env(environment() ))
```




```{r knit_exit()} 
# knitr::knit_exit()
```


```{r render, eval=FALSE	} 

# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root
dir  <- "env_code"
file  <- here(dir, "")

file
"

rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
