/home/jim/code/try_things_here/TIDYVERSE/300_pluck.qmd

TAGS: pluck, purrr 

PURPOSE:    
USE pluck notation.   This is also used in tidyr::hoist and unnesting list-columns.
REF:	https://purrr.tidyverse.org/reference/pluck.html

pluck(.x, ..., .default=NULL)

NOTE:

When using BASE R `[[`,  dput makes much more readable:

TODO:
2023-06-06
-   compare pluck and map (below)

```{r}
library(purrr)
library(tidyr)
```

### use dput to make nested lists easier to read
```{r}
?`[[`

obj1 <- list("a", list(1, elt = "foo"))
obj2 <- list("b", list(2, elt = "bar"))
x <- list(obj1, obj2)

#   compare 

x[[2]][[2]][2]
# $elt
# [1] "bar"
dput( x[[2]][[2]][2])
# list(elt = "bar")

```

### Pluck [REF](https://purrr.tidyverse.org/reference/pluck.html)     
```{r review_purrr_pluck}
# [REF](https://purrr.tidyverse.org/reference/pluck.html)     

obj1 <- list("a", list(1, elt = "foo"))
obj2 <- list("b", list(2, elt = "bar"))
x <- list(obj1, obj2)
dput(x)
# list(
    list("a", list(1, elt = "foo")), 
    list("b", list(2, elt = "bar"))
   )

##  Two identical ways to pull 1 element from list
##  I prefer dput notation!
dput(purrr::pluck(x,1))
dput(x[[1]])
identical(dput(x[[1]]), dput(purrr::pluck(x, 1)))   #T

# 
# 
##  Examples
pluck(x,1,2)   # from 1st element, pluck 2nd item (a list)
dput(pluck(x,1,2))
# list(1, elt = "foo")

pluck(x,1,2,"elt") # [1] "foo"
identical(pluck(x,1,2,"elt"), pluck(x,1,2,2)) #  [1] TRUE
```

```{r}
# compare pluck to map, find 2nd element of each
map(x, 2)
map(x, list(2, "elt"))  # just walk the hierarchy

# mix and match!  (just walk ....)
# Recall:   map(.x, .f, ...),  where ... are args for .f
map(x, pluck, 2, "elt")
```

### Accssor (return certain element of list)
```{r accessor_function}
# pluck() also supports accessor functions:
my_element <- function(x) x[[2]]$elt

pluck(x, 1, my_element)	# grab 1st elment, then apply my_element to it
pluck(x, 2, my_element)

# compare to:     cool!
map(x, my_element)
base::Map(my_element, x)
#base::Map(x, my_element)  NO
```

```{r pluck_my_example}
l1  <- list(a="a", b="b", c="c")
l2  <- list(a=1, b=2, c=3)
l3  <- list(a="joe", b="john", c="moe")
l  <- list(l1=l1, l2=l2, l3=l3)

# crawls the hierachy ....
pluck(l, 1) # first 1st element (a list)
pluck(l, 2 )
pluck(l, 1, 2) # second element of first item
pluck(l, 3, "c")
# pluck(l,c(1,2,3), "c")

```
```{r}

# GIVEN named list of strings, flatten list into single string
format_single_string  <- function(l) {
	map2_chr(names(l), l, ~ paste0(.x ,": ",pluck(.y), "\t\t")  )
}


# EXAMPLE: 
l  <- list(one="a", two="b", three="c")
single  <- format_single_string(l)
```

### example from httr2, youtube-api, simplify list 
```{r}
L=list(id="123", list(title="abc", description="describe"))
L
L1 = unlist(L, recursive=F, use.names=T)   # list, but flattened
dput(L1)
# list(id = "123", title = "abc", description = "describe")
```

### want:   c("id", "title", "descrption")
```{r}

L=list(id="123", list(title="abc", description="describe"))
pluck(L, \(x) "id")
pluck(L, 2, \(x) x[["title"]])
pluck(L, 2,  \(x) x[["description"]])

pluck(L, \(x) c( x[["id"]], x[[2]][["title"]]))
# [1] "123" "abc"




## no need pluck !
L1 = unlist(L, recursive=F, use.names=T)   # list, but flattened
dput(L1)

```

