---
title: "079_lists_tidyr_purrr_unpack_nested_lists.Rmd"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## Study github with gh:: package, uses github API
* REF:  https://github.com/r-lib/gh
* gh() returns JSON in R format, ie lists of lists
* First, find endpoint (url for search you want)

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools) 
library(gh)
library(magrittr)
library(purrr)
library(dplyr)
library(tibble)
```

basic 
```{r simple}
users <- gh("/users", .limit = 50)  # 50 elements in root; each has 18 elem

users
length(users)

gh("/licenses")

gh("/users/jimrothstein")
gh("/users/jimrothstein/repos")
# Useful!
gh("/rate_limit")
```
basic  GET /repos/owner/repo/pulls
```{r ex1 }
# GITHUB_PAT is set in ~/.Renviron
# to retrieve it:
gh::gh_token()
Sys.getenv("GITHUB_PAT")

owner <- "hadley"
repo <- "r-pkgs"
pr_list <-
  gh("/repos/:owner/:repo/pulls", owner = owner, repo = repo,
     .token=Sys.getenv("GITHUB_PAT"),
     state = "all", .limit = Inf)
length(pr_list)   #503
```
GET /users/owner/repos
```{r query}
# This is endpoint:  https://api.github.com/repos

# limit to 10 records
# q is search query

res  <- gh("/users/jimrothstein/repos")
gh("/users/jimrothstein/repos", q="function")

# stolen from:  https://github.com/xvrdm/ghhelper/blob/master/R/backend.R

gh_response_to_tibble <- function(gh_res) {
  tibble::tibble(data_col = unclass(gh_res)) %>%
    tidyr::unnest_wider(col = "data_col")
}
gh_response_to_tibble(res)  # 26 x73
```
GET /repos/owner/repo/issues
```{r issues }

# This is endpoint:  https://api.github.com/repos
gh("/repos/jimrothstein/yt_api/issues")
x  <- gh("/repos/jimrothstein/try_things_here/issues")


#######   sep post processing
str(x, max.level=1)  # list of 9 (issues)

l  <- x %>% pluck(1)  # 1 of n(issues)
length(l)  #list of 25
l

type  <- map_chr(l, ~typeof(.))
type
length(type)

t  <- tibble(name=names(type),
             type=type)

t  <- t %>% mutate(value=unlist(l))
t

```

GET /search/code
```{r search}
# This is endpoint:  https://api.github.com/search/code
# limit to 10 records
# q is search query

# response has passed through fromJSO
res  <- gh("/search/code", 
   q="dplyr+language:r")

res$total_count
length(res$items) #30

length(res$items[1]) #1
res$items[[1]]   # 1 of 30
res$items[[1]]$name



```

GET /issues  is for ASSIGNED USER
```{r get_issues}
q=paste0("tibble+language:R", "+labels:'help wanted'")
q

# NULL
res  <- gh("/issues", 
   q="tibble")
res
```


GET search/issues
```{r}

res  <- gh("/search/issues", 
   q="language:'R'+labels:'help wanted'", 
   .limit=10)

res  <- gh("/search/issues", 
   q="language:R+labels:'help wanted'", 
   .limit=10)

# POST
# ==================================
str(res, max.level=1)  # list of 3

res[1]  # total count 
res[2]  # incommplete?
str(res[3], max.level=1) # list of 1 with 30 items

# first of 30 lists
x  <- purrr::pluck(res, 3, 20)
names(x )
x[["comments_url"]]
x[["title"]]
x[["comments"]]
x[["body"]]
unlist(x)
```


FAILS:
GET /search/labels
```{r labels}
# must give specific repo?
# documentation says NO quantifiers in q ???
res  <- gh("/search/labels", q="help wanted")

```


  
GET /search/issues
ignoring language:R ???
```{r issues}
# response has passed through fromJSO

res  <- gh("/search/issues", 
   q="language:'R'+labels:'help wanted'", 
   limit=10)

str(res, max.level=1)  # list of 3

res[1]  # total count 
res[2]  # incommplete?
str(res[3], max.level=1) # list of 1 with 30 items
res[[3]]
rres[[3]][[1]]  # 1st fo 30
res[[3]][[1]]$body  # 1st fo 30


# first of 30 lists
x  <- purrr::pluck(res, 3, 10)
names(x )
x[["comments_url"]]
x[["title"]]
x[["comments"]]
x[["body"]]
unlist(x)
```


```{r errors}
# Error
gh("GET /users/{username}/repos", username= "jimrothstein")
my_repos <- gh("GET /users/{username}/repos", username = "gaborcsardi")
```

```{r my_repos}
owner <- "jimrothstein"
# repo <- "r-pkgs"

# error
pr_list <- gh("/repos/:owner", owner = owner, 
     .token=Sys.getenv("GITHUB_PAT"),
     state = "all", .limit = Inf)
length(pr_list)   #503
```

GET /users/user/repos
```{r}

x  <- gh("/users/:username/repos", username = "jimrothstein")
typeof(x)
str(x, max.level=1, vec.len=1  )

purrr::pluck(x, 1)  # first in list  of 26
l  <- purrr::pluck(x, 1) # list of 74
names(l)


jsonlite converts to dataframe
```{r json}
jim_repos  <- jsonlite::fromJSON("https://api.github.com/users/jimrothstein/repos")
jim_repos
t  <- tibble::as_tibble(jim_repos)
dim(t)  # 26 x 73
head(t)
str(jim_repos, max.level=1, vec.len=1)
```

GET /search/code
```{r}

res  <- gh("/search/code", 
   q="dplyr+language:r+labels:'help wanted'", 
   limit=10)

# 1st of 30
res[[3]][[2]]$name
res[[3]][[2]]$repository$description
res[[3]][[2]]$repository$html_url

head(as_tibble(res[[3]][[2]]$repository))

jsonlite::fromJSON("https://api.github.com/search/code?q=tibble")

```

GET /repos/user/repo/issue
```{r json_search}

jim_issues  <- jsonlite::fromJSON(
         "https://api.github.com/repos/jimrothstein/try_things_here/issues?q=tibble")
t  <- as_tibble(jim_issues)
t
head(t[11], n=1L)
names(t)
# one row
t[1,]

t[1,][11]

pluck(t[1,], 11)
length(pluck(t[1,], 11)) # 18

head(t)   # 6 x 25
names(t)
```
gh::has_name()
```{r has_name}
# component by component
# for list, only 1 level
has_name  <- function(x) {
  nms  <- names(x)
  if (is.null(nms)){
    rep_len(FALSE, length(x))
  } else {
    !( is.na(nms) | nms=="")
}
  }


has_name("")
has_name(c(a=1, b=2))
has_name(c(1, b=2))
has_name(list(1,2))
has_name(list(a=1,2))
has_name(list("", b="", c=3))
has_name(list(NULL, b="", 3))
has_name(list(a=1, 2, c=list(1,2)))
has_name(x=NULL)

```

gh::has_no_names()
```{r has_no_names}
has_no_names  <- function(x) all(!has_name(x))
has_no_names( c(a=1, 2))
has_no_names( c(1, 2))
```
vapply(X, FUN, FUN.VALUE)
```{r vapply}
x  <- list("a", 2, TRUE )
vapply(x, is.character, logical(1))
vapply(x, is.character, logical(1))
```

#### 14_tidyr - easier
```{r}

# hoist(<tibble>, <list-column>, <specifc name>) -> <tibble> with <list_c

repos <- tibble(repo = gh_repos)

# each user has up to 30 repos, we want list of all the repos
unnest_longer(repos, repo)

repos %>% unnest_longer(repo) %>% 
  hoist(repo, 
  login = c("owner", "login"), 
  name = "name",
  homepage = "homepage",
  watchers = "watchers_count")

repos %>% 
  hoist(repo, owner = "owner")
```

```{r}
knitr::knit_exit()

```
