---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  3   
fontsize: 10pt   
geometry: margin=0.4in,top=0.25in   
TAGS:  conditions,errors,simpleCondition, simpleError,handler
---


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read
library(jimTools)
library(rlang)
```

## Notes on R `condition system`
  *  overall mechanism to handle conditions (aka: exceptions, errors, warnings)
  *  **code** signals (or throws) a condition.
  *  `condition system` matches condition and invokes the best handler function.


**Condition** 

  *  S3 object, constructor is simpleCondition()
    minimal contains 2 char[1] (strings): "message" and "call"
  *  message is for user, call refers to function where condition occurred.  
  *  Hadley prefers to not include call.
  *  Hierarchy:  condition; subclasses:  simpleCondition, exception, warning
    ... Each can subclassed, such as simpleException.
  *  Built-in simple*() constructors

  *  More details in code.

**Handler**   (aka error handling)

  *  a regular R function, with at least one arg for condition object.
  *  Two flavors:
      *  Exiting handler (mostly for errors) does not allow code to continue.
      *  Calling handler (mostly for non-errors) handles (aka modify or
        swallow) the condition and allows code to continue, just a regular
        function returns control.




### simpleCondition:  built-in subclass of condition object
  *  not much in book.
```{r simpleCondition}
formals(simpleCondition)
args(simpleCondition)

# create one
s  <- base::simpleCondition("jim")

# examine s
s
str(s, max.depth=1)
class(s)
inherits(s, "condition")  # T
typeof(s)  # boring list
conditionMessage(s) # jim
conditionCall(s)   # call where occured?

# simpleError 
e  <- simpleError("jim1")
e
str(e, max.depth=1)

# see how it work, first base:
tryCatch( {stop("jim1")
          print("never gets here")
    },

         error = function(e) e )

#  This creates an error, but does not signal (or throw it) 
tryCatch( {simpleError("jim1")
          print("never gets here")
    },

         error = function(e) e )

#  Ahh, bbut this thows error, then catches, and never return to print line.
tryCatch( {signalCondition(errorCondition("jim1", call="I made this up"))
          print("never gets here")
    },

         error = function(e) e )

# Examine the object returned.
r  <- tryCatch( {signalCondition(errorCondition("jim1", call="I made this up"))
          print("never gets here")
    },

         error = function(e) e )

# we have the condition, it is r
str(r)
inherits(r, "condition")  # T
conditionCall(r)
```

### So far, focused on simpleCondition and none of other classes.   Before
proceeding, now examine handlers
```{r handlers}
```


```{r knit_exit(), include = FALSE  , eval=FALSE } 
# knitr::knit_exit()
```




```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

## In doubt?   USE  knitr and do not waste time!

## Want to embed Latex, stick to pdf output (html?   never sure!)

# NOTE:   .tex only works with PDF
# NOTE:    tex will NOT work with html
# NOTE:    md_document is considered HTML, and latex commands may fail.
# in general, pdf will look nicer


## GOOD PRACTICE:
#  Refer to files relative to project root, which should remain as working dir.
#  So why using `here` ?

{
file  <- "008_basic_conditions.Rmd" 
file  <- basename(file)
file  <- here("R", file)   # in this repo using `R`, not `rmd`
file
}


## Want github to display nicely rendered output?
##  *  include md_format
#   *  include output_dir = "out"
#   *  if .gitignore balks,  then add !out/**   to close .github

rmarkdown::render(file,
    #output_format = "pdf_document",
    #output_format = "html_document",
    output_format=c("html_document", "md_document"),
    #output_dir = "out",
    output_file = "out")

# ==========================================
#  Example 1:  pdf| same name as source file | in wd |
# ==========================================

rmarkdown::render(file,
    output_format = "pdf_document")

# ============================================================
#  Example 2:  pdf | same name as source file | in pdf/|
# ============================================================
# pdf/  will be created

if (knitr::is_latex_output())  (
rmarkdown::render(file
    output_format = c("pdf_document") ,
    output_dir = "pdf")
)

if (knitr::is_html_output()) (

rmarkdown::render(file,
    output_format = c("html_document", "md_document"), 
    output_dir = "html")

)
```
