
---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
TRY:   advanced-R,condition
---

###  Example of custom condition,
  *  Section    of Advanced R

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = T   ) # easier to read
library(rlang)
```


Here I try to unpack some of the contents of Hadley's section on custom
expections.   Nothing new here.  But sometimes I find it helpful to work through
simpler code before using tidyverse, rlang etc features.

### Simple
  *  Here I force an error under certain conditions.   Then handle it.
  *  The handler for error is anonymous function.   Note that e$message or
    base::conditionMessage(e) result in same thing.
```{r simple}

tryCatch(
  {
    a <- 0
    if (a == 0) abort(message = "Dummy")
  },
  error = function(e) {
    e$message
  }
)
```

### To break this down:
  *  tryCatch mechanism
  *  need to create a  condition (error or warning) , then signal (or throw) the condition
  *  Catch and handle the condition.
  *  In case above, abort() combines the signal and the creation of the
    condition.

To examine things, it will be a little easier to wrap the try portion of
tryCatch in a function

```{r study}


tryCatch(
  {
    a <- 0
    if (a == 0) abort(message = "Dummy")
  },
  error = function(e) {
    e$message
  }
)

get_val.1  <- function() {
  a = 0
  if (a == 0 ) abort(message="Dummy")

}
# with no tryCatch  Error
get_val.1()


tryCatch(
  {
    get_val.1()
  },
  error = function(e) {
    e$message
  }
)

# When get_val.1() runs, it is same as above. 
#
#  abort does at least 2 things:   create the custom condition, identified with
#  this class.
#  Abort continues to throw (or signal) this condition


```{r custom error}

get_val.2  <- function() {
  a = 0
  if (a == 0 ) abort(message="Dummy", 
                     class="get_value_ERROR", 
                     val = a)
}

get_val.2()

tryCatch(
  {
    get_val.2()
  },
  error = function(e) e$message
)

# now catch the specific errror
tryCatch(
  {
    get_val.2()
  },
  get_value_ERROR = function(e) print("hi"),
  error = function(e) e$message
)

```

#### Create the Condition
```{r}
# Use errorCondition allows custom class
my.error  <- errorCondition(message = "jim1", class="error.X", call="I made this up")

inherits(my.error, "condition")
inherits(my.error, "error")
inherits(my.error, "error.X")


tryCatch( {signalCondition(my.error)
          print("never gets here")
    },

          error.X = function(e) print("error.X"), 
         error = function(e) e )

```

#### Finally, the handler
```{r handler_error.X}
error.X_handler  <- function(e) print("error.X")

tryCatch( {signalCondition(my.error)
          print("never gets here")
    },

         error.X = error.X_handler ,
         error = function(e) e )
```


#### Hadley's code
```{r}
stop_custom <- function(.subclass, message, call = NULL, ...) {
  err <- structure(
    list(
      message = message,
      call = call,
      ...
    ),
    class = c(.subclass, "error", "condition")
  )
  stop(err)
}

err <- catch_cnd(
  stop_custom("error_new", "This is a custom error", x = 10)
)
class(err)
err$x
```
#### my changes to hadley's code
```{r my_changes}

# division_condition
division_condition  <- function(message="msg_division_condition",
                                class = "class_division_condition",
                                call = NULL,
                                ...) {
  condition <- structure(
                  list(
                      message = message,
                      call = call,
                      ...
                      ),
                  class = c("class_division_error", "error","errorConditon", "simpleCondition", "condition")
  )
}

err <- catch_cnd(
  signalCondition("error_new", "This is a custom error", x = 10)
)
class(err)
err$x

c  <- division_condition()
str(c)

# why NULL?  Doesn't signal throw it?
signalCondition(c)
stop(c)

tryCatch(signalCondition(division_condition()),
  class_division_condition  = function(e) print("class_division"),
  error  <- function(e) print("error"),
  msg_division_condition  <- function(e) print("msg_division"),
  )



```

```{r render, eval=FALSE, include=FALSE 	} 
# from ~/.Renviron
output_dir  <- Sys.getenv("OUTPUT_DIR")
output_dir
{
file  <- "/home/jim/code/book_advanced_R/R/008_custom_conditions.Rmd" 
file  <- basename(file)
file  <- here("R", file)
file
}


## Want github to display nicely rendered output?
##  *  include md_format
#   *  include output_dir = "out"
#   *  if .gitignore balks,  then add !out/**   to close .github

rmarkdown::render(file,
    output_dir = output_dir,
    output_format = "html_document")

)
```
