---
title: "100_using_match.fun().Rmd"
date:  
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true 
fontsize: 10pt
geometry: margin=0.5in,top=0.25in
#  md_document:
#  html_document:
#    toc: true
#  	toc_float: true
#    css: styles.css
---

###  USE of match.fun
  *  Why?  What is purpose?   Aids in selecting from a LIST of functions

Below is example of one use for match.fun().   My question is:
Are there OTHER situations where `match.fun()` is used and why?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  
                      comment="      ##",  
                      error=TRUE, 
                      collapse=TRUE)
library(magrittr)
library(lobstr)
library(jimTools)
``` 

See also:
From here(  https://stackoverflow.com/questions/6403852/how-to-call-a-function-using-the-character-string-of-the-function-name-in-r)




Here is ONE example where match.fun helps us code with `mean` and `"mean"` as functions:
  
```{r match}

# First, create list 
f_list  <- list("mean", mean)

# do a few checks
sapply(f_list, typeof)
sapply(f_list, is.function)


# Suppose we want to use first entry of f_list as a function 

#   This will fail:
f  <- f_list[[1]]
f
f(c(3,4))  # FAIL

# But this succeeds:
g  <- match.fun(f_list[[1]])
g
g(c(3,5))


f_list  <- list("*", `*` , `[`,  "sin", sin, "mean", mean)





```{r render, eval=FALSE, include=F	} 
setwd("~/code/book_advanced_R/")
getwd()
here()
file <- "book_advanced_R/R/100_functions_match.Rmd" 
file  <- basename(file)
file
dir <- "R"


rmarkdown::render(input = file,
                  output_format = "html_document",
                  output_file = "out"
                  )

```

