<!--
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
-->	 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

* Lots purrr, dplyr examples in PRIOR code.
* Don't see SIMPLE examples.
* X' = map(X, f) where X is a set

```{r setup, include=FALSE		}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "      ##",
  error = TRUE,
  collapse = TRUE
)
```

```{r library, include=FALSE		}
library(jimTools)
library(purrr)
library(dplyr)
``` 

purrr::rerun (replaces replicate())
```{r rerun}
# create a list, or repeat any expression
purrr::rerun(10, "hello")

rerun(3, date())

# this creates chr[]
purrr::map_chr(1:10, ~"hello")
```

rep
```{r rep}
# returns chr[]
base::rep("hello", 10)

# double[]
base::rep(c(1, 2), 4)

# create atomic vectors
rep(1 + 1, 3) # 2 2 2
rep(date(), 2) # chr[]
```

SIDE EFFECTS
* most functions return something (list ...)
* side effect means ENV changed, global, a file.  Usually not desired.
* Sometimes, ex print(), side effect is benefit.
```{r sideeffect}
l <- c("jim", "joe", "sally")
# SIDE EFFECT:  print each element of l
z <- map_chr(l, print)

# in addition, returns chr[] z
z
```
WALK
```{r walk}
# WALK,   calls sideeffect
#   AND returns .x
z <- purrr::walk(l, ~ print(length(.)))
z
```

# more advanced
```{r}
# SIDE EFFECT,  print goes to console (1st 10 lines)
# THEN print a list
map(rep("hello", 10), print)

# map returns LIST of just 1 elment, chr[]
map("hello", ~ rep(.x, 10))

# ERROR
if (FALSE) {
  map_chr("hello", ~ rep(.x, 10))
}
```

```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file <- ""
file <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file, dir)
jimTools::ren_github(file, dir)
```
