---
title: 
output: 
  pdf_document:
    latex_engine: xelatex
  md_document:
fontsize: 12pt
geometry: margin=0.5in,top=0.25in  
params:  
  d:  !r Sys.Date()
#  html_document:
#    toc: true
#  	toc_float: true
#    css: styles.css
---

**To Render** 

```{r render, eval=FALSE, echo=FALSE	, include=FALSE }		
library(here)
library(ggplot2)
library(scales)

ren <- function() {
  rmarkdown::render(here::here("book_advanced_R", "R", "010_functional_factories.Rmd"),
    output_dir = "~/Downloads/print_and_delete"
  )
}
ren()

ren2 <- function() {
  rmarkdown::render(here::here("book_advanced_R", "R", "010_functional_factories.Rmd"),
    output_format = "github_document",
    output_dir = "~/Downloads/print_and_delete"
  )
}
ren2()
```

```{r one, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE, comment = "      ##", error = TRUE, collapse = TRUE)

#
```

```
10.1, 
````

```{r 10.1_code}
power1 <- function(exp) {
  function(x) {
    x^exp
  }
}
square(2)
```

````
10.2.1 Environments
````
PATTERN:
OUTER FUN  (aka factory)

-- INNER FUN
-- aka manufactured function
-- it's enclosing env = execution env of outer FUN
-- normally an execution env is ephemeral, not in this case!

````


```{r 10.2.1_code}
# we are 'promise' state
# exp not evaluated, so where is it?

square
cube

# Observe, same parent env, but different enclosing env
env_print(square)
env_print(cube)

# Examine exp in each enclosing env
fn_env(square)$exp
fn_env(cube)$exp

# LOOK:  above must have forced eval!
x <- 3
square(2)
```

```` 
10.2.2  
Bug!  
lazy...searches for value of env
```

```{r 10.2.2_code}
x <- 2
square <- power1(x)

fn_env(square)$exp # this forces an evaluation!

x <- 3
square(2)
```

```
Solution: use force(exp), so will exp, b/c it will be returned
```

```{r}
power2 <- function(exp) {
  force(exp)
  function(x) {
    x^exp
  }
}

x <- 2
square <- power2(2)
square # code still shows "exp" but env now knows
fn_env(square)$exp
x <- 3
square(2)
```

```
10.2.4 Stateful
```
```{r 10.2.4_code}
set.seed(0602)
new_counter <- function() {
  i <- 0

  function() {
    i <<- i + 1
    i
  }
}

counter_one <- new_counter()
counter_one()
fn_env(counter_one)$i
env_print(fn_env(counter_one))

counter_two <- new_counter()
```

```{r 10.2.6_3}
rm(i)
rm(x)

pick <- function(i) {
  e <- current_env()
  cat(capture.output(e))
  l <- info()
  cat(capture.output(l$current),
    capture.output(l$caller),
    capture.output(l$caller_parent),
    sep = "\n"
  )

  force(i)
  one <- function(x) x[[i]]
}

# pick(i)  # Error, demands value for i

i <- 3
f <- pick(i)


# create
# code
f

# EXAMPLES
letters <- letters
f(letters) # c

l <- list("one", "two", "threee")
f(l) # threee
info()
```


\pagebreak
