---
title: "R4DS Notes From Advanced R"
output: md_document
#  html_document:
#    toc: true
#  	toc_float: true
#    css: styles.css
---

#### Experiment
Beginning with Chapter 4 use .Rmd to self-document output + any specific notes.
Save the .pdf for future self.
To review rmarkdown.
To make easy to comment on slack.
To do everything that "Gaya" does.

#### Workflow
Once code is working, RENDER to /home/jim/Documents/print_delete
Using:
:RSend
rmarkdown::render("book_QA.Rmd", output_format="pdf_document", output_dir="pdf")
Apparently, output_dir is flaky.

# Chapter 4 - Basics
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment="      ##",  error=TRUE, collapse=TRUE)

library(tidyverse)
library(kableExtra)
library(lobstr)
```
#### compare: paste0 has default sep=""
```{r}
paste0("hello","world","!")
paste("hello","world","!" )
```

# Chapter 4
::: Review basics :::
```{r}
x <- 1:10
x[]

x[0]
```

#### 4.2.6 #3 Matrix Subset
::: dim > 1, matrix IS vector with attributes, `[` returns either matrix or 1D :::
```{r}
m  <- matrix(1:16, byrow=T, nrow=4)
str(m)
is.vector(v) && is.matrix(m) && is.integer(m) && is.atomic(m)
attributes(m)

m[,] 
m[0,]
m[,0]

m[,-2]
m[0,-2]

m[2,]
m[2,,drop=F]
```

::: m[select] seems to saying, form (i,j) from row 1 of select; grap that
element from m; repeat with row 2...:::
```{r}
m  <- outer(1:5, 1:5, FUN=paste, sep=",")
m
select <- matrix(ncol=2, 
								 byrow=T, 
								 c(1,1, 3,1, 2,4)
)
select
m[select]
```
::: vector :::
```{r}
x  <- 1:10
x[]
x[0]
x[NA]
x[NULL]
```
::: matrix is a vector, with attributes :::
```{r}
m  <- matrix(1:16, byrow=T, nrow=4)
str(m)
dim(m)
attributes(m)
is.vector(m)
is.matrix(m) && is.integer(m) && is.atomic(m)
m[,] # everything
m[0,]# no rows, col headings only
m[,0]
m[,-2]
m[0,-2]
m[2,]
m[2,,drop=FALSE]
```

::: m[select] seems to saying, form (i,j) from row 1 of select; grap that element from m; repeat with row 2...:::

::: more matrix :::
```{r}

# FUN refers to binary operation used
outer(1:5, 1:5, FUN=paste, sep=",")
outer(1:5, 1:5, FUN="*")
outer(1:5, 1:5, FUN="+")

# now, subset m
# lose dimension!
m[c(1,1,1,1), drop=FALSE	]

z  <- matrix(c(1,1,1,1), nrow=2, byrow=TRUE)
z
m[z]
m[upper.tri(m)]
m[upper.tri(m) , drop=FALSE	 ]
```
::: Compare sizes:   1:10, rep(1:10, 10), 1:100 :::
reprex({code},foo=)
```{r}
reprex::reprex({
x1 <- 1:10
x2 <- rep(1:10,10)

lobstr::obj_size(x1)
lobstr::obj_size(x2)

lobstr::obj_sizes(x1,x2)
lobstr::sxp(x1, expand="altrep")

lobstr::sxp(x2, expand="altrep")

dput(x1)
dput(x2)

}, outfile="foo")
```
```{r}
x1  <- 1:10
tracemem(x1)
.Internal(inspect(x1))

x2  <- rep(1:10,10)

.Internal(inspect(x2))

x3  <- 1:100

obj_size(x1)
obj_size(x2)
obj_size(x3)

ref(x1,x2,x3)
obj_size(x1,x2)
obj_sizes(x1,x2)

```
#### Problem Sets
```{r}
# =================
# Quiz Question 1
# =================
```

```{r}
# positive ints
# ===============
<<<<<<< HEAD
=======
v  <- 1:8
>>>>>>> fix_advanced_R
v[1:3]
ref(v[1:3])

# negative ints
# ===============
v[-c(1:8)]
v[-1]
v[-1:-8]  # Works

v[c(1:3, 5:6)]
v[c(c(1:3), 5:6)] # SAME

# logical
# =========
set.seed(2020)
l <- sample(c(TRUE,FALSE	),size=26, replace=TRUE) # logical
l
v[l]  # returns only TRUE
v[!l]  # inverse
v[ l || !l] # back to original


# note:
v[c(TRUE, FALSE	)] #recyling (Hadley says not to)
v[-l]  # hmmm, drops only first
reverse  <- -l # ooops, converts to int[]
TRUE || TRUE

# character
# ===========
v <- letters

w  <- stats::setNames(v, paste("candy_",letters,sep=""))
names(w)
w[c("candy_a","candy_b")]

# note:
v["a"]  # NA
names(v) # no names
w[!c("candy_a")]  # error
v <- NULL

# misc
# ======
v[]
v[0]
<<<<<<< HEAD
```
=======
v[NULL] #error
v[NA]
v[rep(NA,8)]
v[c(NA)]

v[1] <- NULL
v[1] <- NA
v
```

>>>>>>> fix_advanced_R

#### 4.2.6 Exercises
# ====================

#### 4.2.6 #1
```{r}
# FIX
# mtcars[mtcars$cyl = 4, ],  NEED LOGICAL
mtcars[mtcars$cyl == 4, ]

# mtcars[-1:4, ],  DO NOT negative and postive indices
nrow(mtcars)	
nrow(mtcars[-(1:4), ])
nrow( mtcars[-c(1:4), ])

#mtcars[mtcars$cyl <= 5] # missing columns
mtcars[mtcars$cyl <= 5,]

# mtcars[mtcars$cyl == 4 | 6, ]  # see r4ds 5.2.2
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
mtcars[mtcars$cyl %in% c(4,6),]
dput(mtcars)
```
#### 4.2.6 #2
```{r}
# Fix
x  <- 1:5
x[NA]
x[c(NA,NA,NA,NA,NA)]

x[NA_integer_]
x[TRUE] # recycles
x[NULL]

# integer
typeof(x) # integer

typeof(NA) #logical
typeof(NA_integer_) # integer

```
```{r}

m[c(1,2,3),c(1,1,1)]
m[c(1,2,3),c(1,1,1), drop=FALSE]
# Returns matrix of logicals, same as size

m  <-  matrix(c(1,1,1,1), nrow=2)
m
upper.tri(m)
lower.tri(m)

```
:::  4.2.6 #4 :::
```{r}
m <- matrix(1:16,  nrow=4, byrow=TRUE)
m
diag(m) # returns vector


{ # works
	x	<- vector("double", length(1:4))
	for (i in 1:4) {x[i] <- m[i,i]}
	x
}

{
	# not work
	list(1:4)
	lapply(list(1:4), d)
	vapply(1:4, d)
}

# Works
f  <- function() {
	size = 250
	m  <- matrix(1:size^2, byrow = T, nrow=size)
	d	<- function(i){m[i,i]}
	purrr::map_dbl(1:nrow(m), d)
}

```

::: WIP, construct an assorted list of `typeof` objects :::
```{r}
ex  <- list(
  `[`,
  `[[`,
   `c`,
  `+`, 
  sum,
  c(1,2), 
  c(TRUE,FALSE), 
  function(x) {x+1},  
  "abcd", 
  environment()
)
ex

deparse(ex)

g  <- function(x) { 
  paste(deparse(x), "=", typeof(x))
}

purrr::map(ex, g)

```



# pick out diagonals
```{r}
m  <- matrix(1:25, byrow=T, ncol=5)
m

sapply(1:5,function(x) {m[x,x]})

```

