---
title:  "0700_why_closure_stopwatch.Rmd"
output: 
  pdf_document:
    toc: TRUE 
    toc_depth: 2
    latex_engine: xelatex
fontsize: 12pt
geometry: margin=0.5in,top=0.25in
---

####  Why Closure, StopWatch example:
  *  REF:  Gillespie, Lovelace   pg63
  *  One use:  closure (function + env) is way to CACHE variables.


When I first started hearing words `closure`, `environment` or even `lambda` I
startd getting confused.   Actually, the basic ideas can be grasped without
needed to understand the fancy words.   


All we need to know is that the variable `start_time` remains available to
self$stop, when ever this function is called, ie `start_time` is cached.

Normally, when a function finishes execution like `stopwatch` below all of its
variables disappear.   So if self$stop is run much later, it normally would not
have access to `start_time`.   


```{r stopwatch}
# cache start_time
stopwatch  <- function() {
    start_time  <- NULL
    current_time  <- NULL

    self  <- list()

    self$start  <- function() start_time  <<- Sys.time()

    self$current  <- function() difftime(Sys.time(), start_time)

    self$stop  <- function() difftime(Sys.time(), start_time)

    return(self)
}
```

```{r invoke}

        watch1  <- stopwatch()
        watch1$start()

            watch  <- stopwatch()
            watch$start()
            watch$current()
            watch$stop()


        watch1$stop()

```

Looking under the hood, there is a little more complexity that a programmer use
efficiently. That is where the fancy words come in. 

I found these couple of stackoverflow answers clarified many of the meanings.
Then things became much clearer.

  *    https://stackoverflow.com/a/36879264/5057251



Below, the variable start_time is cached for future use by the function
self$stop().   Normally, once a 
