---
title: 9009_namespace
format:
  html:
    embed-resources: true
    toc: true
execute:
  enabled: false   # do not execute chunks
---

Namespaces Collect examples here

::: very_small
very small?
:::

::: small
small?
:::

### Todo:
- Import (load only)
- Depends (loads & attaches, so use sparsely; like library)
- detach(pkg) - 1st step when unloadNS is called.
- Dreaming?
ls(R_BaseEnv)
R_BaseNamespace
namespace:R_BaseNamespace
  
namespace:base
package:base
- 
# NAMESPACES - See Ch 7.4.3

## REF

-   how R finds - https://blog.thatbuthow.com/how-r-searches-and-finds-stuff/
-   SEE R packages book
-   **Note** that lazy-loaded datasets are not in the package’s namespace so need to be accessed via ::, e.g. survival::survexp.us.
-   R-internals
-   **Note** never use @export ds; put in data/

## Loop through ds?

use `::`(pkg, ds)

```{r}

f = function(ds){ deparse(substitute(ds))
  }

```

## HELP

```{r}
apropos("environment")
apropos("namespace")
apropos("getNamespace")
apropos("namespace")

```

## Packages (Mostly)

### all installed packages

```{r}
installed.packages()
installed.packages() |> class()
ls()
```

### attached packages vs search()

```{r}
(.packages()) #22
search()  #24
z=paste0("package:", .packages())

z %in% search()
search()[!z %in% search()]

search()[!z %in% search()]
base::setdiff(z, search())
```

### is pkg loaded?

```{r}
"stats" %in% loadedNamespaces()  # TRUE
"admiral" %in% loadedNamespaces()   # T
"admiraldev" %in% loadedNamespaces()  # T
"broom" %in% loadedNamespaces()  # F

```

### package versions?

```{r}
sessionInfo()   # uses package version from disk, not memory
getNamespaceVersion("admiral")  # better
```

## getAnywhere (examine all loaded NS)

```{r}
?getAnywhere
argsAnywhere(derive_params)
getAnywhere(filter)
getAnywhere(ae)

# fct in pkg admiral
getAnywhere("count_vals")
getAnywhere(count_vals)

argsAnywhere(derive_params)
getAnywhere(filter)
getAnywhere(ae)

# fct in pkg admiral
getAnywhere("count_vals")
getAnywhere(count_vals)

# dataset in pkg (must be loaded)
library(pharmaversesdtm)
getAnywhere("ae")   # finds in package:pharamaversesdtm

# load = error
if (F) load("mtcars")  # errro

# but will not find if unloaded
unloadNamespace("pharmaversesdtm")
```

## Namespace (mostly)

### get pkg namespace

-   asNamespace , returns \<environtment: namespace:stats\>
-   as.environment, must include package: , returns \<environment: package:stats\> + attrs

```{r}
asNamespace("stats")  # do not prefix with "package:"

as.environment("package:stats")
```

### getNamespace, returns environment

```{r}
?help.start
library(pkgload)
getNamespace("admiral")   # Loads NS for admiral
isNamespaceLoaded("admiral")
unloadNamespace("admiral")

# fake notation, do not use
if (F) {
namespace:base
package:base
}

```

### objects in NameSpace

When lazyLoad on, datasets DO NOT appear in NS

```{r}
ls(getNamespace("dplyr"))  # works, getNameSPace returns environment

## Need better way ...
ls(getNamespace("admiral")) == "count_vals"  # T
getAnywhere("count_vals")

```

### is object in ns?

```{r}
z=ls(getNamespace("admiral"))

c("slice_derivation") %in% z   # TRUE
```

## unload NS

unloadNamespace- forces; first it detach(NS)

```{r}

?pkgload::unload   # NS, pkg

?unloadNamespace
unloadNamespace("stats")   # error, b/c needed by others


loadNamespace("admiral")
unloadNamespace("admiral")
"admiral" %in% loadedNamespaces()  # check

?pkgload::unload

```

## Datasets (mostly)

### Find datasets data(), class packageIRQ

```{r}
# returns obj of class packageIRQ
data() |> class()

# data in one package
data(package = "tidyr")

# data in one package (not loaded)
isNamespaceLoaded("broom")  # F
data(package = "broom")
isNamespaceLoaded("broom") # F
```

### Find all 'loaded' ds? means ...?

so far, only getAnywhere("ae") finds ds 'ae' in pkg (pharamaversesdtm)

```{r}
?getAnywhere
getAnywhere("mtcars")$where   # where won't print the ds
getAnywhere("ae")$where
getAnywhere("ae")$visible

```

### compare: data("ae") vs pharamaversdtm::ae

if pkg loaded and attached, easy to find "ae"

```{r}
search()
library("pharmaversesdtm")  # makes it easy
isNamespaceLoaded("pharmaversesdtm")
unloadNamespace("pharmaversesdtm")

ae

# is in environment
"ae" %in% ls(as.environment("package:pharmaversesdtm")) #T

"ae" %in% ls(as.environment("package:pharmaversesdtm"))  #T

## but if you unload
unloadNamespace("pharmaversesdtm")
search()
pharmaversesdtm::ae   # loads, not attach pkg;  still works
z = pharmaversesdtm::ae
z

ae   # error
"ae" %in% ls(as.environment("package:pharmaversesdtm"))  # error

```

## unload pkg

```{r}

```

### ds in pkg; NOT 'exported', put into data/

```{r}
data()   # all loaded ds
data(package = "admiral")  # ds in package admiral; will NOT load NS admiral
ls(getNamespace("admiral"))   # will load NS admiral, ~ 176 objects


isNamespaceLoaded("admiral")
unloadNamespace("admiral")
data("admiral_adsl")   # not found
data("ae")  # not found



# ae is dataset in pharmaversesdtm, in /data
library(pharmaversesdtm)  # must be loaded
ae   # found it
getAnywhere("ae")   # works

## pharamaversesdtm does NOT export much;  data/ is not called 'exported'
c("ae") %in% getNamespaceExports("pharmaversesdtm") # F

# but "ae" is in data/
str(data(package = "pharmaversesdtm"))
c("ae") %in% data(package = "pharmaversesdtm")$results

# because uses IQR format
z=data(package = "pharmaversesdtm")
z |> class()   # packageIQR
"ae" %in% z   # WHY FALSE?
"ae" %in% z$result   # NOW it is true!


ls(package:stats)
ls(package:admiral)   # admiral must be on search()


ls(ns:stats)
ls(namespace:stats)
ns:stats
namespace("stats")
ls(imports:stats)

z=getNamespaceImports("admiral")
class(z)
```

## attach ds to search path; to see all its columns

```{r}
search()
base::attach(mtcars)
search()
```

## probe package, imports, exports (seems to match file NAMESPACE)

```{r}
getNamespaceImports("stats")   # list
getNamespaceExports("stats")

getNamespaceImports("admiral")
getNamespaceExports("admiral")

```

## CASE: pharmaversesdtm

```{r}
sdtm = "pharmaversesdtm"   # I can't spell
library("pharmaversesdtm")

##  remove pkg and its NS ------------------------ 
unloadNamespace("pharmaversesdtm")
search()
isNamespaceLoaded("pharmaversesdtm")
ae  # not found
vs

##  ------------------------ 

##  loads , does not attach pkg ----------------------------
pharmaversesdtm::ae

# works
`::`(pharmaversesdtm, ae)
?`::`

# fails
`::`(pharmaversesdtm, ae, ts)
sapply(c("ae", "ts"), `::`pharmaversesdtm)

data(package = "pharmaversesdtm")  # no attach

##  ------------------------ 
##  fails, functional form?

f = function(x) {
  cat(enexpr(x))
  `::`(pharmaversesdtm,!!x)
  }

sapply(c("ae", "vs"), f)

f(expr(ae))
f("ae")

## call ------------------------ fails
call("`::`", "pharmaversesdtm", "ae")




getAnywhere("ae")$where
getAnywhere("ae")$visible  # T

search()
as.environment("package:pharmaversesdtm")   # package req'd
ls(as.environment("package:pharmaversesdtm"))  # lists "ae"
"ae" %in% ls(as.environment("package:pharmaversesdtm"))  #T


getNamespace("pharmaversesdtm")  # no "package:", returns envir)
ls(getNamespace("pharmaversesdtm")) # get_terms
ls(envir=getNamespace("pharmaversesdtm"))
get_terms


getNamespaceExports("pharmaversesdtm") # no exports  character(0)

# easy way to get NS
asNamespace("pharmaversesdtm")
as.env
# WHY FALSE?
"ae" %in% ls(getNamespace("pharmaversesdtm")) #F


```
