---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 4
        toc_float: true
        mathjax: default
  pdf_document:   
    latex_engine: lualatex  
    toc: true
    toc_depth:  4   
fontsize: 11pt   
geometry: margin=0.4in,top=0.25in   
TAGS:  frame
---
~/code/MASTER_INDEX.md

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = T   ) # T easier to read
```

```{r library, include=F 		}
#library(jimTools)
``` 

####    USE CASE:  Inside function f, label current environment and parent
environment

  * Return list with the objects in each frame.
```{r 001}
Z  <- 0
    f  <- function() {
        cat("In f execution frame \n")
        pframe  <- parent.frame()
        cframe  <- environment()
        Z  <- 10
        
        # default, ls uses calling envir 
        list(ls(),
             ls(name=cframe),
             ls(name=pframe))
        list(Z,
             get("Z", envir=cframe),
             get("Z", envir=pframe))
    }
f()

```


####    USE CASE:  Same as above, but show other ways to show pframe, cframe
identical. 
  * cframe = current or active frame
  * pframe = calling frame (in stack 1 frame up, parent)
  * eframe = enclosing frame, where function was first DEFINED.

```{r 002}
Z  <- 0
    f  <- function() {
        cat("In f execution frame \n")
        pframe  <- parent.frame()
        cframe  <- environment()
        cframe2  <- sys.frame(which= sys.parent(n=0))
        cframe3  <- sys.frame(sys.nframe())
        cat(identical(cframe,cframe2), "\n")
        cat(identical(cframe2,cframe3), "\n")
        Z  <- 10
        
        # default, ls uses calling envir 
        list(ls(),
             ls(name=cframe),
             ls(name=pframe))
        list(Z,
             get("Z", envir=cframe),
             get("Z", envir=pframe))
        
    }
f()
```

### 003 now interesting....
  * cframe = current or active frame
  * pframe = calling frame (in stack 1 frame up, parent)
  * eframe = enclosing frame, where function was first DEFINED.
```{r 003}

Z  <- 0
    f  <- function() {
        cat("In f execution frame \n")
        pframe  <- parent.frame()
        cframe  <- environment()
        eframe  <- environment(f)
        Z  <- 10
        
        # default, ls uses calling envir 
        list(ls(),
             ls(name=cframe),
             ls(name=pframe))
        list(Z,
             get("Z", envir=cframe),
             get("Z", envir=pframe),
             get("Z", envir=eframe))
    }

g  <- function() {
    Z  <- 100
    f()
}
g()

```


#### USE CASE:

```{r 004}
Z  <- "Z in global"
f = function() {
    Z = "inside f" 
    g = function() {
        Z = "inside g"
        pframe  <- parent.frame()
        cframe  <- environment()
        eframe  <- environment(f)
        grandframe  <- parent.frame(n=2)
        list(Z,
             get("Z", envir=cframe),
             get("Z", envir=pframe),
             get("Z", envir=eframe),
             get("Z", envir=globalenv()),
             get("Z", envir=grandframe))

             
    }
}

h  <- function() {
     Z  <- "Z inside h" 
     cat("In h execution frame \n")
     r  <- f()
     r()
}
h()
```

```{r render, eval=FALSE, include=FALSE 	} 
output_dir  <- Sys.getenv("OUTPUT_DIR")
output_dir

{
file <- "USE_CASES.Rmd"
## dislike `here`?
## get dir/file name separately, then combine
 the_dir  <- dirname(file)
 the_dir
 file  <- basename(file)
 file
 file  <- file.path(the_dir, file)
 file

#   file  <- here("rmd", file)
file
}

rmarkdown::render(file,
                  #output_format="html_document",
                  output_format="pdf_document",
                  output_dir=output_dir
                  )
```
