---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 3
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  3    
fontsize: 10pt   
geometry: margin=0.3in,top=0.25in   
---

###  This set of pages to include: 
  -  Vectors (atomic & lists)
  -  Simple vector examples
  -  Use of purrr .... and analogus base functions
  -  Use of lapply* family, unlist
  -  Notation for subset, contents etc
  -  Common patterns
  -  Vectorize a function, with mapper functions


```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = T   ) # easier to read
library(jimTools)
library(DiagrammeR)


```


### Vector hierachy
  *  REF:  r4ds 20.2, adv-r ch3

```{r graph}
DiagrammeR::grViz("
  digraph rmarkdown {
    vector -> atomic -> character
    atomic -> numeric
    atomic -> logical
    numeric -> double
    numeric -> integer
    integer -> factor
    double  -> date
    double -> POSIXct
    vector -> list
  }
  ", height=600)
```

Common functions of vectors
```{r commond_functions}
typeof(c("a")) # actually all objs
length(c("a"))

```
### Groupings (Define)
  *  Vector, Atomic, Numeric are groupings ("abstract", never created)
  *  Meanings of these functions, differ from intution:
     *  is.vector(), is.atomic(), is.numeric()

### Atomic
  * is.*
  * as.*
```{r atomic}
```
### NULL is its own ??
```{r null}
```

### Like other R objects, vectors can have Attributes
  *  dimensions
  *  class (S3, which allows generic functions to process by class attr)
  *  factor (int[], attr = levels)
  *  date (double[])
  *  df, tibble (list)
```{r}
f  <- as.factor(c("A", "B"))
attributes(f)
attributes(unclass(f))

x  <- Sys.Date()
x
attributes(x)
unclass(x)
attributes(unclass(x))

t  <- Sys.time()
attributes(t)
attributes(unclass(t))
```

### Create object with arbitrary attributes
```{r}

a  <- structure(.Data = 1:3, 
                x = "jim", 
                y = "blue")

attributes(a)

a  <- structure(.Data = 1:3, 
                class = factor)

attributes(a)


# internally, stored as int[]
b  <- factor(letters[1:3])

dput(b)
# structure(1:3, .Label = c("a", "b", "c"), class = "factor")

attributes(b)
unclass(b)


```

### using vector()
  *  `mode` (older, pre-OO) better to think mode is **base class** or list
  *  https://rdrr.io/r/base/vector.html
```{r vector}
vector(mode="logical", length=0)

# logical of length 
v  <- vector(mode="logical", length=0) 
v

# logical of length 1
v1   <- vector(mode="logical", length=1) 
v1

# logical of length 2
v2   <- vector(mode="logical", length=2) 
v2


# list of length 5, values are NUL
l1  <- vector(mode="list", length=5)
l1

```

### Subset, Extract
  *  [, [[, $

```{r subset}
a  <- c(1,2,3)
a[2]
a[a>2]
a[a%%2 == 0]

f  <- function(x) x ==1
a[f(a)]


# list
lapply (a, f)
lapply (a, `[`)

tryCatch({ 
         lapply(a, `[[` )},
          error = function(e) {"No good - error" }
)

```

#### list, can subset, append, walk  ... this is hard one.
```{r list}
l.1 <- list(letters, zz=4)
length(l.1)

l[length(l)]
l[1:3]


l.2  <- append(l.1, letters)
length(l.2)


l.2[2]
l.2[2:5]

# my favorite!
l.2[c(2,3,10, 26)]

l.2$zz
names(l.2)
```


```{r knit_exit(), include = FALSE  , eval=FALSE } 
# knitr::knit_exit()
```



```{r render, eval=FALSE, include=FALSE 	} 
{
file  <- "0010_base_vectors_intro.Rmd"
file  <- here("vector", file)
file
}

# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document")
                  output_format = "html_document")

```
