
```


```{r library, include=FALSE		}
get_typeof <- function(x) typeof(x)
``` 

#### expr()  `capture code`
  *  Confused:  relation to R expression?
  *  R expression is NOT a typeof
  *  Expression can use any of `call`, `symbol`, `constant`, `pairlist` 
  *  Each of these is `typeof`
```{r begin}
# a, b are undefined
# expr() simply stores expression (typeof is "language") for future evaluation
e <- expr(a + b)
# a + b
is.expression(e)
# [1] FALSE
typeof(e)
# [1] "language"
```


#### evaluate e 
```{r evaluat}
a <- 1
b <- 2
eval(e)
```

(a) # still get a
```
a few properties of expr()
```{r expr_are_lists}
f <- expr(f(x = 1, y = 2))
f
get_typeof(f) # language

# add element
f$z <- 3
names(f)

f[[1]] # name of the call
f[[2]]
```

love the ast tree
```{r tree}
lobstr::ast(f("a", "b"))
lobstr::ast(a * 2 + b)
lobstr::ast(b + a * 2)
```
17.4 create a funcction call
```{r call2}
rlang::call2("f", 1, 2, 3)
rlang::call2(f, 1, 2, 3)
```
shortcut, !!a (use expression a)
```{r bang-bang}
a <- expr(x + y)
b <- expr(x - y)

# Build a new expression from 2 simpler expressions.
expr(sin(!!a))
call2("sin", a)

expr(sin(!!a / !!b))
# getting longer ....
call2("sin", call2("/", a, b))

expr(!!a / !!b)
expr(!!a + !!b)
```

example, build new expression, from simpler ones.
```{r exampl}
# calc sd()/mean()

cv <- function(x) {
  var <- enexpr(x) # capture user's input
  expr(sd(!!var) / mean(!!var))
}
cv(x + y)
```



expressions
```{r b}
expr(log(e)) # log(e)
expr(log(!!e)) # log(a + b)


expr(log(!!a + b)) # log(10 + b)
expr(log(!!(a + b))) # log(21)

# use with care...
log(!!(a + b)) #   0
eval(log(!!(a + b))) # 0
eval(3) # 3
```

use eval to get number, result
```{r get-result}
log(eval(e)) # 1.1


# error:
log(eval(!!e))
eval(log(e))
eval(log(!!e))
```
more fun with expr() symbol
```{r symbol}
s <- expr(syb)
get_typeof(s) # symbol

s <- expr(x)
typeof(s) # symbol
typeof(expr(x)) # symbol

get_typeof(expr(5)) # double

get_typeof(d <- expr(5)) # double
get_typeof(d) # double

get_typeof(c <- expr("5")) # character

get_typeof(s <- expr(c(x, y))) # language
x <- 2
y <- 3
eval(s) # 2 3
```
```{r plurals}
z <- exprs(several, such, symbols)
get_typeof(z) # language
get_typeof(z[[1]]) # symbol
```

capture expr PLUS env
```{r quo}
quo(letters)
q <- quo(x)
x <- 1
eval(!!q)
```
create sym and call 
```{r call}
# create sym for y
expr(y)
sym("y")

c <- expr(mean(y))
c

get_typeof(c)
c <- expr(f(1, 2, 3))
c
get_typeof(c)

c <- call("mean", y, na.rm = TRUE)
get_typeof(c)
```

base:: 
```{r enquo}
code <- base::expression(a + b)
base::is.expression(code)
a <- 1
b <- 4
base::eval(code)

# same code,
e <- environment()
e$a <- 10
e$b <- 40
base::eval(code, envir = e)
```



```{r render, eval=FALSE	} 
file <- ""
file <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file, dir)
jimTools::ren_github(file, dir)
```
