---
title: "transform"
format: html
editor: visual
---

\footnotesize

###   ToDo
  -   Read Advanced R (first version) on NSE and how to use df in transform, or with, or ...
  -   Convert df to environment, why as.environment fails


You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```


base::transform
  -   allows unquoted operations on column in data.frame
```{r}
args(transform)
transform
transform.default
transform(mtcars, twice = hp * 2)
```

However, I am still stuck with how df can be  used instead of envir
same as with , examine first line of transform.data.frame
eval(expr, envir, enclos)
for expr, uses substitute(list(...))
for envir uses df
for enclo uses parent.frame
```{r}
# debug(f)
f <- function(`_data`, ...) {
  expr <- substitute(list(...))
  envir <- `_data`
  enclos <- parent.frame()
}
debug(f)
f(mtcars, twice = hp * 2)
# undebug(f)
```
##   Create Environemnt -- convert to data.frame
```{r}
e <- new.env(hash = TRUE, size = 3)
assign(x = "a", value = 10, envir = e)
assign(x = "b", value = 100, envir = e)
assign(x = "c", value = 1000, envir = e)
e
ls(e)
df <- data.frame(setNames(lapply(ls(e), get, envir = e), ls(e)))
```
use that data.frame
```{r}
transform(df, twice = 2 * b)
```

##      Reverse:   data.frame to environment
```{r}
```

