--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: lualatex
    toc: TRUE 
    toc_depth:  4 
    keep_tex:  true
#    hightlight:  tango
  html_document:
    toc: TRUE
    toc_depth:  4
  header-includes:  
    -  \usepackage{helvet}  
    -  \renewcommand{\familydefault}{\sfdefault}  
    -  \usepackage{ulem}
    -   \usepackage{xcolor}
# Add to header-includes: (for smaller kable tables)
#  \AtBeginEnvironment{longtable}{\tiny}
# mainfont:  Computer Modern Dunhill Roman  
fontsize: 10pt 
geometry: margin=0.4in,top=0.25in 
TAGS:  deparse, substitute
---
\footnotesize{this is footnotesize}

```{r setup, include=FALSE		}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "      ##",
  error = TRUE,
  collapse = T
) # T easier to read
```

```{r, results = 'asis',echo = T }
htmltools::h3("Test big font")
htmltools::h4("Test medium font")
htmltools::h5("Test smaller font")
```
`r htmltools::h3("Outside chunck")`

***

~/code/MASTER_INDEX.md

\textcolor{red}{red}

#### REF: 
```
https://statisticsglobe.com/parse-deparse-expression-r-function
Murdoch, RJournal, Dec 2010
https://journal.r-project.org/archive/2010/RJ-2010-010/index.html
https://stackoverflow.com/questions/46834655/whats-the-difference-between-substitute-and-quote-in-r
```


#### parse:  string  -->    expression
```{r parse}
# args(parse)
e <- parse(text = "2^2")
e

##  e is expression, which is based on list
is.expression(e)


## Rmk:  a is unknown, but fine for expression.
e2 <- parse(text = "2*a")

## Rmk2:  However, this fails because `2a` is not valid expression.

tryCatch(
  {
    e3 <- parse(text = "2a")
  },
  error = function(e) e
)


##  An expression can have attributes.
attributes(e)
```

````
####    expression, more details
```{r more_detail}`r ''`

length(e)

e[[1]]

## Using character vector, sapply will provide name 
    l  <- c("class", "typeof", "attributes","is.language", "is.expression", "is.pairlist")
    sapply(l, do.call, args=list(e))
    sapply(l, do.call, args=list(e[[1]]))

##  also works
{
    if (F) {
    l  <- list(class, typeof, attributes, is.pairlist)
    lapply(l, do.call, args=list(e))
    } 
}

```
````


####  parse, deparse  are not exact inverses.  Attributes are added.

```{r inverse}
e <- parse(text = "a   <-  sin(x)")


identical(e, deparse(e))

f <- function() {
  x
}
e <- parse(text = "f")



x <- quote(5 + 5)
x
is.call(x) # T
typeof(x) # language
is.pairlist(x) # F
as.list(x) # breaks the code


z <- quote(x + y)
is.call(z) # T


e <- quote({
  print(x)

  x + y

  x <- 3
})

as.list(e)
```

####    eval:  Takes a string, returns R expression
An expression is unevaluated and is not a character vector (string).
An error is returned if eval can not find a value.
eval does NOT apply quote.   So if x has a value, R considers x+1 to be double, not expression.
```{r eval}
args(eval)


x <- 10
eval(x)

##  e is not string
e <- expr(x)
is.character(e)
e
eval(e)

##
eval(expr(x))

##  An error
y <- 20
rm(y)
if (F) eval(expr(x + y))


##
y <- 12
eval(x + y)
eval(expr(x + y))

##  expr is cummulative
eval(expr(expr(x + y)))
eval(expr(expr(expr(x + y))))

##  observe,  a character vector makes no sense to evaluate, it just is what it
##  is. But quote(x) is an expression and is evaluated.
eval("x")
eval("5+5")
eval(quote(x))

##  observe quote delays lookup for x

eval(quote(x + 1), env(x = 1000))


##  Explain this: (hint:  eval takes an "expression")
eval(x, env(x = 1000))
eval(x + 1, env(x = 1000))
eval((x + 1), env(x = 1000))
eval(eval(x + 1), env(x = 1000))
eval(print(x + 1), env(x = 1000))
eval(eval(x + 1), env(x = 1000))

##
eval(expr(x + 1), env(x = 1000))
eval(quote(x + 1), env(x = 1000))
eval(quote(x + 1), env(x = 1000))
```


```{r knit_exit(), include=F, eval=F} 
knitr::knit_exit()
```

```{r render, eval=FALSE, include = F	} 
{
  file <- "/home/jim/code/try_things_here/BASE/0205_base_parse_expression.Rmd"
  file <- normalizePath(file)
  file
}

rmarkdown::render(file,
  output_format = "pdf_document",
  # output_format="html_document",
  output_dir = "~/Downloads/print_and_delete"
)
```
# z is 'expression'

added 5 MAY
```{r}
z
x <- 4
# now eval the expression
base::eval(z)
y
```

##    basic NSE
```{r}
a <- 2
b <- 10
z <- quote(a + b)
eval(z) # 12

z
(zz <- as.list(z))
```
# change
```{r}
zz[[1]] <- `*`
zz

## but zz is list
(as.call(zz))
eval(as.call(zz)) # 20
```
##    Quote can return.... (SEE Lionell)
https://rpubs.com/lionel-/programming-draft
