--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
    highlight: tango
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  match.call, do.call, useMethod, match.args
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

####    Quote
```{r quote_eval}
  x <-6
  y  <- quote(x)
# x
  quote(6)
# [1] 6

```

####    Eval is opposite of Quote
```{r eval}
  eval(y)
# [1] 6
```

####  Use case:   plot title
```{r}
a  <- 2
bquote(a == .(a))
plot(1:10, a*(1:10), main = bquote(a == .(a)))
```
####  Assignment, at run-time
```{r}
##  as an assignment
a = "A"
bquote('=' (.(as.name(a)), 4))
bquote('<-' (.(as.name(a)), 4))

# construct an assignment
x = 1000
bquote('<-' (.(as.name(a)), .(x)))

# eval or run the code!
eval(bquote('<-' (.(as.name(a)), .(x))))
A
# [1] 1000

```


####  Use case:  partial insert? 
```{r}
name = "jim"
bquote(name == .(name))
# name == "jim"

bquote(.(name))


paste0("hi, my name is ", bquote(.(name)))

if (F) print("hi")

# bquote will throw error, but do not understand why ignores `if`

if (F) {
  bquote(value is .(name))
}   # FAILS

```
bquote('=' (.(as.name(a)), 4))
####  Quote, repeated use adds "layers"
```{r properties}
x  <- 6

quote(x)
# x

quote(quote(x))
# quote(x)

##  quote adds "layers"
identical(quote(x), quote(quote(x)))
# [1] FALSE


##  eval each "layer"
eval(quote(x))
# [1] 6

eval(quote(quote(x)))
# x

eval(eval(quote(quote(x))))
# [1] 6

```

#### Referencial Transparency
```{r ref_transparent}

### Rmk:  Lack of referential transparency, is result for x and its value the
### same?
###
x  <- 6
quote(x)
# x

quote(x  <- 6)
# x <- 6

quote(6)
# [1] 6
```

#### Quote, to form unevaluated expression
```{r example3}
## capture unevaluated expression
     z = quote(mtcars |> subset(cyl == 4) |> nrow())

##  to evaluate it
    eval(z)
```
```
####    substitute(expr, envir or list)
        substitute does NOT evaluae any arguments
```

####   Create expression, using expression
```{r simple}

    x <- 1
    y <- 2
    a <- expression(10 + x + y)
    a
    is.expression(a)
    
    a[[1]]
    eval(a)

```
####    Substitute and then eval
```{r sub}

substitute(x+y, list(x=quote(1), y=quote(2)))
# 1 + 2


eval( substitute(x+y, list(x=quote(1), y=quote(2))))
# [1] 3

ans  <- substitute(x+y, list(x=quote(1), y=quote(z)))
# 1 + z

## why?
eval( substitute(ans, list(z=2)))
# 1 + z

eval(substitute(x+y, list(y=2)))
# [1] 3
```


```{r}

## why
ans  <- quote(x + y + 10)
substitute(ans, list(x=2))
# ans
y  <- NULL
eval(ans)
```


#### substitute, inside function retrieve value user sent
  *  deparse as quoted; character string 
```{r substitute}

# =========================
#deparse, substitute KEEP
# =========================
f  <- function(x)	substitute(x)
g   <-function(x) deparse(substitute(x)) 

f(unknown) # returns unquoted
# unknown
g(unknown) # returns in quotes, as char[1]
# [1] "unknown"


                      
f(a + b + c)
g(a + b + c)

# quite differnt
typeof(f(a + b + c))
typeof(g(a + b + c))
```

#### bquote,  partial evaluation in expression
```{r bquote}
##  bquote(expr, where=, slice)
  a =2
  quote(a == a)
  bquote(a == a)

## partial: evaluat .( ) 
  bquote(a == .(a))
  
  substitute(a == A)
  
##    If it is a list, the equivalent of list2env(x, parent = emptyenv()) is returned    
  substitute(expr = a == A, env = list(A = a))
  
##  Check environment  
  z = as.environment(list(A = a))
  z
  ls(envir = z)
  ls.str(envir = z)
  parent.env(z) 
#   >   z
# <environment: 0x55d7743de420>
# >   ls(envir = z)
# [1] "A"
# >   ls.str(envir = z)
# A :  num 2
# >   parent.env(z) 
# <environment: R_EmptyEnv>
 
  substitute(expr = a == A, env = list(A = a))
  
```


#### Compare sustitute vs bquote
#### !so 63231060, bquote return call with only partial evaluation, not
expression!
```{r partial}

## accepted ans (substitute)
    a <- 2
    b <- 1
    y = substitute(expression(b + a),list(b=1))
    # expression(1 + a)

typeof(y)
# [1] "language"


y[[1]]
# expression

is.call(y)
# [1] TRUE

is.expression(y)  # what?
# [1] FALSE

eval(y)
# expression(1 + a)

## another way, nope, not an expression
    a  <- 2
    b  <- 1
    x  <- bquote(.(b) + a)
    # 1 + a

    typeof(x)
    # [1] "language"

    is.pairlist(x)
# [1] FALSE

    x[[1]]
# `+`
    x[[2]]
# [1] 1

    is.call(x)
    # [1] TRUE

    is.expression(x)
    # [1] FALSE

    eval(x)
# [1] 3

    
```


```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
