--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  4 
fontsize: 10pt 
geometry: margin=0.4in,top=0.25in 
TAGS:  dput
---

####  For text (non-binary) storage:  use dput/dget or dump/source (muliple
obj)  

	*		Use dput() outputs to console a structure or something easy cut & paste.
  *  dput() returns a R object,  portable.  REF:  see Peng Chapter 7  
  *  dump: for multipe objects/source()  

```{r library, include=FALSE		}
``` 

#### EXAMPLE:   dput()  

```{r dput1}
y  <- list(a=1, b=2)
dput(y)

# keeps meta data
dput(mtcars[1:2,])
```

#### EXAMPLE:  dput simple data.frame  

```{r dput2}

test <- data.frame(x = 1, y = 1)
dput(test)


## probe
# structure
output <- dput(test) # output is to console
output
typeof(output)  # list

# write structure to file
if (FALSE )
  dput(test, "junk.txt")

```

#### EXAMPLE:  capture.output 
```
     Evaluates its arguments with the
     output being returned as a character
     string or sent to a file.  Related to
```

```{r capture_output}
# capture.output()

v  <- capture.output(dput(test))
v  
typeof(v) # char[]


# collapse to single ch[]
paste0(v, collapse="/n")


# function?
# as chr[] vector
capture.output(dput(paste0))

# single chr[]
s  <- paste0(capture.output(dput(paste0)), collapse = "/n")
s

```

#### save structure as expression, expr()
  *  called a defused expression, like blueprint
  *  so prevents execution of code
  *  see help for `expr`
```{r dput_expr}
z  <- rlang::expr(dput(test))
z
typeof(z)

c  <- eval(z)
```

```{r knit_exit()} 
knitr::knit_exit() 
```

```{r render, eval=FALSE, include = F	} 
file <- "BASE/0200_base_dput_structure.Rmd"
file  <- basename(file) 

dir <-"BASE/"
file  <- paste0(dir, file)
file

rmarkdown::render(file, 
       output_format="pdf_document",
       output_dir = "~/Downloads/print_and_delete"
       )


```
