---
---

PURPOSE:    more examples of do.call()
TODO
2023-06-02
-   more do.call examples (from r-bloggers this REF)
-   show how do.call can be constructed, manipulated  (or better with  call object)

REF: https://www.r-bloggers.com/2023/05/the-do-call-function-in-r-unlocking-efficiency-and-flexibility/
### match.call , in real life
  *  use case:
  *  do.call(what, args), where `what` is function, args is list
  *  here:  match.call()[-1] drops 1st element (function) 

##  match.call() returns call (only allowed inside function)
```{r begin}
f <- function(x, y, z) {
    z = as.list(match.call())
    cat(unlist(z[-1]), "\n")   ## args only

  do.call("sum", as.list(match.call()[-1])) 
}

f(1, 2, 3)
```




rbind a list  of atomic
Try with df, explain as written NOT a df
```{r}
vectors <- list(c(1, 2, 3), c(4, 5, 6), c(7, 8, 9))   ## list   
combined_matrix <- do.call(rbind, vectors)            ## 

##  NOTE
    class(combined_matrix) # [1] "matrix" "array" 
    typeof(combined_matrix) # [1] "double"


combined_matrix
```



