--- 
title: "295_base_assign_get.qmd"
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  4 
fontsize: 10pt 
geometry: margin=0.4in,top=0.25in 
TAGS:  get, assign 
---

```{r}
?assign
?get
```

assign binds a value to variable (as string) in chosen envirnment
```{r}
rm(list = ls())

# binds x  to value `hello`
assign("x", "hello")
x
assign("x", "bye")
ls()


get() returns value of named object (as string or symbol)
```{r}
x  <- 5
get("x")


y  <-  2+1
get("y")   # 3
```


```r
x  <- 5
get("x")   # [1] 5

# these fail - explain
# note:   neither `x` nor `"x"` are objects ??
get(as.name(x))
get(as.character(x))

# these work
get(as.character(quote(x)))
get(quote(x))
```


```r
rm(list = ls())
f  <- function(x) {
  x}

f(2)   # 2

f(x=2) # 2


f(z=30) # error



# 1. What is actually passed to `f` ? `10`, `z <- 10`?
#

ret = f(z  <- 10) # nothing printed
z  # 10,   because evaluated in calling environment (globalenv)?
ret   # 10



a = 2
b = 3

ret  <- f(x  <- a + b)
x
ret

```
