--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
        mathjax: default
  pdf_document: 
    latex_engine: xelatex
fontsize: 11pt 
geometry: margin=0.5in,top=0.25in 
TAGS:  ifelse,switch,unname, normalizePath, path.expand
---

## base R only,  collect here
*   unlist
*   dput  
*   save

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=F) 
```

### assign
```{r}
assign("x", 5)                         # must be quotes (string) 
x
```

### Not so keen on \code{else if} construct.
```{r if}

# ======================
## if - else if - else
# ======================

f  <- function(x) {
  if (x <0) {cat("less than 0")}
	else if(x<1) {cat("less than 1")}
	else if(x<2) {cat("less than 2")}
	else if(x<3) {cat("less than 3")}
	else if(x<4) {cat("less than 4")}
	else {cat("4 or greater")}
}

x <- -1
f(x)
f(x <- x+1)
f(x  <- x+1)
f(x  <- x+1)
x
```

### But ifelse(boolean, true , false) is very handy. 
```{r ifelse}

#  
# ======================
##		ifelse
# ======================
ifelse(5>3, "yes", "no")

x  <- c(1,5,10)

# return vector of length(x) and each element 'yes' or 'no'
ifelse(x > 5, "yes", "no")  # 'no' 'no' 'yes'
```

### switch
```{r}

## as function, x must evaluate to integer or character
choose  <- function(x) {
  stopifnot(is.character(x))

    base::switch(x, 
             A = "first", 
             B = "second",
             C = "third"
)

}

## choose one
    x  <- 3
    x <- 100
    x  <- "B"

##
    choose(x)
```

#### I find this useful, based on parameter \code{type} do something with x 
```{r switch2}

choose2  <- function(x=NULL, type=character(1)){

  base::switch(type,
               video = paste0("print videos",length(x)),
               playlist = "print playlists",
               channel = "print channel")

}
x  <- NULL
choose2(x,"channel")
choose2(x,"playlist")
choose2(x, "video")


### in named vector, names are attributes 
```{r unname}
# ----------------------
## create named vector
# ----------------------
    x  <- 1:5
    # [1] 1 2 3 4 5

    names(x)  <- letters[1:5]
    x
    # a b c d e 
    # 1 2 3 4 5 

    attributes(x)


# ----------------------------------------
## now, remove the names: couple of ways
# ----------------------------------------

  x.1  <- unname(x)
# [1] 1 2 3 4 5


## another way to remove names
  x.2  <- x
  names(x.2)  <- NULL
  x.2
# [1] 1 2 3 4 5

## also
  x.3  <- as.numeric(x)
# [1] 1 2 3 4 5
```

### expand relative file name
```{r normalizePath}

file <- "BASE/0296_base_R_misc.Rmd"
normalizePath(file)
path.expand(file)
path.expand("~")
```

```

