---
---
*0502_lists_1_2_3.qmd*
##  TODO: 

####	PURPOSE:
-   Demonstrate ways counting helps lists !
-   REF: https://stackoverflow.com/a/49911554/5057251
##		-----------------------------------


```{r}
L <- list(1, c(2, 3), list(4, list(5)))

##  count number of elements
length(L) # [1] 3

##  capture elements 1, 2
L[[1]] # [1] 1
L[[2]] # [1] 2 3


##  flatten?
unlist(L) # [1] 1 2 3 4 5


## What will return ?
L[[3]]
# [[1]]
# [1] 4
#
# [[2]]
# [[2]][[1]]
# [1] 5
#
#
##  Correct, but not very helpful to me.  Try this:
L <- list(1, c(2, 3), list(4, list(5)))
L
# [[1]]
# [1] 1
#
# [[2]]
# [1] 2 3
#
# [[3]]
# [[3]][[1]]
# [1] 4
#
# [[3]][[2]]
# [[3]][[2]][[1]]
# [1] 5
#
#
dput(L)
# list(1, c(2, 3), list(4, list(5)))
# [[1]]
# [1] 1
#
# [[2]]
# [1] 2 3
#
# [[3]]
# [[3]][[1]]
# [1] 4
#
# [[3]][[2]]
# [[3]][[2]][[1]]
# [1] 5
#
#
dput(L[[3]]) # list(4, list(5))

##  dput makes much easier to follow, I think.
```


```{r}
dput(mtcars)
```


### LEGACY

##		-----------------------------------
##		-----------------------------------
```{r}
library(tibble)
library(constructive) # alternative to dput, experimental:  see  remotes::install_github("cynkra/constructive")
```

##	atomic, data.frame

```{r}
v <- letters[1:10] #  [1] "a" "b" "c" "d" "e" "f" "g" "h" "i" "j"
str(v) #  chr [1:10] "a" "b" "c" "d" "e" "f" "g" "h" "i" "j"


df <- data.frame(
  x = v,
  y = v,
  z = v
)

df
str(df)
# 'data.frame':	10 obs. of  3 variables:
#  $ x: chr  "a" "b" "c" "d" ...
#  $ y: chr  "a" "b" "c" "d" ...
#  $ z: chr  "a" "b" "c" "d" ...
```

##  Probe
```{r}
probe <- function(x, max.depth = 1, list.len = 3, vec.len = 1) str(x, max.depth, list.len, vec.len)

#   even simple gets messy
L <- list(1, c(2), list(3, list(4, list(5))))
probe(L)
probe(L, max.depth = 2)
probe(L, max.depth = 3)
```

##  dput (control = ??)

##  Pop off outer list
```{r}
L[1] # subset
# [[1]]
# [1] 1
#
dput(L[1]) # get suppress line2 , 3?
# list(1)
# [[1]]
# [1] 1
#
#
#
## easier way to see effect of [[]]
dput(L[[1]], control = NULL) # remove line 2
# 1
# [1] 1
#
#
#
dput(as.data.frame(L[[1]]))
# structure(list(`L[[1]]` = 1), row.names = c(NA, -1L), class = "data.frame")
#   L[[1]]
# 1      1
dput(tibble(L[[1]]))
# structure(list(`L[[1]]` = 1), class = c("tbl_df", "tbl", "data.frame"
# ), row.names = c(NA, -1L))
# # A tibble: 1 × 1
#   `L[[1]]`
#      <dbl>
# 1        1
```




##	df more complicated



###	Giant structure, Hadley
{
hadley_repos <- jsonlite::fromJSON("https://api.github.com/users/hadley/repos")
hadley_repos


#	learn 30 x 70; list.len is # of cols
str(hadley_repos, 
		list.len=5,
		max.level=1, 
		vec.len=1      
)

f(hadley_repos, 5)

length(hadley_repos)
# [1] 78
#
str(hadley_repos, 
		list.len=78,
		max.level=1, 
		vec.len=0      
)

X=hadley_repos$topics
length(X)
str(X,
		max.level=1,
		vec.len=1
		)


#### opensecrets (error)
###
apikey  <- Sys.getenv("API_KEY")
nj  <- jsonlite::fromJSON( 
paste0("http://www.opensecrets.org/api/?method=getLegislators&id=NJ&apikey=", 
       apikey, "&output=json"))
nj


###
###
devtools::install_github("jennybc/repurrrsive")
library(repurrrsive)
data(package = "repurrrsive") # lists in new tab
browseVignettes(package="repurrrsive") # none
help.search("repurrrsive") # opens help

library(purrr)
data(package = "purrr") # none
# browser opens!
browseVignettes(package="purrr") # none
help.search("purrr") # opens help

#	-------------------------------------------
###	root object
x  <- discog

## too huge
if (F) str(x)

###	# elements
str(x, max.level=0)
# List of 155

length(x)
# [1] 155
#
#
-------------------------------------------
####	poke, each is list of 5 elments
str(x[[1]])
str(x[[2]])
str(x[[155]])


#	function to help poke!
f  <- function(x, a,b,c) {
	str(x, 
			max.level=a,
			list.len=b, 
			vec.len=c
	)

}

#	
f(x, 0,0,0)
# List of 155


#	show all 155 elements, but .. listtle of each
f(x, 1,155,1)


# now include singletons (ie to level 2)
f(x, 2,155,1)


# examine 'base_information"  a list
#
f(x = x[[1]], 1, 3, 1)

#	all elements of this x[[1]], no further
f(x = x[[1]], 1, 5, 1)


#	all 5 elment of x[[1]], but 1 level deeper
f(x = x[[1]], 2, 5, 2)


f(x = x[[1]], 3, 5, 2)

f(x = x[[1]], 5, 5, 2)

-------------------------------------------


#	assemble list of lists
A = list("one", "two")
B = letters[1:5]
C = character()
D = matrix(1:9, ncol=3)

L3 = letters[1:3]
E = data.frame(item=1:10, value=sample(L3, 10, replace = TRUE))

X = list(A,B,C,D, E)
str(X)
str(X, max.level = 0)
str(X, max.level = 2, vec.len=1)

# assign something to 3rd element
X[[3]]  <- c("hello", "bue")

#	walk the tree, find 3 element then its 2nd element
X[[c(3,2)]]


