### /home/jim/code/try_things_here/BASE/604_matrix_pkg_update_HELPER.qmd

PURPOSE:
HELPER:   update packages, using matrix tools (and not dt)


TODO:
  *   fix comments
  *   LEGACY:   pick out useful

Updating R packages should be routine task.  But my setup has a little twist.
I have two libraries: `r .libPaths()` one for standard packages ("sys") and another directory for user-installed packages ("user").  
The twist is that the "sys" library is read-only. When R  tries to update a package in this directory, which it can't, R asks to use user directory, which I approve.

The result, however, is that a package in sys can also be present in user, but a newer version.

(There are many workarounds for this.  RStudio seems to handle the updates fine.  I could change the permissions on the *stdlib* directory, but I do not want to.)

The work-around is probably over the top, but I good  exercise in  comparing two sets of packages before upgrading.

Notation:
  *   lib_sys refers to .libPaths()[2] for stdlib
  *   lib_user refers to .libPaths()[1] for user

First, 
### currently installed, in each of two libraries
```{r}
lib_user = installed.packages(lib.loc = .libPaths()[[1]])
head(lib_user[, 1:3])
lib_sys = installed.packages(lib.loc = .libPaths()[[2]])
head(lib_sys[, 1:3])
```

### Note these are matrix objects and not data.frame
```{r}
is.matrix(lib_user)
is.matrix(lib_sys)

# expect no package overlap, but ...

lib_user[lib_user[, "Package"] %in% lib_sys[,1],1:2, drop=F]

# check
"class" %in% lib_user[, 1] && "class" %in% lib_sys[, 1]
```

### `old` in each of two libraries I have (user/sys) 
The function `old.packages` find the candidates for upgrades.   I say candidates because some may be duplicates.
```{r}
cand_user = old.packages( instPkgs = installed.packages(lib.loc = .libPaths()[[1]]))
cand_sys = old.packages( instPkgs = installed.packages(lib.loc = .libPaths()[[2]]))
cand_user
cand_sys
is.matrix(cand_sys)


## check:  update candidates should already be installed: 
## all cand_user,2 elements of lib_user,2? (or NULL)
cand_user[, 1] %in% lib_user[, 1]
cand_sys[, 1] %in% lib_sys[, 1]
```

### Update candidates in `user` can be updated immediately.
```{r}
!is.null(cand_user[,1])
if (!is.null(cand_user[, 1])){
update.packages(ask=F, old = cand_user )
}
```

### Candidates cand_sys in sys and need to be more careful.
Case I 
For those pkgs in candidates in `sys` and already installed in `user`  =>  
DO NOT Upgrade

```{r}
dim(cand_sys[cand_sys[, "Package"] %in% rownames(lib_user),, drop=F])

if( F) 
update.packages(old = cand_sys, instlib = .libPaths()[[1]]
)
```

Case II 
For those pkgs in candidates in `sys` but not installed in `user: (ie never were upgraded)
Upgrade 

```{r}
cand_sys[!cand_sys[, "Package"] %in% rownames(cand_sys),, drop=F]

# if zero rows, then none of the candidates from lib2 have been previously updated and installed in lib1
head(cand_sys[, 1:3])
knitr::knit_exit()
```

# --------------
#       LEGACY  - extract useful only
# --------------
### Examine candidate packages against what is currently installed.
```{r}

## Candidate
(cand_packages = z[, 1])
is.matrix(cand_packages)
is.character(cand_packages)

# check all installed 
ins = installed.packages()

# check candidate packages to see installed
ins[ins[, "Package"] %in% cand_packages, 1:2, drop=F]

# CASES: compare candidate to installed
# Any candidates in [1] and NOT in [2] .. upgrade
# Any candidates NOT in [1], but in [2], .. upgrade
# Any candidates in BOTH [1] and [2]  ... upgrade only [1]

z[, "LibPath"] == .libPaths()[[1]]  |
z[, "LibPath"] == .libPaths()[[2]],1:2,drop=F]

# candidates installed in both [1] and [2]
# "Matrix" in [2] and also [1], prior upgrade
z[
z[, "LibPath"] == .libPaths()[[1]]  &
z[, "LibPath"] == .libPaths()[[2]],1:2,drop=F]


# candidates installed in either [1] , not [2]
z[
z[, "LibPath"] == .libPaths()[[1]]  |
!z[, "LibPath"] == .libPaths()[[2]],1:2,drop=F]

# candidates installed in [2], but not [1]  (updade)
(u = z[
!z[, "LibPath"] == .libPaths()[[1]]  |
z[, "LibPath"] == .libPaths()[[2]],1:2,drop=F])

# in [2], but not in [1] update
update.packages(lib.loc = .libPaths()[[2]], 
                ask = F,    ## TRUE means asks for each pkg
                oldPkgs = u,         # ONLY these packages updated
                instlib = .libPaths()[[1]],       # gives WARNING that R_HOME not writable
                checkBuilt=T)

# check


is.matrix(
z[z[, "LibPath"] == .libPaths()[[2]],1:2, drop=F] 
  )
is.matrix(
z[z[, "LibPath"] == .libPaths()[[2]],1:2 ]
)


```
### installed.packages
```{r}

e = installed.packages()
is.matrix(e)

## Check any of these packages installed more than once?
e[e[, "Package"] %in% c("Matrix", "KernSmooth", "mgcv", "nlme", "spatial"),1:2, drop=F]
```



<!-- BEGIN HERE:  subset matrices, based on condition -->




### EXAMPLE:  Finally, subset installed
```{r}
x = installed.packages()
x

# matrix Subset 

res  <- x[x[, "Package"] == "yaml",,  drop=F ]
is.matrix(res)
rownames(res)
colnames(res)
attributes(res)

dimnames(res)
dimnames(res)[[1]]
dimnames(res)[[2]]
```

```{r}

# "gert" package installed on in user dir
(res  <- x[x[, "Package"] == "gert", , drop=F ])



```{r}
(res  <- x[x[, "Package"] == "class", , drop=F ])

## from these 2 rows, choose is sys
(res2  <- res[res[, "LibPath"] == .libPaths()[[2]], , drop=F])

## this is the row  to upgrade !   user
(res2  <- res[res[, "LibPath"] == .libPaths()[[1]], , drop=F])

# about 29, standard packages  in sys

(res  <- x[x[, "LibPath"] == .libPaths()[[2]], c("Package", "LibPath"), drop=F])

```








