---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: show
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:   
    latex_engine: xelatex  
    toc: true
    toc_depth:  2   
    fontsize: 11pt   
    geometry: margin=0.5in,top=0.25in   
---

<!-- 
code_folding: hide   to turn off

set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw

vim: to format all urls for md
s/https.*/[&]()/g
-->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = FALSE   ) # easier to read

file  <- knitr::current_input() 
library(jimTools)
```



plot poisson distribution
```{r dpois}
lambda = 1    # rate of 1 per unit
x  <- 0:10    # how many counts?
z  <- dpois(x = x, lambda = lambda)
plot(z, main ="Poisson Distribution: upto n=10 counts, lambda=1",
     ylab = "dpois",
     xlab = "counts, 1:n"
)

```

***


## Motivating Poisson:
  * REF  http://www.econometricsbysimulation.com/search?q=poisson
  * Suppose, on averge, receive 9 letters in mail each day.
  * And suppose, can model post office as poisson.
  * Then expect sd to be sqrt(9) and so expect actual number of letters to vary
    between 3 and 12  (2 sd).

## Begin with Binomial.
  * u = expected number of events (letters)
  * in N = number of draws

```{r}

set.seed(2021)
draws_approx  <- function(u, N) sum (rbinom(N, 1, u/N))

draws_approx(9, N=9)  # this says,  p = 1 for each draw

draws_approx(9, N=18) # here p = 1/2, like flipping a coin, so total heads will
vary.
```
## Redo above, using my notation:
*  p = prob of success (u/N)
*  u = expected number of successes (letters)
*  in N = number of draws
*  B = repeat with this many samples
```{r}

B   <- 1000
u   <- 9
N  <-  18


# expect to see an avergae of 9 (successes from 18 draws at p = 1/2)
# expect sd to be 3 (if poisson holds)
# double[B], each is success for N in 1 sample
x  <- rbinom(B, N, u/N)
mean(x)
sd(x)


# suppose N really much larger, ie prob = u/N  much smaller
# getting closer!
N  <- 100
x  <- rbinom(B, N, u/N)
mean(x)
sd(x)


# now N huge, p = u/n tiny
N  <- 10^6
x  <- rbinom(B, N, u/N)
mean(x)
sd(x)

hist(x)
boxplot(x)
plot(density(x))

```


## Using sample to estimate expect value and variance, compare to Poisson.


```{r knit_exit(), include = FALSE  , eval=FALSE } 
# knitr::knit_exit()
```

/newpage

```{r plot_density, eval=FALSE  }
density_object <- density(mpg$hwy)
pdf("~/Downloads/print_and_delete/out.pdf")
plot(density_object)
dev.off()
```

```{r plot_png_example, eval=FALSE  }
filename  <- "~/Downloads/print_and_delete/out.png"
png(filename = filename,
         width = 480, height = 480, units = "px", pointsize = 12)
plot(density_object)
dev.off()
```
```{r render, eval=FALSE, include=FALSE 	} 
# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root

file  <- here("", "")
file  <- "rmd/01030_poisson_distribution.Rmd"

# in general, pdf will look nicer
rmarkdown::render(file,
                  #output_format = "pdf_document",
                  output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
