<!--
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
-->

--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

* Checking normality, qq, REF Hannay
   [REF:  hannay](https://faculty.nps.edu/rbassett/_book/continuous-random-variables.html)

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

```{r library, include=FALSE		}
library(jimTools) 
library(tibble)
library(ggplot2)
``` 

Ch 9.1.1 - fake population 
```{r one}
set.seed(2020)
heights<-rnorm(1200, mean=5.5, sd=4/12.0)
mean(heights)  # 5.49

hist(
     heights, 
     col='coral', 
     main='Heights of Schreiner Students', 
     ylab='Proportion of Students', 
     xlab='Height (ft)', 
     freq=FALSE)

```
Be lazy, sample N=50
```{r}
set.seed(1942)
N=50
one_sample_x  <- sample(heights, replace = FALSE  , size=N)

mean(one_sample_x)  #5.53 
# this just sd of sample, NOT sd of means
sd(one_sample_x)   # 0.327  

```

* Sample average is an `estimator`, is it unbiased? (Chapter 9.3)
* compare to theory, E[sample avg]
* replicate()
```{r estimator}
# take B samples (large)
B <- 10000
sample_mean  <- function(N=50){
  mean(sample(heights, replace=FALSE , size=N))
}

# store is set of mean heights
store  <- replicate(B, sample_mean(N))
mean(store) # 5.49
sd(store)  # 0.0486 (vs 0.047 - theory)_

theory_sd  <- sqrt( (1/3)^2/N )
theory_sd   # 0.0471
hist(
     store, 
     col='coral', 
     xlab='Sample mean', 
     main='Distribution of Sample Means N=50')

# scale
hist(
     scale(store), 
     col='coral', 
     xlab='Sample mean', 
     main='Distribution of Sample Means N=50')
```

Ch 9.4, qq & normality?
```{r normality}
IQR(store)/sd(store)  # 1.35 (vs 1.3)

qqnorm(store)
qqline(store, col="red")


# What percent are below 5.3, using as above mean= 5.5, sd=4/12
pnorm(q=c(5.3),5.5, 1/3)    #27%
```



```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "rmd/099_hannay_chapter_09_sampling.Rmd" 
file  <- basename(file) 
dir <- "rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
