	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
# Spaces, no Tabs
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---


* Checking normality, qq, REF Hannay
   [REF:  hannay](https://faculty.nps.edu/rbassett/_book/continuous-random-variables.html)
<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw
-->

<!--  commments -->

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools) 
library(tibble)
library(ggplot2)
``` 


Create SAMPLE
```{r assess_normality}
# REF:  hannay https://faculty.nps.edu/rbassett/_book/continuous-random-variables.html
# Chapter 8.5 Assessing Normality

set.seed(2020)
B  <- 1000
fake.data  <- rnorm(B) # B points

# B bins
# exclude 0, 1 (z not finite)
seq(1/(B+1), B/(B+1), 1/(B+1))
```
```{r histogram}

# histogram

```

Some ratios, fake.data
```{r ratios}
sd(fake.data) # 1.12 (vs 1.0 theory)

# Given area to left, what is quantile (value for x)
quantile(fake.data, 1/4)
quantile(fake.data, 1/2)
quantile(fake.data, 3/4)

# find IQR
quantile(fake.data, 3/4) - quantile(fake.data, 1/4)
IQR(fake.data) # 1.3 (vs 1.3 theory

# ratio
IQR(fake.data)/sd(fake.data)   # 1.16 (vs 1.3 theory)
```


scale, repeat for scaled fake.data
```{r scale}
z  <- scale(fake.data)
sd(z)

# quantiles
quantile(z, 1/4)
quantile(z, 1/2)
quantile(z, 3/4)



# find IRQ
quantile(z,probs=c(3/4)) - quantile(z, probs=c(1/4)) 
IQR(z)
# (theory) because (look at CDF-1 graph)
qnorm(p = c(3/4)) - qnorm(p = c(1/4))


# ratio
IQR(z)/sd(z)
```

qq
```{r plot}
t  <- tibble(fake.data = fake.data)

g  <- ggplot(t, aes(sample=fake.data) ) +
  stat_qq() +
  stat_qq_line()

# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")
```

TODO:  
Follow Hannay (above)
But replace with UNFIMR distr
```{r quantile}

# 100 points from UNIFORM DISTR, 0-1
set.seed(1492)
n  <-20 
t  <- 30*tibble(sample = sort(runif(n))  )

# Find sample quantiles corresponding to 1/20, 2/20 % ... 20/20 %  etc
probs  <-  seq(0 +1/n, 1, 1/n)

# corresponding pts, % points to left
x  <- quantile(t$sample, probs=probs)
x
t  <- t %>% mutate(quantiles = x, percent=probs)
t

# PLOT just sample points
g  <- ggplot(t, aes(x=sample,y=1) ) +
geom_point()
# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")

# PLOT quantiles corres to specific percentages
s  <- t %>% filter(percent == c(0.25, .50, .75, 1.00))


## NEXT Plot both sample data AND indicate quantiles.
g  <- ggplot(s, aes(x=quantiles,y=0, color="blue") ) +
geom_point() +
geom_point(aes(x=t$sample, y=1)) + 
# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")

```

```{r qq_plot}
g  <- ggplot(mtcars, aes(sample=mpg) ) +
	stat_qq() +
	stat_qq_line()
# save last plot
ggsave("~/Downloads/print_and_delete/qq.png", device = "png")


# 32 x 1
t  <- tibble(mpg = mtcars$mpg)
q_sample  <- quantile(mtcars$mpg, probs = seq(0, 1, 0.05), na.rm = FALSE,
              names = TRUE, type = 7 )

p_normal  <- pnorm(seq(-2, 2., 0.05))
p_normal
q_normal  <- qnorm(seq(0,1,0.05))

q_sample  <- quantile(mtcars$mpg, probs = seq(0, 1, 0.05), na.rm = FALSE,
              names = TRUE, type = 7 )
q_normal

```
***

```{r knit_exit()} 
knitr::knit_exit() 
```

/newpage

```{r render, eval=FALSE	} 
file <- "" 
file  <- basename(file) 
dir <-"rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
