
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
--- 
title: Template for .Rmd 
date: "`r paste('last updated', 
    format(lubridate::now(), '%H:%M, %d %B %Y'))`"
output: 
  pdf_document: 
    latex_engine: xelatex
toc: TRUE 
toc_depth:  1 
fontsize: 12pt 
geometry: margin=0.5in,top=0.25in 
---

<!-- 
set cul   "cursorline
cc=+1			"colorcolumn is 1 more than tw

vim: to format all urls for md
s/https.*/[&]()/g
-->

Simple Linear Regression
- REF:  Rafael: https://www.youtube.com/watch?v=3KFwJd5-JD8
- use basic plot(), abline() etc.

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "      ##",
                      error = TRUE,
                      collapse = TRUE)
```

<!--  RENDER:  see bottom -->

```{r library, include=FALSE		}
library(jimTools)
``` 

Data
```{r begin}
set.seed(2021)
x  <- 1:10
e  <- rnorm(10, mean = 0, sd = 1)
e
y  <- 10 * x + 2 * e

y
# if not X11 use:

pdf("~/Downloads/print_and_delete/01_data.pdf")
plot(x, y, col=1)
dev.off()

```

Guess slopes, plot
```{r}

y_hat  <- 9.8 * x
y_hat

pdf("~/Downloads/print_and_delete/02_guess.pdf")
plot(x, y, col = 2 )
abline(a = 0, b=9.8)
abline(a = 0, b = 8.0, col=3)
dev.off()
```

residuals, with guess
```{r residuals}
r  <-  y - 9.8 * x
r

pdf("~/Downloads/print_and_delete/03_residuals.pdf")
plot(x, r, col = 2, ylim = c(-5, 5))
abline(h = 0)
dev.off()

```
fit:  Use lm, no y-intercetp, residuals better
```{r}
fit  <-  lm (y ~x - 1)

y_hat  <- fit$coef * x
y_hat

r  <- y - y_hat 
r


pdf("~/Downloads/print_and_delete/04_lm_residuals.pdf")
plot(x, r, col = 2, ylim = c(-5, 5))
abline(h = 0)
dev.off()

```

TSS  = ESS + ModelSS
```{r sum_of_squares}
ess  <- sum(r^2)
ess


modelss  <-  sum(y_hat ^ 2)
modelss


ess + modelss

totalss  <- sum(y ^ 2)
totalss


# R^2
modelss / totalss   # 0.999

# compare to fit
summary(fit)        # 0.999

```





***

```{r knit_exit()} 
knitr::knit_exit()
```

/newpage

```{r render, eval=FALSE	} 
file <- ""
file  <- basename(file)
dir <- "rmd"

jimTools::ren_pdf(file,dir)
jimTools::ren_github(file, dir)
```
