---
title:  "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: hide
        toc: true 
        toc_depth: 2
        toc_float: true
        highlight: pygments
        theme: cerulean
        mathjax: default
editor_options: 
  chunk_output_type: console
TAGS: sampling, rep, sample, replicate, lapply
---


<!--
	 vim:linebreak:nowrap:nospell:cul tw=78 fo=tqlnr foldcolumn=3 
-->
### PURPOSE: Basic sampling, 
* Verbs:   sample, rep, repetition
* Collect all examples HERE
* [source Rafal Chapter 14:](https://rafalab.github.io/dsbook/probability.html#discrete-probability)

```{r setup, include=FALSE		}
knitr::opts_chunk$set(echo = TRUE,  
											comment="      ##",  
											error=TRUE, 
											collapse=TRUE) 
```

#### replicate vector
```{r rep}
# return a vector, containing x  3 times (3 elments)
rep(x=5,times = 3)  # ( 5 5 5	)x 3

# return one vector containing "a" "z"  6 times (12 elements)
v <- rep(x=letters[c(1,26)], times=6)
v

# be sure understand this patternpattern:
v  <- rep (c(letters[1:3]), c(3,2,1))
v

```

## generate seq
```{r seq}
head(seq(from = 1, to = 10, by=0.1))

x <- seq(1:4)
x

rep(x, 3) # 1...10 1 ...10  1...10
```

## take a sample, returns vector
*  size = repeat this many times (length of return vector)
```{r sample}
sample(x=c(0,1), size= 7, replace=TRUE)

sample(x=c("blue","red"), size=10, replace=TRUE, prob=c(0.8,0.2))

sample(x=1:5, size=4,  replace=TRUE)

# note:
m <- matrix(data=1:100, nrow=25) # fills 1st column, then second ...
head(m)
```



#### replicate, apply verb repeatedly
* HOWEVER, replicate is wrapper for sapply, which tries to `simplify`
* n = repeat number of times
* expr = ususally a call, such as mean(rnorm(10)) to repeat
* return DEPENDS 
```{r replicate}
replicate(n =3 , mean(1:100))
replicate(n =3,  mean(rnorm(10)))

# STUDY:  return 3 samples, each sample of size=4 as a COLUMN, in matrix
replicate(n =3,  sample(c(0,1),size=4, replace=T)) 
knitr::knit_exit()
```

Construct list of vectors for lapply
  *  replicate is version of sapply
  *  simply=FALSE ; returns list of vectors
  *  DEFAULT would return array or matrix

```{r lapply }
n  <- 3
l  <- replicate(3, rnorm(5), simplify = FALSE )
l

lapply(l, mean)
```

### median
```{r median}
# From normal distribution, take N values at random.  Compute median. To
# generate the sampling distribution of the median,  repeat this B times.
#
# Compute point estimate and 95% CI for TRUE  median.  

N  <- 100    # 1 sample
B  <- 10^4 

#  Generate a list of length B, each element containing double of length N
l  <- replicate(B, rnorm(N), simplify = FALSE)
if (F) str(l)

# For each of the B elements of l, compute the median.
m  <- lapply(l, median)

if (F) length(m); str(m)

# the Sampling Distribution appears Normal
hist(unlist(m))

# What can we say about the properities of Sampling Distribution.
# Is it normal?   If so, can we compare its mean and standard error and compare
# to theory?
#

```

### range
```{r range}
# From normal distribution, take N values at random.  Compute range. To
# generate the sampling distribution of the range,  repeat this B times.
#
# Compute point estimate and 95% CI for TRUE  range.  

N  <- 100    # 1 sample
B  <- 10^4 

#  Generate a list of length B, each element containing double of length N
l  <- replicate(B, rnorm(N), simplify = FALSE)
if (F) str(l)

# For each of the B elements of l, compute the median.
m  <- lapply(l, range)
m.1 <- lapply(l, function(e) {r  <- range(e); r[2] - r[1]})


if (F) {length(m); str(m)}

# the Sampling Distribution appears Normal
hist(unlist(m))
hist(breaks=20, unlist(m.1))

# What can we say about the properities of Sampling Distribution.
# Is it normal?   If so, can we compare its mean and standard error and compare
# to theory?
#

```
### sine (careful with # repetitions)
```{r sine}
# From sine wave distribution (x=0 to x=pi), take N values at random.  Repeat this B times.
# Compute point estimate and 95% CI for TRUE  median.  
N  <-  100
B  <- 10^2
y  <- replicate (B, runif(N, min=0, max=1), simplify=FALSE)
length(y)
inv  <- function(y) (acos(1-2*y))
x  <- lapply(y, inv)

m  <- lapply(x, mean)

pi/2

hist(unlist(x))

# PDF = sin
# CDF = integral sin bet 0 and x 

CDF  <- function(x) {1/2*(1 - cos(x))}
curve(CDF, from = 0, to = pi)

# given y, solve for x
x  <- acos(1 - 2*y)
x
hist(x)

curve(sin)

N  <- 100  
B  <- 10^4 
l  <- replicate(B, rnorm(N), simplify = FALSE)

m  <- lapply(l, median)
hist(unlist(m))
```
LEGACY (below)  move to 1060_sample binom?
======
game
```{r game}
# game:  roll 2 dice, 7 or 11?
# construct matrix of 2 cols, so simulate throwing 2 die, 50x.

B  <- 50  # experiments (rows)
n  <- 2   # each experment, 2 dice

s <- sample(1:6,B*n, replace=TRUE) # returns vector
m <- matrix(s, nrow= B, ncol=n)
head(m)
```

replicate is wrapper to repeatedly apply sapply
```{r replicate2}
# replicate
```

```{r norm}

#### 0050_from_norm
df <- data.frame(matrix(rnorm(1000), ncol =10))
head(df)	

```
```{r coin flip1}
N  <- 10000
roll <- sample(x=c(-1,1),size=N,replace=TRUE, prob=c(0.5,0.5))
roll  <- as.integer(roll)
is_integer(roll)
is_vector(roll)# T
is_atomic(roll)
str(roll)


t  <- tibble(raw =roll,
	     cum_total = cumsum(raw))
print(t, n=20)
tail(t)
plot( x=1:N, y=t$cum_total)

```
```{r coin_flip}


#### 0060_coin_flip_running_avg
# 1/n converges slowly!
size <- 100000
t <- tibble(x=sample(x=c(1,0),size=size, replace = TRUE),
			run_avg = dplyr::cummean(x)
			) %>% 
	mutate(line = 1:size)

t %>% head()

t$run_avg %>% tail()
t %>% ggplot(aes(x=line, y=run_avg)) +
	geom_point()
	

# create 'urn' with 5 balls
beads <- rep(c("red", "blue"), times = c(2,3))
beads

# pick sample (of 1)
event <- sample(beads,size=1)
event

# many events
set.seed(2012)
B <- 10000
events <- replicate(B,sample(beads,size=1))
table(events)  #  5988 blue, 4012 blue
t <- tibble(events)
	
# shortcut...
events1 <- sample(beads, size = B, replace = TRUE)
table(events1)  # 5988, 4012

x <- rnorm(10000,0,1)
mean(x <= 0)  # cute,  returns area, percentile
```




```{r knit_exit(), eval = FALSE   } 
knitr::knit_exit() 
```

```{r render, eval=FALSE, include=FALSE 	} 

# TODO:  file is found when knitr runs (see above)

# file must be of form:
# dir/name_of_this_file    where dir is relative to project root
getwd()
{
file <- "1050_sampling_basic.Rmd" 
file  <- here("STATS", file)
file
}
# in general, pdf will look nicer
rmarkdown::render(file,
                  output_format = "pdf_document",
                  #output_format = "html_document",
                  output_dir =  "~/Downloads/print_and_delete/")
```
