---
title:  "`r knitr::current_input()`"
date: "`r lubridate::today()`"
output:
    html_document:
        toc: true 
        toc_depth: 3
        toc_float: true
font_size: 10
---
COIN FLIPS
https://askgoodquestions.blog/2019/10/07/14-how-confident-are-you-part-1/
TODO:
-   overestimating se ?
-   better line segments
-   Large B,  use histogram to depict # do not containe TRUE p

## What does 95% confident actually mean??

Return # heads in sample of 75, fair coin: 
`rbinom (n=1, size = 75, prob= 0.5)`:  ren string of yes/no, 1,0  (size # trials)

From this ONE sample, estimate the number of heads and expected 95% CI (2 st)

Repeat B times (large number), calc 2 standard deviations each time, and plot expect upper and lower.
Compare this Theory

For counts:  mean should be p*n, standard dev squrt(n*p*q)
For proportion:   p, sqrt(p*q/n) 
$$   $$

```{r BEGIN}
set.seed(1492)

## Take a Sample  (ie string of yes/no, 1,0)
r  <- sample(x = 0:1,
             size = 75,
             replace = TRUE, 
             prob = c(0.6, 0.4))
r
sum(r) #31
(p_bar = sum(r)/n) # [1] 0.4
```


rbinom may be easier
```{r}
t  <- tibble::tibble()
prob = 0.5
n = 1000
B = 10^4
n*prob # [1] 50
sqrt(prob*(1-prob)/n) # [1] 0.05
2*sqrt(prob*(1-prob)/n) # [1] 0.10   (2 sd)
.05*B   # 50

old  <- options(digits = 4)
one_sample  <- function(DUMMY = NULL) {  # 1 sample
    # p_hat is percentage heads 
    p_hat  <- rbinom(1, size=n,  prob = prob)/n

    ## And the range, for any single p_bar, standard error
    se_p_hat = sqrt( (p_hat * (1 - p_hat))/n)
    ## CI,   For 95% use = 1.96, OR
    L  <- p_hat - 1.96*se_p_hat
    U  <- p_hat + 1.96*se_p_hat
    contains_TRUE =  (L < prob)   & ( prob < U )
    new  <- tibble::tibble(p_hat = p_hat, L=L, U=U, contains_TRUE = contains_TRUE)
    t <<- rbind(t, new)
    invisible(t)
}


#   Run experiment many, many times (expected value)

for (i in 1:5) {
t  <- tibble::tibble()
z = lapply(1:B, one_sample)
cat(nrow(t[!t$contains_TRUE,]), "\n")          # expect 5 out 100 outside 95%CI 
}

nrow(t[t$p_hat > 0.50, ])
nrow(t[t$p_hat < 0.50, ])
nrow(t[t$p_hat ==  0.50, ])

# view
t
t[order(t$L), ]
nrow(t[!t$contains_TRUE,])

if (F) {
hist(t$p_hat)
hist(t$L)
hist(t$U)
}
```

## Plot sample estimate with error bars
```{r}
library(ggplot2)
head(t)
(t = t[order(-t$U), ])
t = cbind(t, n= as.integer(rownames(t)))
names(t)
head(t)
ggplot(t, aes(x = L,xend = U,y = n,   , yend =n    )  ) +
  geom_segment(arrow = arrow(length=unit(0.10,"cm"), ends="first", type = "closed")) +
  geom_line(x = .5)
```
```{r}
#   :w
https://ggplot2.tidyverse.org/reference/geom_segment.html
df2 <- expand.grid(
  lineend = c('round', 'butt', 'square'),
  linejoin = c('round', 'mitre', 'bevel'),
  stringsAsFactors = FALSE
)

df2 <- data.frame(df2, y = 1:9)
df2

ggplot(df2, aes(x = 1, y = y, xend = 2, yend = y, label = paste(lineend, linejoin))) +
  geom_segment(
#     lineend = df2$lineend, linejoin = df2$linejoin,
     size = 3, arrow = arrow(length = unit(0.3, "inches"))
  ) +
#  geom_text(hjust = 'outside', nudge_x = -0.2) +
  xlim(0.5, 2)
```


## plots

## Error Bars
```{r}
library(ggplot2)

# Define the top and bottom of the errorbars
p <- ggplot(t[order(t$U), ], aes(p_hat, 1:nrow(t)))
p

p +
  geom_point() +
  geom_errorbarh(aes(xmax = U , xmin = L , height = .1))


p +
  geom_point() +
  geom_errorbarh(aes(xmax = resp + se, xmin = resp - se, height = .1))

```



```{r plot}

file  <- "out"
png(filename = paste0("~/Downloads/print_and_delete/" , file,".png"),
         width = 480, height = 480, units = "px", pointsize = 12,
          bg = "white",  res = NA,
         type = c("cairo", "cairo-png", "Xlib", "quartz")
)

 hist(t$p_hat)
# boxplot(x, col="green", names=c("UrOx"))

stripchart(x = x, 
           ylim=c(0, 225),
           pch = 1, 
           cex = 0.5,
           frame = TRUE, 
           vertical = TRUE,
           main = "24-hour Ur Ox")

abline(h = 35, col=2)

# plot(x, fake, ylim = c(-5, 5), sub = "t-value = 21.5", type = "p")
# abline(h = fake_mean, col=2)
dev.off()
```


