---
editor_options: 
  chunk_output_type: console
---

PURPOSE:   Scratch area.   Goal is to annotate 70_ and hoist
#
#  0430_tidyr_firefox_bookmarks_WORKS.R
##  Compare to base::rapply
##  TODO
##      -   2023-05-31

```{r}
library(tibble)
library(jsonlite)
library(dplyr)
library(listviewer)
library(tidyr)
```   

```{r}
the_dir = "~/code/try_things_here/rrapply/bookmarks"
the_file <- paste0(the_dir, "/", "bookmarks-2023-05-24_FF.json")
the_dir
the_file

##  no attempt to simplify
##  x is a list of lists
x <- jsonlite::read_json(the_file)
#
```



# ----------------------------------------------------------------------------------------

### Attempt:  dput, str
### This approach is brute force, step by step into list.
### PROBLEM: dput
### Very helpful for seeing the structure, but easily becomes mess.
### tools like str and dput  easily recurse and for very large files can tie up machine.
```{r}
#   x is list of lenth 10, 9 are singletons, the 10th is list   
listviewer::jsonedit(x)     ## list(guid, title, ... children=list(...))

## 
dput(x[1:9])

# first 9, As a list:
# list(guid = "root________", title = "", index = 0L, dateAdded = 1684955179130000, 
#     lastModified = 1684985711947000, id = 1L, typeCode = 2L, 
#     type = "text/x-moz-place-container", root = "placesRoot")
#   
#   10th, is list of lists and recurses!
if (F) # x$children

length(x$children)
str(x$children, max.depth=1, list.len=4)


menu = x$children[[1]]
toolbar = x$children[[2]]
unfiled = x$children[[3]]
mobile = x$children[[4]]


dput(menu[1:9])
##  

## but this recurses
if (F) dput(x$children[1:9])

if (F) dput(x[[10]])
# [1] 4
listviewer::jsonedit(x$children) ## list of 4 ( list(..), list(..), list(..), list(..))
```


# ------------------------------------------------------------------
###  Much better is capture lists in the very top list
###  And use various tools  (tidyr, hoist etc  and NOT with lists)
# ------------------------------------------------------------------
```{r}
x1 <- x$children %>% tibble()
##  4 x 1 tibble, each cell: named list of 10 ellements
x1
#   .                
#   <list>           
# 1 <named list [10]>
# 2 <named list [10]>
# 3 <named list [10]>
# 4 <named list [10]>

##  each of 4 chlidren holds another list of 10 
str(x1, max.depth=1, list.len=4)

```

# ----------------------------------- SKIP-------------------------------------
## BRUTE FORCE, skip
```{r}

names(x1) <- "first"
x2 <- x1$first %>% tibble()
# # A tibble: 4 × 1
#   .                
#   <list>           
# 1 <named list [10]>
# 2 <named list [10]>
# 3 <named list [10]>
# 4 <named list [10]>
#
length(x2)   # 1 column
##  1st row of x2    
x2[[1]][[1]]
names(x2[[1]][[1]])   # children is list of list
str(x2[[1]][[1]], max.depth=1, list.len=9)   # stop just before children
# ----------------------------------- END SKIP-------------------------------------
#
x2 <- x1 %>% hoist(".", # pull out from list-column
    title = "title",
##    id = "id",
    children = "children"
)
x2
# # A tibble: 4 × 3
#   <chr>   <list>         <list>
#   title   children       .
# 1 menu    <list [87]>    <named list [8]>
# 2 toolbar <list [14]>    <named list [8]>
# 3 unfiled <list [1,196]> <named list [8]>
# 4 mobile  <list [212]>   <named list [8]>


x3 <- x2 %>% unnest_longer("children")
x3

names(x3)
unique(x3$title)

x4 <- x3 %>% hoist("children",
    uri = "uri",
    title1 = "title"
)

head(x4)
# # A tibble: 6 × 5
#   title uri                               title1 children     .
#   <chr> <chr>                             <chr>  <list>       <list>
# 1 menu  <NA>                              Mozil… <named list> <named list>
# 2 menu  https://github.com/jarun/Buku     jarun… <named list> <named list>
# 3 menu  https://github.com/r-lib/remotes  r-lib… <named list> <named list>
# 4 menu  https://usethis.r-lib.org/articl… pr gi… <named list> <named list>
# 5 menu  https://quantifyinghealth.com/in… Inter… <named list> <named list>
# 6 menu  https://support.apple.com/guide/… Locat… <named list> <named list>
x4 %>% dplyr::select(c("title", "uri", "title1"))
# # A tibble: 1,509 × 3
#    title uri                                                        title1
#    <chr> <chr>                                                      <chr>
#  1 menu  <NA>                                                       "Mozi…
#  2 menu  https://github.com/jarun/Buku                              "jaru…
#  3 menu  https://github.com/r-lib/remotes                           "r-li…
#  4 menu  https://usethis.r-lib.org/articles/pr-functions.html       "pr g…
#  5 menu  https://quantifyinghealth.com/interpret-logistic-regressi… "Inte…
#  6 menu  https://support.apple.com/guide/icloud/locate-a-device-mm… "Loca…
#  7 menu  https://bookdown.org/markhoff/social_network_analysis/     "R | …
#  8 menu  https://dplyr.tidyverse.org/articles/programming.html      "Env …
#  9 menu  https://dcl-prog.stanford.edu/tidy-eval-section.html       "12 I…
# 10 menu  https://www.meetup.com/twincitiesrug/                      "Twin…
# # ℹ 1,499 more rows
# # ℹ Use `print(n = ...)` to see more rows
x4 %>% select(c("title1"))
# # A tibble: 1,509 × 1
#    title1
#    <chr>
#  1 "Mozilla Firefox"
#  2 "jarun/buku: |  bookmark manager"
#  3 "r-lib/remotes: Install R packages from GitHub, GitLab, Bitbucket, git…
#  4 "pr github! r helpers"
#  5 "Interpret Logistic Regression Coefficients [For Beginners] – QUANTIFY…
#  6 "Locate a device in Find Devices on iCloud.com - Apple Support"
#  7 "R | book | Methods for Network Analysis"
#  8 "Env var   Or    Data var "
#  9 "12 Introduction | Functional Programming r tidyeval "
# 10 "Twin Cities R User Group (TCRUG) (Minneapolis, MN) | Meetup"
# # ℹ 1,499 more rows
# # ℹ Use `print(n = ...)` to see more rows
```
